### 2022-04-10 R/GBIF_ocurrence_data.Rmd
Alex fixed the code. I nee dto fix path names

---
title: "GBIF occurrences"
author: "Alex Bush"
date: "17/05/2021"
output: html_document
---

# Species Data

```{r}
rm(list=ls())

spp.list = read.csv("../q10_meta-master/data/species_list_Q10.csv")
spp      = unique(spp.list$species_full)

spp[71]  = "Enithares"                 # Switch 'Enithares sp' to whole genus
spp[73]  = "Euchemotrema leaii"        # Synonym: Euchemotrema leai
spp[133] = "Merlangius merlangus"      # Synonym: Merlanguis merlangus
spp[175] = "Pinctada mazatlanica"      # Synonym: Pinctada mazatlantica
spp[202] = "Hoplobatrachus tigerinus"  # Synonym: Rana tigrina
spp[203] = "Lithobates virgatipes"     # Synonym: Rana vergatipes
spp[239] = "Zoarces viviparus"         # Synonym: Zoarces viviparous

# Note:
# > Halozetes fulvus is a species of mite endemic to Marion Island, hence no records in GBIF

```

# DOWNLOAD DATA FROM GBIF

```{r }
library(rgbif)
library(scrubr)
library(maps)

for(i in 1:length(spp)){
  if( !file.exists(paste0("../Data/GBIF/",gsub(" ","_",spp[i]),"_occurrences.RData")) ){
    # Download GBIF occurrence data for this taxon
    gbif_data = occ_data(scientificName = spp[i], hasCoordinate = TRUE, limit = 20000)
    # Save
    save(gbif_data, file=paste0("../Data/GBIF/",gsub(" ","_",spp[i]),"_occurrences.RData"))
    # No of records available:
    print(paste0(spp[i],": ",gbif_data$meta$count))
    # If "Records found" is larger than "Records returned", need to increase the 'limit' argument above
    # See help(occ_data) for options and limitations
    rm(gbif_data)
  }
}

# Quite a few have more than 20k records so need to update the download:
xspp = c("Alligator mississippiensis","Amphiura filiformis","Anolis carolinensis","Asterias rubens","Barbus barbus","Bithynia tentaculata","Boreogadus saida","Carassius auratus","Carassius carassius","Caretta caretta","Chrysemys picta","Crocodylus porosus","Cyprinus carpio","Gadus macrocephalus","Gadus morhua","Gambusia holbrooki","Gasterosteus aculeatus","Lepomis cyanellus","Lepomis macrochirus","Lepomis megalotis","Limax maximus","Limecola balthica","Limnodynastes peronii","Lithobates clamitans","Micropterus salmoides","Myoxocephalus scorpius","Mytilus edulis","Notophthalmus viridescens","Oncorhynchus mykiss","Oncorhynchus tshawytscha","Ophiothrix fragilis","Osmerus mordax","Perca flavescens","Perca fluviatilis","Pimephales promelas","Pomoxis nigromaculatus","Porcellio scaber","Pseudopleuronectes americanus","Pungitius pungitius","Rhinella marina","Rutilus rutilus","Salmo trutta","Salvelinus fontinalis","Salvelinus namaycush","Scomber japonicus","Trachemys scripta","Vipera berus","Zoarces viviparus")

for(i in match(xspp,spp)){
  # Download GBIF occurrence data for this taxon (hard limit is 100000)
  gbif_data = occ_data(scientificName = spp[i], hasCoordinate = TRUE, limit = 100000)
  # Save
  save(gbif_data, file=paste0("../Data/GBIF/",gsub(" ","_",spp[i]),"_occurrences.RData"))
  # No of records available:
  print(paste0(spp[i],": ",gbif_data$meta$count," == ",nrow(gbif_data$data)))
  # If "Records found" is larger than "Records returned", need to increase the 'limit' argument above
  # See help(occ_data) for options and limitations
  rm(gbif_data)
}

```

# CLEAN RECORDS

```{r }
doi.list = list()

for(i in 1:length(spp)){
  # Load GBI occurrence file
  load(paste0("../Data/GBIF/",gsub(" ","_",spp[i]),"_occurrences.RData"))

  # get the columns that matter for mapping and cleaning the occurrence data:
  xcols = c("decimalLongitude","decimalLatitude",
            "individualCount","occurrenceStatus",
            "coordinateUncertaintyInMeters",
            "institutionCode") #"references"
  coords = gbif_data$data[,names(gbif_data$data)%in%xcols] 
  coords = data.frame("Species" = spp[i], coords)
  #head(coords)

  # Map the occurrence data:
  map("world", 
      xlim = range(coords$decimalLongitude), 
      ylim = range(coords$decimalLatitude))  
  points(coords[ , c("decimalLongitude", "decimalLatitude")],pch=16, col="red", cex=0.5)

  # Remove records of absence or zero-abundance (if any)
  if( c("individualCount") %in% names(gbif_data$data)  ){
    #sort(unique(coords$individualCount))   # notice if some points correspond to zero abundance
    if( any(coords$individualCount==0 & !is.na(coords$individualCount)) ){
      coords = coords[!which(coords$individualCount==0 & !is.na(coords$individualCount)),]
    }
  }

  # Check for different indications of "absent", which could be in different languages! and remember that R is case-sensitive
  if( (c("individualCount") %in% names(gbif_data$data)) & c("occurrenceStatus") %in% names(gbif_data$data)  ){
    #sort(unique(coords$occurrenceStatus))
    absence_rows = which(coords$individualCount==0 | 
                           coords$occurrenceStatus %in% c("absent","Absent","ABSENT","ausente","Ausente","AUSENTE"))
    if(length(absence_rows) > 0){ coords = coords[-absence_rows, ] ; print(paste0("Removing ",length(absence_rows)," rows of 'absence' data.")) }
  }

  # Functions from the 'scrubr' package identify records with lat/long issues
  if( c("coordinateUncertaintyInMeters") %in% names(gbif_data$data)  ){
    #nrow(coords)
    coords      = coord_incomplete(coord_imprecise(coord_impossible(coord_unlikely(coords))))
    #nrow(coords)
  }

  # Eliminate presences with reported coordinate uncertainty (location error, 
  # spatial resolution) larger than 1 km (as this will be the grain of 
  # observation for the climate data later):
  if( c("coordinateUncertaintyInMeters") %in% names(gbif_data$data)  ){
    coords = coord_uncertain(coords, coorduncertainityLimit = 1000)
    #nrow(coords)
  }

  # Add DOIs to list for eventual citation for all species:
  doi.list = c(doi.list, gbif_citation(gbif_data))

  # Save
  save(gbif_data, file=paste0("../Data/GBIF/",gsub(" ","_",spp[i]),"_occurrences.RData"))

  rm(gbif_data)
  print(i)
}

# Keep only unique DOI's
doi.list = unique(doi.list)
# Save DOIs
save(doi.list, file="../Data/GBIF/List_of_GBIF_data_DOIs.RData")

```

****

```{r}
gbif = data.frame("Species"          = NA,
                  "decimalLongitude" = NA,
                  "decimalLatitude"  = NA)

for(i in c(1:length(spp))[-95] ){
  load(paste0("../Data/GBIF/",gsub(" ","_",spp[i]),"_occurrences.RData"))
  gbif = rbind(gbif, data.frame("Species"=spp[i],gbif_data$data[,c("decimalLongitude","decimalLatitude")]))
  print(paste0(spp[i],"   :::    ",nrow(gbif_data$data)))

}
gbif         = gbif[-1,]
gbif         = gbif[!duplicated(gbif),]
gbif$Species = gsub(" ","_",gbif$Species)

# Save
write.csv(gbif, paste0("../Data/GEE/All_GBIF_occurrences.csv"), row.names = F)

#map("world")
#points(gbif[ ,c("decimalLongitude","decimalLatitude")],pch=16, col="red", cex=0.5)

```

```{r}
rm(list=ls())

gbif = read.csv(paste0("../Data/GBIF/All_GBIF_occurrences.csv"))
names(gbif)[2:3] = c("longitude","latitude")

spp.list = read.csv("../q10_meta-master/data/species_list_Q10.csv")
spp      = spp.list$species_full
spp[71]  = "Enithares"                 # Switch 'Enithares sp' to whole genus
spp[73]  = "Euchemotrema leaii"        # Synonym: Euchemotrema leai
spp[133] = "Merlangius merlangus"      # Synonym: Merlanguis merlangus
spp[175] = "Pinctada mazatlanica"      # Synonym: Pinctada mazatlantica
spp[202] = "Hoplobatrachus tigerinus"  # Synonym: Rana tigrina
spp[203] = "Lithobates virgatipes"     # Synonym: Rana vergatipes
spp[239] = "Zoarces viviparus"         # Synonym: Zoarces viviparous

widedat  = read.csv("../q10_meta-master/data/CombinedData_wide.csv")
widedat$spp_names = paste(widedat$genus,widedat$species)
habitat  = widedat$habitat[match(spp.list$species_full, gsub("_"," ",widedat$spp_names))]
# Change habitat of Shortnose Sturgeon (Acipenser brevirostrum) to freshwater
habitat[4] = "f"

gbif             = read.csv(paste0("../Data/GEE/All_GBIF_occurrences.csv"))
names(gbif)[2:3] = c("longitude","latitude")
gbif$habitat     = habitat[match(gbif$Species, gsub(" ","_",spp))]

# Subsets
gbif_terr        = gbif[gbif$habitat=="t",1:3] # Take terrestrial subset for MODIS LCM extraction
gbif_fresh       = gbif[gbif$habitat=="f",1:3] # Take freshwater subset (also for MODIS LCM extraction)
gbif_mar         = gbif[gbif$habitat=="m",1:3] # Take marine subset to extract NOAA SST data

gbif_terr$Fid    = c(1:nrow(gbif_terr))
gbif_fresh$Fid   = c(1:nrow(gbif_fresh))
gbif_mar$Fid     = c(1:nrow(gbif_mar))

write.csv(gbif_terr,  paste0("../Data/GBIF/Terrestrial_GBIF_occurrences.csv"), row.names = F)
write.csv(gbif_fresh, paste0("../Data/GBIF/Freshwater_GBIF_occurrences.csv"), row.names = F)
write.csv(gbif_mar,   paste0("../Data/GBIF/Marine_GBIF_occurrences.csv"), row.names = F)

```