### 2024-07-07 ms.qmd
Re-read and revised a few sections

---
title: "Limited physiological plasticity and reduced relative variance in physiological rates of ectotherm populations under climate change"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
format:
  docx:
    reference-doc: ./bib/template.docx
    output-file: "ms_revised.docx"
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
editor: 
  markdown: 
    wrap: 72
---

```{r, setup}
#| label: setup
#| echo: false
#| include: false
##################################
# Clean workspace
##################################
    rm(list=ls())

##################################
# Loading packages & Functions
##################################
    pacman::p_load(tidyverse,  metafor, flextable, raster, rasterVis,  patchwork, RColorBrewer, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, ggbeeswarm, magick, ggimage, gt, performance, ggthemes, mapproj)

#devtools::install_github(repo = "ErikKusch/KrigR")
## Function to produce colour palette
# Note that temp is stored as a factor of 10 C larger, so needs to be divided. Also, we can create a custom colour palette function using colorRampPallette
colfunc <- colorRampPalette(rev(brewer.pal(10,"RdBu")))
#test colfunc(10) will generate the exact Hex codes for 10 different colours can 10 to 100 and you get 100 colour hex codes. Very useful webpage: https://pjbartlein.github.io/REarthSysSci/rasterVis01.html

    # loading functions
    source("./R/functions.R")

  options(digits = 2, scipen = 3)
```

```{r, data_setup_summary}
#| label: data_setup_summary
#| echo: false
##########################
## Load processed datasets
##########################
data_wide <- read.csv("./output/data/data_final_wide.csv")
data_long <- read.csv("./output/data/data_final_long.csv")

data_long2 <- data_long %>% mutate(type = factor(type, levels = c("acute", "acclim")))

##########################
## Summarize data
##########################

 data_wide$lat <- as.numeric(data_wide$lat)
    new_papers <- data_wide[-grep("_o", data_wide$record_num),]
    old_papers <- data_wide[grep("_o", data_wide$record_num),]

    # habitat
      spp_habitat <- data_wide %>%
                      group_by(habitat) %>%
                      summarise(spp_n = length(unique(species_full)))
      tot_spp <- sum(spp_habitat[,2])

    # Taxa
      taxa_sum <- data_wide %>%  
              group_by(habitat, phylum, class) %>% 
              summarise(spp = length(unique(species_full)), n = n()) %>% data.frame()

      taxa_sum_long <- data_long2 %>% group_by(habitat, phylum, class) %>% 
              summarise(spp = length(unique(species_full)), n = n()) %>% data.frame()

    # Traits 
      traits_habitat <- data_long2 %>% filter(grepl("lnRR_Q10",name))  %>% 
                      group_by(habitat, trait_category) %>%
                      summarise(spp_n = length(unique(species_full)), k = n())
      traits_all <- data_long2 %>% filter(grepl("lnRR_Q10",name)) %>% 
                      group_by(trait_category) %>%   
                      summarise(spp_n = length(unique(species_full)), k = n())

      # Acclimation time vary across habitat?

      acclim_time <- data_wide %>% 
                      group_by(habitat) %>% 
                      summarise(mean = mean(acclim_period, na.rm = TRUE),
                                  sd = sd(acclim_period, na.rm = TRUE),
                                  n = n())
      test_m_t <- t.test(acclim_period ~ habitat, data = data_wide %>% filter(!habitat == "f"))
      test_f_t <- t.test(acclim_period ~ habitat, data = data_wide %>% filter(!habitat == "m"))
      test_f_m <- t.test(acclim_period ~ habitat, data = data_wide %>% filter(!habitat == "t"))

#########################
## Phylogeny
#########################
 ## Tree data
    tree_data <-  data_long %>%
                  group_by(species_full)                                                 %>%
                  mutate(n = length(habitat))                                            %>%
                  dplyr::select(species_full,  habitat, n)                               %>% 
                  distinct(species_full, habitat, n)
    #rownames(tree_data) <- tree_data$species_full
    #tree_data <- tree_data[,2:3]

 ## Load the phylogenetic tree
                tree <- read.tree("./tree/tree_binary_SN_Q10.tre")   
      tree$tip.label <- gsub("_", " ", tree$tip.label)
      tree$tip.label <- gsub("Enithares sp", "Enithares ciliata", tree$tip.label)

      # Do a tree check with original data
        #  tree_checks(data_long, tree, dataCol = "species_full", type = "checks")

##################################
# Tree checking and pruning
##################################
  # Have a look at species in data and tree. 
  tree <- tree_checks(data_long, tree, dataCol = "species_full", type = "prune")

# Check that taxa dropped
 # tree_checks(data_long, tree, dataCol = "species_full", type = "checks")

##################################
# Branch lengths and cor matrix
##################################
      # Now that we have the tree file, we need to use it to create a phylogenetic correlation matrix using Grafen's method. This assumes Brownian motion and is the best we can do without a molecular phylogeny. The topology of the tree should be fairly good. We will explore a few different options with p to determine the best power function
      tree_p0.5 <- compute.brlen(tree, method = "Grafen", power = 0.5) # Probably most sensible power
      tree_p0.6 <- compute.brlen(tree, method = "Grafen", power = 0.6) # Probably most sensible power
      tree_p0.7 <- compute.brlen(tree, method = "Grafen", power = 0.7) # Probably most sensible power
      tree_p0.8 <- compute.brlen(tree, method = "Grafen", power = 0.8) # Probably most sensible power
      tree_p0.9 <- compute.brlen(tree, method = "Grafen", power = 0.9) # Probably most sensible power
      tree_p1.0 <- compute.brlen(tree, method = "Grafen", power = 1.0) # Probably most sensible power

     # Create null nodes 
      tree_p0.5$node.label <- NULL # We don't use nodes and will cause a warning so create null node names(
      tree_p0.6$node.label <- NULL # We don't use nodes and will cause a warning so create null node names(
      tree_p0.7$node.label <- NULL # We don't use nodes and will cause a warning so create null node names
      tree_p0.8$node.label <- NULL # We don't use nodes and will cause a warning so create null node names
      tree_p0.9$node.label <- NULL # We don't use nodes and will cause a warning so create null node names
      tree_p1.0$node.label <- NULL # We don't use nodes and will cause a warning so create null node names
```

<p style="text-align: center;">

Daniel W.A. Noble^1^ $\ddagger$, Fonti Kar^2^, Alex Bush^4^, Frank
Seebacher^3^ $\dagger$, & Shinichi Nakagawa^2^ $\dagger$

</p>

## Affiliations:

^1^ Division of Ecology and Evolution, Research School of Biology, The
Australian National University, Canberra, ACT 2600, Australia
\
^2^ Ecology and Evolution Research Centre, School of Biological, Earth
and Environmental Sciences, University of New South Wales, Sydney, NSW,
Australia
\
^3^ School of Life and Environmental Sciences, University of Sydney,
Sydney, NSW 2006, Australia
\
^4^ Department of Biology, Lancaster University, Liverpool, UK
\
$\dagger$ contributed equally
\
$\ddagger$ corresponding author, daniel.noble\@anu.edu.au

**Running head**: Reduced physiological rate variation in ectotherms

## **Abstract**

Climate change is expected to result in warmer and more variable thermal
environments globally. Maintaining variability in physiological rates
and adjusting them in response to extreme temperatures (plasticity) is
essential for allowing populations to adapt to climate change. Yet,
incorporating both plasticity and changes in phenotypic variation when
predicting the impacts of climate change on populations is often
ignored. We developed effect size estimates that quantify how
physiological rate variance changes relative to the mean as temperature
increases, which we refer to as 'relative variance'. We show by
meta-analysing \>1900 effects from 226 species that relative variance in
physiological rates decreases at higher temperatures driven by the lower
proportional increase in physiological rate variance relative to the
mean. Freshwater ectotherms are capable of acclimating and have the
smallest reductions in relative variance. Marine organisms also
acclimated, but plasticity traded-off with a reduction in relative
variance in physiological rates at higher temperatures. Relative
variance reductions were particularly pronounced for terrestrial
ectotherms, and this coincided with a lack of capacity for acclimation,
highlighting the vulnerability of terrestrial ectotherms to climate
change. Our results show that, while there are weak acclimation
responses in some habitats (marine and freshwater), such responses may
trade-off with reductions in relative physiological rate variance that
may have important evolutionary and ecological ramifications that affect
our understanding of how climate change will impact populations now and
in the future.

## **Introduction**

Climate change is expected to result in warmer but also more variable
thermal environments globally [@Suarez2023; @Easterling2000;
@Ummenhofer2017]. Greater thermal variability is predicted to result in
strong selection pressures that lead to genetic adaptation and/or the
evolution of adaptive phenotypic plasticity -- both of which are
considered important for population resilience to contemporary
human-induced climate change [@Seebacher2015-wi; @Merila2014;
@Chevin2015; @Chevin2017; @Chevin2010; @Nunney2016; @seebacher2023can;
@cooke2021one]. Without plasticity or adaptation, high extinction rates
are expected unless organisms can migrate to track suitable habitats
[@Cahill2012; @Nunney2016].

Reversible phenotypic plasticity, such as physiological acclimation, is
relatively rapid and can be fine-tuned to environmental conditions
making it the first ‘line-of-defense’ against environmental change
[@Dewitt1998; @Scheiner2020]. For example, physiological rates are known
to speed up as temperature increases because of the thermodynamic
effects on chemical reaction rates -- so called 'acute' temperature
responses (@fig-concept). Acute temperature increases that push
organisms from their thermal optimum can be mitigated by plastically
adjusting reaction rates in response to chronic temperatures through
acclimation [@Seebacher2015-wi; @Havird2020-mj]. Physiological
acclimation is driven by endocrine and epigenetic processes that change
the underlying physiology to allow organisms to maintain physiological
performance around a fitness optimum despite changes in the environment
[@seebacher2019histone; @little2013thyroid; @taff2016endocrine].
Acclimation therefore alters acute thermal sensitivity to offset the
potentially negative effects of acute temperature changes (e.g., higher
energetic demands). Acclimation, however, does not normally result in
physiological rates returning to their original state (‘complete’
compensation - *sensu* @huey1999testing), but rather are only partially
compensated [@huey1999testing; @Havird2020-mj].

Reversible forms of phenotypic plasticity, such as acclimation, are
expected to evolve in populations when environmental variability is high
but predictable and the costs of plasticity are low [@Chevin2017;
@Scheiner2020; @Nunney2016; @Dewitt1998; @Reed2010]. Thermal variation
and predictability are known to differ across habitats (terrestrial,
marine and freshwater) [@steele2019comparison], and as such, we expect
organisms within these habitats to vary in their capacity for plastic
adjustments in physiology. Species occupying terrestrial habitats have
been shown to exhibit weak acclimation abilities, and so, are likely to
be particularly vulnerable to climate change given their greater
probability of experiencing thermal extremes that overwhelm
physiological homeostasis [@Seebacher2015-wi; @Gunderson2015-hz;
@hoffmann2013upper; @morley2019physiological]. In contrast, marine and
freshwater habitats appear to have greater physiological acclimation
capacity [e.g., @Seebacher2015-wi; @Pottier2022]. However, the focus of
research up to now has been primarily on mean physiological responses
neglecting how variability in physiological rates might also be impacted
at higher temperatures.

As mean physiological rates increase with temperature there is reason to
believe that the intrapopulation variability might also be impacted.
NEED MORE ABOUT WHY VARIANCE IS EXPECTED TO CHANGE!

Changes in variability of traits governing energy demand are expected to
have important consequences on the flow of energy within and between
populations, communities, and ecosystems [@hendry2016key;
@seebacher2023can; @barneche2021warming; @Sanderson2023; @Bolnick2011].
More variable populations, genetically and/or phenotypically, are
predicted to be associated with broader niches, reduced intraspecific
competition, increased growth rate, decreased vulnerability to
environmental change and lower extinction risk [@forsman2015rethinking;
@forsman2014effects; @Bolnick2011; @hendry2016key; @hart2016variation].
In addition, if phenotypic and genetic variation are correlated and
traits are linked to fitness, changes in variation can limit responses
to selection impacting a populations capacity to evolve
[@Pelletier2012]. Therefore, maintaining intrapopulation variability in
physiological rates in a warmer world may be important for population
resilience to climate change.

Here, we use meta-analysis to re-evaluate the degree to which aquatic
and terrestrial ectotherms are capable of physiological plasticity. We
then developed new effect sizes to quantify how variance in
physiological rates change with temperature to ask the following
questions: 1) How much is variance in physiological rates expected to
change, if at all, as temperatures rise? 2) Are temperature effects on
plastic adjustments in physiological rates larger than changes in
variance across aquatic and terrestrial ectotherms? 3) How does
acclimation and changes in variance relate to different life-stages and
types of traits, and does this depend on habitat? 4) Are changes in
plasticity or variance in physiological rates impacted by past climate
history? 5) How are means and variances in physiological rates expected
to change under climate change?

## **Materials and Methods**

### *Literature collection*

We compiled literature on ectothermic animals that measured
physiological rates (e.g., metabolic rate) at two or more temperatures
after having been acclimated (or acclimatized) at these temperatures. We
used data from a previous meta-analysis [@Seebacher2015-wi] and updated
@Seebacher2015-wi's data by extracting data from suitable studies from
our own searches that followed the same search protocol. For full
details on the search protocol see the *Supplementary Materials*, where
we also provide a PRISMA flow diagram of our extraction process
(@fig-prisma).

### *Data Compilation*

We extracted means, standard deviations, and sample sizes for
physiological rates at the two test temperatures. If there were more
than two test temperatures, we chose only the test temperatures that
fell within the most likely natural range of temperatures experienced by
the species in question. We extracted these data from text, tables or
figures of a given paper. Data were extracted from figures using the R
package *metaDigitise* [@Pick2019-mg]. We also recorded the phylum,
class, order, genus and species, and the latitude and longitude from
where the experimental animals were sourced. For studies that did not
provide latitude and longitude for the population, we searched for
similar studies by the same lab group to identify where the population
was likely to have been sourced. If the experimental animals were
derived from the wild, we recorded the nearest latitude and longitude of
the field collection site. If the animals were sourced from a commercial
supplier, we took the latitude and longitude of the supplier. When it
was not possible to find latitude and longitude using these methods, we
looked up the distribution of the species in question and took the
latitude and longitude of the centroid of the species' distributional
range.

### $Q_{10}$ Based Effect Sizes and Sampling Variances for Means and Variances

Following @Noble2022-ty we calculated a series of temperature-corrected
effect sizes that compared mean physiological rates ($lnRR_{Q_{10}}$) as
well as the variability in physiological rates ($lnVR_{Q_{10}}$ and
$lnCVR_{Q_{10}}$) (@fig-concept). These effect sizes are similar to the
traditional temperature coefficient ($Q_{10}$), but with formal
analytical approximations of their sampling variances. Sampling
variances for effect sizes allowed us to make use of traditional
meta-analytic modelling approaches.

```{r, fig-concept}
#| label: fig-concept
#| fig-width: 8
#| fig-height: 12
#| fig-cap: "Calculations of acute and acclimation effect sizes. (A) Two idealised thermal performance curves for animals held at 'cold' ('blue') temperatures and warm ('red') temperatures. Shaded blue and red areas are the thermal optima of the performance curves. Physiological rates are measured for a sample of ectotherms at two different temperatures along the thermal performance curves ($T_{1}$ = 20$^\\circ$C and $T_{2} = 30$^\\circ$C) for both curves. At each temperature a mean physiological rate (R) (points) and its standard deviation (SD) (error bars above and below mean) are estimated. R1.1 and R1.2 are the rates and associated SD (subscripted) for the cold acclimated animals at temperature 1 and 2, respectively. R2.1 and R2.2 are the rates and associated SD (subscripted) for the warm acclimated animals at temperature 1 and 2, respectively. An example of how acute and acclimation $lnCVR_{Q_{10}}$ are calculated from the treatments within the study is provided on the right-hand side of the figure with reference to each of the four possible groups. Note that, two acute effect sizes can be calculated, one for the cold acclimated animals and one for the warm acclimated animals. (B) Species are expected, *a priori*, to vary in their thermal performance curves (thin lines) around an average (thick black line). We restricted our data to areas of each species' performance curve that fell within the natural thermal range of the species (thick lines on each species-level curve). However, given it was not possible to measure the full performance curve for each species some test temperatures within studies may have converged on or moved past the thermal optima. In such cases, we expected our $Q_{10}$ effect sizes to be smaller as indicated by comparing the black dashed lines to grey dashed lines."

fig_concept <- magick::image_read("output/figures/fig1_concept_full.png")
plot(fig_concept)
```

#### *Comparing changes in mean physiological rates*

To compare mean physiological rates, we calculated the log $Q_{10}$
response ratio, $lnRR_{Q_{10}}$ [@Noble2022-ty] as follows:

$$
  lnRR_{Q_{10}} = ln\left( \frac{R_{2}}{R_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
$$ {#eq-lnq10}

Where, $R_{1}$ and $R_{2}$ are mean physiological rates at temperatures
$T_{1}$ and $T_{2}$, respectively. Log transformation of this ratio
makes the effect size normally distributed. @eq-lnq10 is essentially a
temperature corrected equivalent to the log response ratio (lnRR)
[@Hedges1999-de; @Lajeunesse2011-fq] when the numerator and denominator
are measured at different temperatures. This allows comparisons of the
means from two temperature treatments directly regardless of the
absolute measurement temperatures. The sampling variance for @eq-lnq10
can be computed as follows (as described in @Noble2022-ty):

$$
  s_{lnRR_{Q_{10}}} = \left( \frac{SD_{2}^2}{R^2_{2}N_{2}} + \frac{SD_{1}^2}{R^2_{1}N_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }^2
$$ {#eq-Vlnq10}

Here, $SD_{1}^2$ and $SD_{2}^2$ are the standard deviations, and $N_{1}$
and $N_{2}$ are the sample sizes, of the groups measured at $T_{1}$ and
$T_{2}$, respectively (@fig-concept A).

#### *Comparing variance in physiological rates*

@Nakagawa2015-fo proposed analogous effect size estimates to *lnRR* that
allow for comparisons of changes in variance between two groups, the log
variance ratio (*lnVR*) and the log coefficient of variation (*lnCVR*).
*lnVR* and *lnCVR* are ratios that describe the relative difference in
trait variability between two groups. We refer readers to
@Nakagawa2015-fo for the equations describing *lnVR* and *lnCVR*, but
these can easily be extended to their $Q_{10}$ analogues (and associated
sampling variance) as follows:

$$
lnVR_{Q_{10}} = ln\left( \frac{SD_{2}}{SD_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
$$ {#eq-lnq10VR} $$
s_{lnVR_{Q_{10}}} = \left( \frac{1}{2\left( N_{2}-1 \right)} + \frac{1}{2\left(N_{1} - 1\right)} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }^2
$$ {#eq-slnq10VR}

@eq-lnq10VR and @eq-slnq10VR describe the change in physiological rate
variance (@eq-lnq10VR) across a 10°C temperature change along with its
sampling variance (@eq-slnq10VR). While this is a useful metric, as
discussed by @Nakagawa2015-fo there is often a strong mean-variance
relationship that needs to be accounted for in analysing changes in
variance. As such, we calculated the coefficient of variation, which
standardizes changes in variance for changes in means as follows:

$$
lnCVR_{Q_{10}} = ln\left( \frac{\text{CV}_{2}}{\text{CV}_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
$$ {#eq-lnq10CVR} $$
s_{lnCVR_{Q_{10}}} = \left[ \frac{(\text{SD}_{1})^2}{N_{1}({R}_{1})^2} + \frac{(\text{SD}_{2})^2}{N_{2} ({R}_{2})^2} + \frac{1}{2\left( N_{1}-1 \right)} + \frac{1}{2\left(N_{2} - 1\right)} \right]{ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }^2
$$ {#eq-slncvr}

where $CV$ is the coefficient of variation defined as $SD / R$. We refer
to $lnCVR_{Q_{10}}$ as relative variance because variance changes are
relative to the mean.

#### *Calculating acute and acclimation* $lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates

$R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$ can all be taken from
samples of organisms measured acutely at two temperatures or after
having been acclimated these same temperatures (@fig-concept A). For
studies that measure acute and acclimated responses we used the mean,
standard deviation, and sample size to derive both acute and acclimation
$lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates. For
studies that only measured $R_{1}$, $R_{2}$, $SD_{1}^2$ and $SD_{2}^2$
acutely we could only calculate acute versions of these effect size
estimates. For all effect sizes the higher temperature was in the
numerator and the lower of the two temperatures in the denominator. As
such, positive effect sizes suggest that the mean (i.e.,
$lnRR_{Q_{10}}$) or relative variance (i.e., $lnCVR_{Q_{10}}$) is larger
at the higher of the two temperatures (numerator) when standardized to
10°C. Importantly, $lnRR_{Q_{10}}$, $lnCVR_{Q_{10}}$ and
$lnVR_{Q_{10}}$, as with $Q_{10}$ more generally, all assume that the
effect of temperature on physiological rates (or changes in relative
variance) is log-linear (see *Supplementary Materials* for further
discussion). We test and control for any violations of these assumptions
in our analysis (see below) ([@fig-concept]B).

### *Moderator Variables*

We recorded or derived a series of moderator variables from each study
that are expected to have an impact on our effect size estimates. This
included the duration of acclimation in days given that acclimation
responses are expected to depend on how long chronic temperature
exposure occurs (longer exposure = better acclimation response)
[@Seebacher2015-wi]. We also recorded if the sample of animals were
derived from captive or wild stocks, the life-history stage of the
animals used ("adult" or "juvenile") and the habitat type ("freshwater",
"marine" or "terrestrial") given that @Seebacher2015-wi show that these
factors can impact $Q_{10}$ estimates. Physiological rate measures
varied widely across the studies but could generally be grouped into
discrete trait categories [@Seebacher2015-wi]. While we explored
differences across more detailed trait categories (see *Supplemental
Materials* for details on the types of traits), most traits could be
categorised into two broad categories that included whole-organism
measures, which all integrate a diversity of physiological and
biochemical processes, and biochemical processes (e.g., enzyme reaction
rates, proton leak). Traits that could not be categorised into these two
we classified as ‘Other’.

### *Climate Data*

To understand how climate is related to a species’ physiological
acclimation abilities and changes in relative variance we used the
coordinates reported by each study to extract temperature data from
terrestrial and aquatic environments. It was unclear whether climate at
the locations of captive reared organisms would be representative of a
population's climate history - particularly for species reared under
captive condition for many generations. Given that we were interested in
understanding climate driven effects on acclimation capacity we only
used studies on wild populations for climate analyses.

Monthly average temperature data were extracted from the ERA5 climate
model, available from the Copernicus climate data store
[@Hersbach2020-hx]. For each population and species in the dataset we
extracted a 72-year period (1950-2022) of either surface air temperature
(0.01° resolution) for both terrestrial and freshwater taxa, or sea
surface temperature for the marine taxa (at 0.25° resolution) using the
*ncdf4* R package [vers. `r utils::packageVersion("ncdf4")`,
@Pierce2021-lc]. We chose surface temperature because we believed that
it was more likely to reflect the micro-thermal environment experienced
by terrestrial and freshwater ectotherms at those locations.

Using the thermal time-series data for each location we calculated
metrics of thermal variability across months and years as well as
estimates of thermal predictability (i.e., autocorrelation). To estimate
thermal variability, we calculated the coefficient of variation
($\frac{SD}{M}$, where SD = standard deviation in temperature and M =
the mean temperature for each year). We also estimated thermal
predictability, by calculating the auto-regressive time lag across
months (i.e., a measure of how correlated temperatures were between
months), however, identifying biologically relevant lags for such
diverse taxa is challenging. As such, we present a coarse analysis using
this metric of thermal predictability in the *Supplemental Materials*.

Lastly, to illustrate the effects that climate warming could have on
physiological rate variance we also extracted climate projections and
calculated thermal variability into the future. We used the CanESM2
climate model (2005-2100) [vers. 1.2.0; @koen_hufkens_2019_2647541]
under a high emissions scenario (RCP8.5).

### *Meta-Analysis*

We analysed our data using multilevel meta-analytic (MLMA) and
meta-regression (MLMR) models in R (vers.
`r paste0(R.Version()$major, ".", R.Version()$minor)`) using *brms*
[vers. `r utils::packageVersion("brms")` @Burkner2017-wg;
@Burkner2018-ur; @noauthor_2021-az] and *metafor* [vers.
`r utils::packageVersion("metafor")` @Viechtbauer2010-fn]. We fit both
Bayesian and frequentist approaches to ensure that our results were
consistent, and to create orchard plots more easily [vers. 2.0,
@Nakagawa2021-ls; @nakagawaorchard]. In addition, Bayesian methods
better protect against type I errors in the presence of complex sources
of non-independence [@Nakagawa2021-uf; @Song2021-ba; @Noble2017-mz]. In
all cases, frequentist and Bayesian models resulted in the same
conclusions. For our Bayesian models, we ran 4 MCMC chains, each with a
warm-up of 1000 followed by 4000 sampling iterations keeping every 5
iterations for a minimum of 3200 samples from the posterior
distribution. We used flat Gaussian priors for 'fixed' effects (i.e.,
$N(0, 10)$) and a student t-distribution for 'random' effects (i.e.,
$student_{t}(3, 0, 10)$). We checked that all MCMC chains were mixing
and had converged (i.e., $R_{hat} = 1$). We compared any competing
models using Akaike's Information Criteria (AIC) (if frequentist) or
Wantabe Information Criteria (WAIC) (if Bayesian). We deemed models with
the lowest IC value to be best supported if there was a $\Delta IC$
between the competing models of 2 or more. If two models were within 2
$\Delta IC$ units we went with the most parsimonious model.

#### *Multi-level Meta-analysis (MLMA) Models*

We first fit multi-level meta-analysis (MLMA) models (i.e.,
intercept-only models) for both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$,
that included study, species, trait type, and phylogeny as random
effects to account for non-independence and identify sources of
variability. We refer to this model structure as “Model 1” in the
results. Our MLMA models allowed us to partition the variation in
$lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ among these key sources while
accounting for total sampling variance in each. This allowed us to
calculate the proportion of total heterogeneity [i.e., $I_{total}^2$;
*sensu* @Nakagawa2012-oc; @Noble2022-ty] along with various $I^2$
metrics describing the proportion of variance explained by each random
effect level [@Nakagawa2012-oc]. We also present 95% prediction
intervals which describe the expected distribution of effects from
future studies [@Nakagawa2021-ls; @Noble2022-ty].

A phylogeny was derived using the Open Tree of Life (OTL) with the
*rotl* package in R (vers. `r utils::packageVersion("rotl")`)
[@Michonneau2016-if], and plotted using *ggtree* (vers.
`r utils::packageVersion("ggtree")`) [@Yu2017-wj]. We resolved all
polytomies in the tree. Any missing taxa were replaced with closely
related species and branch lengths were computed using Grafen's method
[using power = 0.7, @Grafen1989-pr]. Models fit using correlation
matrices computed with different power (p) parameters (from 0.5 – 1.0)
had nearly identical $AIC_{c}$. As such, we used an intermediate value
of p = 0.7. We used the R packages *ape* (vers.
`r utils::packageVersion("ape")`) [@Paradis2019-gx] and *phytools*
(vers. `r utils::packageVersion("phytools")`) [@Revell2012-vq] to prune
the tree for individual analyses and calculate phylogenetic covariance
(or correlation) matrices used in meta-analytic models.

#### *Multi-level Meta-regression (MLMR) Models*

After quantifying levels of heterogeneity, we fit a series of
multi-level meta-regression (MLMR) models to test our key questions. In
all models, we included the same random effects as we used in our MLMA
models. Acclimation time varied from
`r range(data_wide$acclim_period, na.rm = TRUE)[1]` to
`r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean $\pm$ SD
= `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$
`r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial
ectotherms were acclimated for a much shorter duration (mean $\pm$ SD =
`r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]` days, n =
`r acclim_time[3,4]`) than both freshwater (mean $\pm$ SD =
`r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]` days, n =
`r acclim_time[1,4]`) and marine species (mean $\pm$ SD =
`r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]` days, n =
`r acclim_time[2,4]`). However, rates of acclimation have been shown to
be faster for many terrestrial groups compared to aquatic organisms
\[e.g., amphibians and reptiles have faster rates of acclimation than
fishes; See @Einum2023-lz\]. Nonetheless, to control for these possible
differences, acclimation time was mean-centered (mean = 0) and included
in all our models, though it was not a strong predictor of effect size
variation (*Supplementary Materials*, @fig-accl_time).

In addition to acclimation, all our models corrected for the possibility
of the log-linearity assumption of our effect sizes being violated
([@fig-concept]; and see Supplementary Materials @fig-lnrrq10_temp). We
predicted that, if $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ were not
strictly log-linear than we would see a decrease in average effect size
for studies applying higher temperature treatments because these
temperatures are expected to either converge on or cross the thermal
optima of the performance curve causing reaction rates to decelerate or
decrease beyond $T_{op}$ [@michaletz2024hotter]. The benefit of such an
approach means that we could still take advantage of the excellent
properties of $Q_{10}$ as an effect size while statistically correcting
for the potential non-linearity that would be expected in the data at
high treatment temperatures. Given that our data included a wide range
of species and habitats, we also included a random slope of maximum
temperature that varied across species because we, *a priori*, expected
that each species will vary in their thermal performance curve and the
chosen temperatures experimentally manipulated along the performance
curve. To improve model interpretation, we also mean-centered the
maximum temperature and included it in our models.

With acclimation time and maximum temperature as moderators in all our
models we proceeded to build separate models that tested our core
questions. All estimates from our models are therefore conditioned on an
average level of acclimation time (i.e.,
`r mean(data_wide$acclim_period, na.rm = TRUE)` days) and maximum
temperature (i.e., 23$^\circ C$) across the dataset.

We first tested the degree to which the difference in acute and
acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by
habitat type (i.e., terrestrial, freshwater, and marine). Models
included an interaction between effect type (i.e., acute or acclimation)
and habitat (referred to as “Model 2”). Reduced mean
$lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$
indicates that acclimation to thermal environments results in partial
compensation of physiological rates (i.e., phenotypic plasticity),
whereas no differences between $lnRR_{Q_{10_{acute}}}$ and
$lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of
physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In
contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to
$lnCVR_{Q_{10_{acute}}}$ would imply that changes in between individual
variation in physiological rates across 10°C differ depending on whether
acute or acclimation responses are measured.

Second, we tested whether acute and acclimation $lnRR_{Q_{10}}$ and
$lnCVR_{Q_{10}}$ differed between whole-organism versus biochemical
traits across habitats by fitting an model with an interaction between
type, habitat and trait category (referred to as "Model 3"). We also
explored how acute and acclimation effects varied across more detailed
trait categories, and these results can be found in the *Supplementary
Materials*. Given that residual variance appeared to differ across trait
categories we fit heteroscedastic residual variance models which were
better supported in all cases ($\Delta AIC_{c}$; marine = 58, freshwater
= 120, and terrestrial = 12). We expected that whole-organism traits
would be more likely to maintain relative variation in physiological
function and be less likely to acclimate because whole-organism function
relies on a greater number of biochemical reactions each with different
thermal sensitivies [@fields2001protein; @angilletta2009thermal].

Third, we tested whether different life-stages were more or less likely
to acclimate by fitting a model for each habitat type and including an
interaction between life-stage ('adult' or 'juvenile') and effect type
(referred to as "Model 4"). We predicted that acclimation responses
would be more likely early in development compared to later in
development as this pattern has been shown in previous studies [e.g.,
@moghadam2019heat], but that this should depend on the habitat type
given the different constraints faced by different early life stages
across major habitat types.

Finally, we tested whether the change in $lnRR_{Q_{10_{acclimation}}}$
and $lnCVR_{Q_{10}}$ were predicted by a populations climate variability
(CV). We only used $lnRR_{Q_{10_{acclimation}}}$ for these models
because our predictions were specifically focused on the plastic
adjustments to physiological rates. We used all $lnCVR_{Q_{10}}$ data
because there were no differences between $lnCVR_{Q_{10_{acclimation}}}$
and $lnCVR_{Q_{10_{acute}}}$ and we had no *a priori* reason to believe
climate would impact acclimation and acute $lnCVR_{Q_{10}}$ differently.
To test how climate variability impacted mean effect size in wild
populations we fit models that included an interaction between habitat
type and thermal coefficient of variability (CV) as moderators while
controlling for acclimation time and maximum temperature used in the
experiments (referred to as "Model 5"). We also explored how
environmental predictability, as measured by how correlated successive
temperatures at a given location were across months, might also explain
capacity for acclimation. However, such analyses are challenging to
interpret because the temporal scale that is biologically relevant to
different organisms will be different making the choice of lag to
estimate the correlation difficult to apply across taxa. As such, we
report on these results for a single lag (1 month) in the *Supplementary
Materials* but note they do not differ from CV analyses.

#### *Modelling how climate change will impact relative variance in physiological rates*

To understand the consequences of human-induced climate change on the
potential to impact relative variance in physiological rates we fit a
model that included an interaction between acclimation type, habitat
type, latitude and longitude (referred to as "Model 6"). We assumed that
any change in $lnCVR_{Q_{10}}$ across latitude and longitude could vary
by habitat type (i.e., an interaction between habitat). We used
non-linear tensors for latitude and longitude as any response could be
complicated by local factors (e.g., altitude). Our model included random
effects of species, trait, phylogeny and study. We predicted the
expected change in $lnCVR_{Q_{10}}$ for each wild population in our
dataset at the specific populations latitude and longitude. To do this,
we first converted the predicted $lnCVR_{Q_{10}}$ to a 1°C change as
opposed to 10°C as follows:

$$
lnCVR_{Q_{1}} = \frac{lnCVR_{Q_{10}}}{10}
$$ {#eq-lCVRpred}

@eq-lCVRpred turned the expected change across 10°C to 1°C. We then
multiplied this predicted change by the change in air and sea surface
temperatures at the locations of each population (and species) that is
expected under high emissions scenarios in 2080.

### *Publication Bias*

We explored the possibility for publication bias graphically using
funnel plots, and more formally by including the square root of the
inverse effective sample size ($\sqrt{1/ne}$) in our meta-regression
models [@Nakagawa2022-gy]. Funnel plot asymmetry may suggest a form of
publication bias called the 'file-drawer' effect whereby low-powered
studies are less likely to be published, however, graphical approaches
do not account for sources of non-independence and high heterogeneity
which can drive apparent funnel asymmetry [@Nakagawa2022-gy]. As such,
we included $\sqrt{1/ne}$ as a moderator in a multi-level
meta-regression model that accounted for all the random (i.e, study,
species, trait) and fixed effects (acclimation time, type of effect,
habitat, trait category and the interaction between habitat type and
trait category) (referred to as "Model 7").

```{r lnrrq10}
#| label: lnrrq10
#| echo: false
#| cache: true
#| message: false
#| warning: false
#| include: false

#############
## lnRRQ10 ##
#############
##################################################
# Organise the data and prune tree
##################################################
# Subset the data
lnRRQ10_data <- data_long2 %>% filter(grepl("lnRR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period)) %>% 
                              mutate(spp = species_full, c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE),
                                    max_temp_c = scale(temp_2, scale = FALSE))

# Check and prune tree
tree_lnrr <- tree_checks(data = lnRRQ10_data, tree = tree_p0.7, dataCol = "species_full", type = "prune")
tree_checks(data = lnRRQ10_data, tree = tree_lnrr, dataCol = "species_full", type = "check") # Good

# Create the phylo correlation matrix
phylo_cor <- vcv(tree_lnrr, cor = TRUE)
        A <- vcv(tree_lnrr, cor = FALSE) # For bayes, need cov matrix

# have a quick look on habitat, development and type. Mostly looks sensible but there is not too much data in terrestrial realm.
sum_life <- lnRRQ10_data %>% group_by(life.hist, type, habitat) %>% summarise(n=n()) %>% arrange(habitat)

##################################################
# Grafen mode comparison test with different power
##################################################

rerun = FALSE
if(rerun){
    # Create a list of the distinct trees to loop over
           list  <- list(tree_p0.5, tree_p0.6, tree_p0.7, tree_p0.8, tree_p0.9, tree_p1.0)
    names(list)  <- c("tree_p0.5", "tree_p0.6", "tree_p0.7", "tree_p0.8", "tree_p0.9", "tree_p1.0")

    # Create an empty list to store the models
        mod_list <- list()

    for(i in 1:length(list)){

    tree_lnrr_phylo <- tree_checks(data = lnRRQ10_data, tree = list[[i]], dataCol = "species_full", type = "prune")
    phylo_cor_phylo <- vcv(tree_lnrr_phylo, cor = TRUE)
       model1_phylo <- metafor::rma.mv(effect_size ~ 1, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 | spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_phylo), dfs = "contain", 
                              test = "t", data = lnRRQ10_data) 
             mod_list[[i]] <- model1_phylo 
        names(mod_list)[i] <- names(list)[i]       
        print(paste0("Model using ", names(mod_list)[i], " is finished with ", length(list)-i, " left to run"))
    }

  # Save the model list so that it doesn't need to be re-run again. It takes a LONG time otherwise.
    saveRDS(mod_list, file = "./output/models/lnRRQ10_phylo")

} else {

  # Load the model list
    #mod_list <- readRDS("./output/models/lnRRQ10_phylo")

  # Have a look at the AIC of the various models to compare which power paramater is best, if any.
   #aic <- sapply(mod_list, function(x) AIC(x))

}

##################################################
# Model 1.1: Intercept only model. 
##################################################
rerun = FALSE
if(rerun){
  model1.1 <- metafor::rma.mv(effect_size ~ 1, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 | spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data)

  model1.1_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  1 + (~1 | record_num) + (1 | gr(species_full, cov = A)) + (1 | spp) + (1|trait), data2 = list(A=A), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data)

  # Save model1.1
  saveRDS(model1.1, file = "./output/models/lnRRQ10_1.1")
  saveRDS(model1.1_bayes, file = "./output/models/lnRRQ10_1.1_bayes")
} else{ 
  model1.1 <- readRDS("./output/models/lnRRQ10_1.1")
model1.1_bayes <- readRDS("./output/models/lnRRQ10_1.1_bayes")
}

# Heterogeneity analysis
   w <- with(lnRRQ10_data, 1/V) # weights
  sv <- sum(w*(dim(lnRRQ10_data)[1]-1)) / (sum(w)^2 + sum(w^2))

# extract posterior sds, but square them to turn to variance
  posteriors_1.1 <- posterior_samples(model1.1_bayes, pars = c("^sd", "sigma"))^2
  colnames(posteriors_1.1) <- gsub("__Intercept", "", colnames(posteriors_1.1))
  colnames(posteriors_1.1) <- gsub("sd_", "", colnames(posteriors_1.1))

    res <- posteriors_1.1[,"sigma"]
   stdy <- posteriors_1.1[,"record_num"]
  phylo <- posteriors_1.1[,"species_full"]
    spp <- posteriors_1.1[,"spp"]
  trait <- posteriors_1.1[,"trait"]
  total <- phylo + spp + res + sv + stdy + trait

  # Calculate I2
     tot_i2 <- (sv / total)*100    ; mean(tot_i2)  ; quantile(tot_i2, c(0.025, 0.975))
   phylo_i2 <- (phylo / total)*100 ; mean(phylo_i2); quantile(phylo_i2, c(0.025, 0.975))
     spp_i2 <- (spp / total)*100   ; mean(spp_i2)  ; quantile(spp_i2, c(0.025, 0.975))
    stdy_i2 <- (stdy / total)*100  ; mean(stdy_i2) ; quantile(stdy_i2, c(0.025, 0.975))
   trait_i2 <- (trait / total)*100 ; mean(trait_i2); quantile(trait_i2, c(0.025, 0.975))
     res_i2 <- (res / total)*100   ; mean(res_i2)  ; quantile(res_i2, c(0.025, 0.975))

  # Table
     i2_lnrr <- data.frame(name = c("Phylogeny", "Species", "Study", "Trait"),
                           est = c(mean(phylo_i2),mean(spp_i2), mean(stdy_i2), mean(trait_i2)),
                           lci = c(quantile(phylo_i2, c(0.025, 0.975))[1], quantile(spp_i2, c(0.025, 0.975))[1],
                                   quantile(stdy_i2, c(0.025, 0.975))[1], quantile(trait_i2, c(0.025, 0.975))[1]),
                           uci = c(quantile(phylo_i2, c(0.025, 0.975))[2], quantile(spp_i2, c(0.025, 0.975))[2],
                                   quantile(stdy_i2, c(0.025, 0.975))[2], quantile(trait_i2, c(0.025, 0.975))[2]))

###################################################
# Model 1.2: Model controlling for acclimation time, 
# type and habitat assuming type can vary by habitat
###################################################

if(rerun){
  # More realistic model with 
  model1.2 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + habitat + type:habitat, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data)

  # Additional model fit as a result of reviewer commenting that acclimation time is assumed to be linear. So, fit a non-linear parameter to see if it improves fit (even though data doesn't suggest it is the case).
   model1.2_nl <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + I(c_accl_time^2) + type + habitat + type:habitat, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data)

    AIC(model1.2_nl, model1.2) # Again, basically the same and go with most parsimonious model

  # Secondary sensitivity analysis that uses cluster robust variance estimation to deal with unaccounted sources of non-independence. Not completely dealing with sharced sampling covariances in data as when collected these data were never indicated in past studies so hard to build full VCV matrix.

   model1.2_rve <- robust(model1.2, cluster = lnRRQ10_data$record_num) # Consistent.

  # Baysian model allows us to propagate error better is should be more robust to non-independence
   model1.2_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + type + habitat + type:habitat + (~1 | record_num) + (1 | gr(species_full, cov = A)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A=A), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data)

  # Save model1.2
  saveRDS(model1.2, file = "./output/models/lnRRQ10_1.2")
  saveRDS(model1.2_bayes, file = "./output/models/lnRRQ10_1.2_bayes")

  } else{
        model1.2 <- readRDS("./output/models/lnRRQ10_1.2")
  model1.2_bayes <- readRDS("./output/models/lnRRQ10_1.2_bayes")
}

####################################################
## Model 1.2: Bayesian Contrasts
###################################################
  posteriors_1.2 <- posterior_samples(model1.2_bayes, pars = "^b")
  f_acute  <- posteriors_1.2[,"b_Intercept"]; pmcmc(f_acute)
  f_acclim <- posteriors_1.2[,"b_Intercept"] + posteriors_1.2[,"b_typeacclim"]
  m_acute  <- posteriors_1.2[,"b_Intercept"] + posteriors_1.2[,"b_habitatm"]
  m_acclim <- posteriors_1.2[,"b_Intercept"] + posteriors_1.2[,"b_habitatm"] + posteriors_1.2[,"b_typeacclim"] + posteriors_1.2[,"b_typeacclim:habitatm"]
  t_acute  <- posteriors_1.2[,"b_Intercept"] + posteriors_1.2[,"b_habitatt"]
  t_acclim <- posteriors_1.2[,"b_Intercept"] + posteriors_1.2[,"b_habitatt"] + posteriors_1.2[,"b_typeacclim"] + posteriors_1.2[,"b_typeacclim:habitatt"]

# Q10 values for each category. Effect sizes are quite small really
   f_acute_Q10 <- exp(f_acute) ; mean(f_acute_Q10); quantile(f_acute_Q10, c(0.025, 0.975))
  f_acclim_Q10 <- exp(f_acclim); mean(f_acclim_Q10); quantile(f_acclim_Q10, c(0.025, 0.975))
   m_acute_Q10 <- exp(m_acute) ; mean(m_acute_Q10); quantile(m_acute_Q10, c(0.025, 0.975))
  m_acclim_Q10 <- exp(m_acclim); mean(m_acclim_Q10); quantile(m_acclim_Q10, c(0.025, 0.975))
   t_acute_Q10 <- exp(t_acute) ; mean(t_acute_Q10); quantile(t_acute_Q10, c(0.025, 0.975))
  t_acclim_Q10 <- exp(t_acclim); mean(t_acclim_Q10); quantile(t_acclim_Q10, c(0.025, 0.975))

# % change in mean acclimation relative to acute. Note that this is the change in two Q10 contrasts not the % difference between low and high temp (i.e., lnRR)
  # Calculate the mean lnRR for acclimation and acute, then calculate thir difference and make the relative change relative to acute to talk about % reduction in lnRRQ10 in relation to acute.
  change_f <- ((f_acclim - f_acute) / f_acute)*100; quantile(change_f, c(0.025, 0.975)) # 17% reductino in lnRRQ10
  change_m <- ((m_acclim - m_acute) / m_acute)*100; quantile(change_m, c(0.025, 0.975)) # 16% reuction in lnRRQ10
  change_t <- ((t_acclim - t_acute) / t_acute)*100; quantile(change_t, c(0.025, 0.975))
  avg_change <- (change_f + change_m + change_t) / 3; quantile(avg_change, c(0.025, 0.975)) # Could use weighted mean, but for now this seems fine.

####################################################
# Model 1.3's: Trait type varied by acclimation type
# but data is separated by habitat to simplify
####################################################
  # Add two general trait categories to test for trait differences (whole-organism vs biochemical)
    lnRRQ10_data <- lnRRQ10_data %>% 
                    mutate(trait_broad = ifelse(trait_category %in% c("cardiac", "rest_MR", "sprint", "muscle", "endurance", "max_MR"), 
                                                "tissue/organism",
                                         ifelse(trait_category %in% c("ATPase", "metabolic_enzyme", "antiox", "mito_oxidation", "mito_leak"), 
                                                "biochemical", trait_category)))

  # Filter to different habitats
   lnRRQ10_data_m <- lnRRQ10_data %>% filter(habitat == "m" & trait_broad != "other") 
   lnRRQ10_data_f <- lnRRQ10_data %>% filter(habitat == "f" & trait_broad != "other") 
   lnRRQ10_data_t <- lnRRQ10_data %>% filter(habitat == "t" & trait_broad != "other")

  # Prune tree
   tree_m <- tree_checks(lnRRQ10_data_m, tree = tree_lnrr, dataCol = "species_full", type = "prune")
   tree_f <- tree_checks(lnRRQ10_data_f, tree = tree_lnrr, dataCol = "species_full", type = "prune")
   tree_t <- tree_checks(lnRRQ10_data_t, tree = tree_lnrr, dataCol = "species_full", type = "prune")

   # cor matrices
   m_phylo <- vcv(tree_m, corr = TRUE)
       A_m <- vcv(tree_m, corr = FALSE)
   f_phylo <- vcv(tree_f, corr = TRUE)
       A_f <- vcv(tree_f, corr = FALSE)
   t_phylo <- vcv(tree_t, corr = TRUE)
       A_t <- vcv(tree_t, corr = FALSE)

if(rerun){

#################################
  # Broad Trait Category Models
  # Subset to only trait categories 
  # that are clear. Remove 'other'
#################################
  # Full Bayesian model   
   model1.3_btrait_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type*trait_broad*habitat + 
                                           (1 | record_num) + 
                                           (1 | gr(species_full, cov = A)) + 
                                           (1 + max_temp_c | spp) +(1|trait) + 
                                           (1 + trait_broad |gr(obs, cor = FALSE)),  
                                         data2 = list(A=A), family = gaussian(), 
                                         prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                                         control = list(adapt_delta = 0.98, max_treedepth = 12), 
                                         chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = filter(lnRRQ10_data, trait_broad != "other"))
    saveRDS(model1.3_btrait_bayes, "./output/models/model1.3_btrait_bayes")

  # Metafor models for creating plots. Subset by habitat to make this easier
      # Marine
         model1.3_m_broad <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_broad + type:trait_broad, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_broad| obs),
                              R = list(species_full = m_phylo), dfs = "contain", 
                              test = "t", struct = c("GEN","HCS"),  phi = 0,
                              data = lnRRQ10_data_m)
         saveRDS(model1.3_m_broad, "./output/models/model1.3_m_broad")

      # Freshwater
         model1.3_f_broad <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_broad + type:trait_broad, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_broad| obs),
                              R = list(species_full = f_phylo), dfs = "contain", 
                              test = "t", struct = c("GEN","HCS"), phi = 0,
                              data = lnRRQ10_data_f)
          saveRDS(model1.3_f_broad, "output/models/model1.3_f_broad")

      # Terrestrial
         model1.3_t_broad <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_broad + type:trait_broad, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_broad| obs),
                              R = list(species_full = t_phylo), dfs = "contain", 
                              test = "t", struct = c("GEN","HCS"), phi = 0,
                              data = lnRRQ10_data_t)
         saveRDS(model1.3_t_broad, "output/models/model1.3_t_broad")

  } else{
    model1.3_btrait_bayes <- readRDS("./output/models/model1.3_btrait_bayes")
         model1.3_m_broad <- readRDS("./output/models/model1.3_m_broad")
         model1.3_f_broad <- readRDS("./output/models/model1.3_f_broad")
         model1.3_t_broad <- readRDS("./output/models/model1.3_t_broad")
}

###########################################################
  # Processing Bayesian Posteriors
###########################################################
      # Create new data for predictions
       trait_b_data <- data.frame(  type = c("acclim", "acute"),
                              max_temp_c = 0,
                             c_accl_time = 0,
                                       V = 0,
                                 habitat = rep(c("t", "m", "f"), each = 4),
                             trait_broad = rep(c("tissue/organism", "biochemical"), each = 2))

       # Predict from model. Note this is the summary.
         pred_lnrr_traitb <- fitted(model1.3_btrait_bayes, 
                                     newdata = trait_b_data, re_formula = NA,
                                     summary = TRUE, 
                                     robust = FALSE)
         pred_lnrr_traitb <- cbind(trait_b_data, pred_lnrr_traitb)

       # Predict to get the posterior. Note that we have the full posterior distribution here for each of the 12 categories in the new data. Here, we want to look at the difference between acute and acclimation within each of the categories. 
            pred_lnrr_trait_post <- fitted(model1.3_btrait_bayes, 
                                     newdata = trait_b_data, re_formula = NA,
                                     summary = FALSE, 
                                     robust = FALSE)
        contrast_lnrr_trait_post <-  pred_lnrr_trait_post[,c(1,3,5,7,9,11)] - pred_lnrr_trait_post[,c(2,4,6,8,10,12)] # Acclimation - acute.

        # Contrast statistics
            contrast_lnrr_trait_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("tissue/organism", "biochemical"), each = 1),
                                                    diff = sapply(data.frame(contrast_lnrr_trait_post), mean),
                                                     l95 = sapply(data.frame(contrast_lnrr_trait_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lnrr_trait_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lnrr_trait_post), function(x) p_value(pmcmc(x))))

###########################################################
## Model 1.4 - whether chance between q10 depends on life stage and habitat
###########################################################
rerun = FALSE
if(rerun){
  # More realistic model with 
  model1.4 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type*habitat*life.hist, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data)

   model1.4_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + type*habitat*life.hist + (~1 | record_num) + (1 | gr(species_full, cov = A)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A=A), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data)

  # Save model1.4
  saveRDS(model1.4, file = "./output/models/lnRRQ10_1.4")
  saveRDS(model1.4_bayes, file = "./output/models/lnRRQ10_1.4_bayes")

  } else{
        model1.4 <- readRDS("./output/models/lnRRQ10_1.4")
  model1.4_bayes <- readRDS("./output/models/lnRRQ10_1.4_bayes")
}

#######################################################################
# Bayesian posteriors from model 1.4
#######################################################################
 # Create the data we want to make predictions for. Just more intuitive
  newdata <- data.frame( c_accl_time = 0, 
                          max_temp_c = 0, 
                                type = relevel(factor(c("acclim", "acute")), ref = "acute"), 
                             habitat = rep(c("m", "f", "t"), each = 4), 
                           life.hist = rep(c("j", "a"), each = 2))

# Calculate the design matrix using this new data. Need to be careful here to check that the levels are set up the same as the posterior.
  X <- model.matrix(~max_temp_c + c_accl_time + type*habitat*life.hist, data = newdata)

# Extract posteriors
  posteriors_1_4 <- posterior_samples(model1.4_bayes, pars = "^b")

# Now we can do all the calculations in one shot.
         pred <- apply(t(posteriors_1_4), 2, function(x) as.matrix(X)%*%x)
     sum_pred <- apply(pred, 1, function(x) mean(x))
  sum_predL95 <- apply(pred, 1, function(x) quantile(x, 0.025))
  sum_predU95 <- apply(pred, 1, function(x) quantile(x, 0.975))
            p <- apply(pred, 1, function(x) p_value(pmcmc(x)))

  # Create a summary dataset. 
  newdata <- cbind(newdata, est = sum_pred, L = sum_predL95, u = sum_predU95, p = p)

# Now, lets calculate the contrasts between acute and acclim for a and j within each habitat. We need to use pred for that because it's the entire posterior distribution
  ############
    # Contrast between Acclimation and Acute WITHIN each life-history stage. Double checked a bunch of calculations to make sure the right contrasts were being done. 
  ############
  contrast_lnrr_lh_post <-  pred[c(1,3,5,7,9,11),] - pred[c(2,4,6,8,10,12),] # Acclimation - acute.

        # Contrast statistics
            contrast_stage_lnrr_lh_stats <- data.frame( habitat = rep(c("m", "f", "t"), each = 2),
                                                     stage = rep(c("j", "a"), each = 1),
                                                    diff = apply(data.frame(contrast_lnrr_lh_post), 1, mean),
                                                     l95 = apply(data.frame(contrast_lnrr_lh_post), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lnrr_lh_post), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lnrr_lh_post), 1, function(x) p_value(pmcmc(x))))
    ############
    # Contrast between Acclimation and Acute ACROSS each life-history stage. Double checked a bunch of comparisons.
    ############
          contrast_lnrr_lh_post_type <-  pred[c(4,3,8,7,12,11),] - pred[c(2,1,6,5,10,9),] # Adult - Jeu.

        # Contrast statistics
            contrast_type_lnrr_lh_stats <- data.frame( habitat = rep(c("m", "f", "t"), each = 2),
                                                     type = rep(c("acute", "acclim"), each = 1),
                                                    diff = apply(data.frame(contrast_lnrr_lh_post_type), 1, mean),
                                                     l95 = apply(data.frame(contrast_lnrr_lh_post_type), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lnrr_lh_post_type), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lnrr_lh_post_type), 1, function(x) p_value(pmcmc(x))))

  # Acute differences between a and j across all habitats
     acute_all_lh <- c(contrast_lnrr_lh_post_type[1,], contrast_lnrr_lh_post_type[3,], contrast_lnrr_lh_post_type[5,]) 
    acclim_all_lh <- c(contrast_lnrr_lh_post_type[2,], contrast_lnrr_lh_post_type[4,], contrast_lnrr_lh_post_type[6,])

####################################################
# Sensitivity analysis -- Do Acute Q10's (hot / cold) differ?
####################################################
 # Goal here is to test the assumption that acute Q10's are different or not. Note Reviewer wanted more analysis in this space so added additional models
 lnRRQ10_data_acute <-  lnRRQ10_data %>% filter(grepl("acute", name))

 if(rerun){ 
    model1.s1 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + name, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_acute)
    saveRDS(model1.s1, file = "./output/models/model1.s1")

    model1.s1_int <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + name*habitat, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_acute)
    saveRDS(model1.s1_int, file = "./output/models/model1.s1_int")

    AIC(model1.s1_int, model1.s1) # AIC basically the same

    # Bayesian models to check its the same result
    model1.s1_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + name + (~1 | record_num) + (1 | gr(species_full, cov = A)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A=A), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_acute)

    saveRDS(model1.s1_bayes, file = "./output/models/model1.s1_bayes")

    model1.s1_bayes_int <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + name*habitat + (~1 | record_num) + (1 | gr(species_full, cov = A)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A=A), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_acute)

    saveRDS(model1.s1_bayes_int, file = "./output/models/model1.s1_bayes_int")

} else{
            model1.s1 <-  readRDS(file = "./output/models/model1.s1")
            model1.s1_int <-  readRDS(file = "./output/models/model1.s1_int")
      model1.s1_bayes <- readRDS(file= "./output/models/model1.s1_bayes")
}

post_acute <- posterior_samples(model1.s1_bayes, pars = "^b")
  hot_acute <- post_acute[,"b_Intercept"] + post_acute[,"b_namelnRR_Q10_acute_t2"]
  cold_acute <- post_acute[,"b_Intercept"]
  change_acute <- post_acute[,"b_namelnRR_Q10_acute_t2"]

n_acutes <- lnRRQ10_data_acute  %>% group_by(name, habitat) %>% summarise(n = n(), stdy = n_distinct(record_num), spp = n_distinct(species_full))  %>% arrange(habitat)

```

## **Results**

### *Data Summary*

The final dataset included a total of
`r f <- taxa_sum |> dplyr::filter(habitat == "f") |> dplyr::summarise(spp = sum(spp), k = sum(n)); f[1,1]`
freshwater (fishes =
`r taxa_sum |> dplyr::filter(habitat == "f" & class == "Actinopterygii") |> dplyr::summarise(sp = sum(spp))`,
molluscs =
`r taxa_sum |> dplyr::filter(habitat == "f" & phylum == "Mollusca") |> dplyr::summarise(sp = sum(spp))`,
amphibians =
`r taxa_sum |> dplyr::filter(habitat == "f" & class == "Amphibia") |> dplyr::summarise(sp = sum(spp))`,
reptiles =
`r taxa_sum |> dplyr::filter(habitat == "f" & class == "Reptilia") |> dplyr::summarise(sp = sum(spp))`,
arthropods =
`r taxa_sum |> dplyr::filter(habitat == "f" & phylum == "Arthropoda") |> dplyr::summarise(sp = sum(spp))`,
and a single crustacean and nematode species),
`r m <- taxa_sum |> dplyr::filter(habitat == "m") |> dplyr::summarise(spp = sum(spp), k = sum(n)); m[1,1]`
marine (fishes =
`r taxa_sum |> dplyr::filter(habitat == "m" & class %in% c("Actinopterygii", "Agnatha","Chondrichthyes")) |> dplyr::summarise(sp = sum(spp))`,
annelids =
`r taxa_sum |> dplyr::filter(habitat == "m" & phylum == "Annelida") |> dplyr::summarise(sp = sum(spp))`,
molluscs =
`r taxa_sum |> dplyr::filter(habitat == "m" & phylum == "Mollusca") |> dplyr::summarise(sp = sum(spp))`,
echinoderms =
`r taxa_sum |> dplyr::filter(habitat == "m" & phylum == "Echinodermata") |> dplyr::summarise(sp = sum(spp))`,
reptiles =
`r taxa_sum |> dplyr::filter(habitat == "m" & class == "Reptilia") |> dplyr::summarise(sp = sum(spp))`,
arthropods =
`r taxa_sum |> dplyr::filter(habitat == "m" & phylum == "Arthropoda") |> dplyr::summarise(sp = sum(spp))`,
and a single crustacean and cnidarian species), and
`r t <- taxa_sum |> dplyr::filter(habitat == "t") |> dplyr::summarise(spp = sum(spp), k = sum(n)); t[1,1]`
terrestrial species (annelids =
`r taxa_sum |> dplyr::filter(habitat == "t" & phylum == "Annelida") |> dplyr::summarise(sp = sum(spp))`,
molluscs =
`r taxa_sum |> dplyr::filter(habitat == "t" & phylum == "Mollusca") |> dplyr::summarise(sp = sum(spp))`,
arthropods =
`r taxa_sum |> dplyr::filter(habitat == "t" & phylum == "Arthropoda") |> dplyr::summarise(sp = sum(spp))`,
reptiles =
`r taxa_sum |> dplyr::filter(habitat == "t" & class == "Reptilia") |> dplyr::summarise(sp = sum(spp))`
and amphibians =
`r taxa_sum |> dplyr::filter(habitat == "t" & class == "Amphibia") |> dplyr::summarise(sp = sum(spp))`
along with a single tardigrade species) (@fig-phylo). We had more data
on acute thermal responses (n =
`r trimmed <- data_long2 |> dplyr::filter(grepl("lnRR", name) & !is.na(effect_size) & !is.na(V) & !V == 0); sum1 = trimmed |> dplyr::group_by(name, type) |> dplyr::summarise(n = dplyr::n()) |> dplyr::filter(type =="acute"); sum(sum1$n)`)
compared to thermal responses after an acclimation period (n =
`r sum2 <- trimmed |> dplyr::group_by(name, type) |> dplyr::summarise(n = dplyr::n()) |> dplyr::filter(type =="acclim"); sum2$n`)
because acute responses were reported for each of the two acclimation
temperatures (@fig-phylo). The two acute $lnRR_{Q_{10}}$ effect sizes
([@fig-concept]A) differed significantly from each other ($\beta$ =
`r mean(change_acute)`, 95% CI: `r quantile(change_acute, 0.025)` to
`r quantile(change_acute, 0.975)`, $p_{MCMC}$ =
`r p_value(pmcmc(change_acute))`) with animals acclimated to high
temperatures having slightly higher average $lnRR_{Q_{10}}$ ($\mu$ =
`r mean(hot_acute)`, 95% CI: 95% CI: `r quantile(hot_acute, 0.025)` to
`r quantile(hot_acute, 0.975)`, $p_{MCMC}$ =
`r p_value(pmcmc(hot_acute))`, $Q_{10}$ = `r mean(exp(hot_acute))`)
compared to animals at lower temperatures ($\mu$ = `r mean(cold_acute)`,
95% CI: 95% CI: `r quantile(cold_acute, 0.025)` to
`r quantile(cold_acute, 0.975)`, $p_{MCMC}$ =
`r p_value(pmcmc(cold_acute))`, $Q_{10}$ = `r mean(exp(cold_acute))`)
(@fig-Sacute). However, on average they were in the same direction and
only differed by \~10%. As such, we averaged the two acute
$lnRR_{Q_{10}}$ effect sizes in subsequent analyses.

Most of the effect size estimates came from measurements of metabolic
rates (both resting and maximal -- $N_{species}$ =
`r mr <- traits_all |> dplyr::filter(trait_category %in% c("rest_MR", "max_MR")); sum(mr[,"spp_n"])`,
$N_{effects}$ = `r sum(mr[,"k"])`, considering acute and acclimation
effects together), metabolic enzyme rates ($N_{species}$ =
`r me <- traits_all |> dplyr::filter(trait_category %in% c("metabolic_enzyme", "ATPase")); sum(me[,"spp_n"])`,
$N_{effects}$ =
`r me <- traits_all |> dplyr::filter(trait_category %in% c("metabolic_enzyme", "ATPase")); sum(me[,"k"])`)
and whole-organism performance traits (i.e., measures of speed and
endurance -- $N_{species}$ =
`r per <- traits_all |> dplyr::filter(trait_category %in% c("sprint", "endurance")); sum(per[,"spp_n"])`,
$N_{effects}$ =
`r per <- traits_all |> dplyr::filter(trait_category %in% c("sprint", "endurance")); sum(per[,"k"])`).

```{r, fig-phylo}
#| label: fig-phylo
#| echo: false
#| fig-cap: "Phylogenetic distribution of acute and acclimation $lnRR_{Q_{10}}$/$lnCVR_{Q_{10}}$ estimates across major habitats.  The total number of acute and acclimation  $lnRR_{Q_{10}}$/$lnCVR_{Q_{10}}$ effect sizes are indicated by the coloured bars as well as whether the taxa are marine, freshwater or terrestrial. Silhouettes are only representative taxa of major clades within the tree."

##################
  # Figure 2 plot
  # Phylogeny
##################
   # Map data to tree. Just one variable for now
      d <- ggtree(tree, layout="circular") %<+% tree_data

      # Plot the tree

       col <- c("#0871B9", "#F3B40D", "#1BB908")
      p1 <- d + 
            #geom_tiplab(size=2, offset = 10) + 
            geom_tippoint(aes(color=habitat), size = 2.5) +
            labs(color = "Habitat") + 
            scale_colour_manual(values = col)

      # Create the data matrix that can be used to plot data as a heatmap around the phylogeny
      matrix_data <- as.data.frame(data_wide %>%
                  mutate(trait_category = ifelse(trait_category %in% 
                                                   c("ATPase", "mito_leak", "mito_oxidation"), "Mito. Function", trait_category)) %>%
                  mutate(trait_category = ifelse(trait_category %in% c("endurance", "sprint"), "Performance", trait_category))    %>%
                  mutate(trait_category = ifelse(trait_category %in% c("muscle", "cardiac"), "Muscle Function", trait_category))  %>%
                  mutate(trait_category = ifelse(trait_category %in% c("rest_MR", "max_MR"), "Metabolic Rate", trait_category))   %>%
                  group_by(species_full, trait_category)   %>% 
                  summarise(n = n())                       %>% 
                  spread(trait_category, n, fill = 0))

      row.names(matrix_data) <- matrix_data$species_full
      matrix_data <- matrix_data[,c(2:8)]

      colnames(matrix_data)[c(1,3,6)] <- c("Antioxidants", "Metabolic Enzymes", "Other")

      # Alternatively, summarise total effect sizes per species for acute and acclimation
      matrix_data2 <- data_long %>% filter(grepl("lnRR_Q10", name)) %>% 
                      group_by(species_full, type) %>% 
                      summarise(n = n()) %>% 
                      spread(type, n, fill = 0) %>% data.frame()
                     # mutate(acclim = as.character(cut(acclim, breaks = c(0,15, 20, 25, 30, 40, 50))),
                      #        acute = as.character(cut(acute, breaks = c(0,15, 20, 25, 30, 40, 50))))

      row.names(matrix_data2) <- matrix_data2$species_full
      matrix_data2 <- matrix_data2[,2:3]

      # Plot a heatmap of N for each trait around the phylogeny.
   phylo <-   gheatmap(p1, data.matrix(matrix_data2), offset = 0,  colnames = FALSE, 
               colnames_position="bottom", legend_title = "N", 
               colnames_angle=360, colnames_offset_y = 0, colnames_offset_x = -1.5, 
               hjust=0, font.size=3, width = 0.35) + 
                scale_fill_viridis(alpha = 0.6) + labs(fill="# of Effects") + theme(legend.text =element_text(size = 24),
                                                                                    legend.title = element_text(size = 30, face = "bold"))

  #ggsave(filename = "fig2_unedited_18June.pdf", path = "./output/figures/", phylo, device = "pdf", width = 10, height = 9, units = "in")

# Load and plot Figure 1. Note that this figure was derived using 3 separate code chunks described below. Figures were exported, imported into Adobe Illustrator for minor formatting changes and grid organisation. 
  fig2 <- image_read("./output/figures/Fig2_r1.png")
  plot(fig2)

```

### *Terrestrial and aquatic ectotherms differ in their capacity to acclimate but acclimation does not depend on life-history stage*

Effect heterogeneity was high (only `r mean(tot_i2)`% of the variance
was the result of sampling variability, 95% CI:
`r quantile(tot_i2, 0.025)[1]` to `r quantile(tot_i2, 0.975)[1]`%), and
most variance was explained by the specific study and type of trait
(Study: `r mean(stdy_i2)`% , 95% CI: `r quantile(stdy_i2, 0.025)[1]` to
`r quantile(stdy_i2, 0.975)[1]`%; Trait Type: `r mean(trait_i2)`% , 95%
CI: `r quantile(trait_i2, 0.025)[1]` to
`r quantile(trait_i2, 0.975)[1]`%). Evolutionary relationships among
taxa and species ecology (i.e., species random effect) explained little
variation in acute and acclimation responses (Species: `r mean(spp_i2)`%
, 95% CI: `r quantile(spp_i2, 0.025)[1]` to
`r quantile(spp_i2, 0.975)[1]`%; Phylogeny: `r mean(phylo_i2)`% , 95%
CI: `r quantile(phylo_i2, 0.025)[1]` to
`r quantile(phylo_i2, 0.975)[1]`%). All results are from Model 1 (see
“Meta-Analysis” above).

Capacity for acclimation depended on the habitat. Ectotherms in marine
and freshwater environments showed partial compensation of physiological
rates ([@fig-2]A) amounting to reduced $lnRR_{Q_{10_{acclimation}}}$ of
`r abs(mean(change_f))`% (95% CI:
`r quantile(change_f, c(0.025, 0.975))[1]` to
`r quantile(change_f, c(0.025, 0.975))[2]`) in freshwater and
`r abs(mean(change_m))`% (95% CI:
`r quantile(change_m, c(0.025, 0.975))[1]` to
`r quantile(change_m, c(0.025, 0.975))[2]`) in marine environments. In
contrast, terrestrial ectotherms showed no acclimation (possibly even
inverse acclimation) -- showing a `r abs(mean(change_t))`% increase in
$lnRR_{Q_{10_{acclimation}}}$ (95% CI:
`r quantile(change_t, c(0.025, 0.975))[1]` to
`r quantile(change_t, c(0.025, 0.975))[2]`, [@fig-2]B). All results are
from Model 2 (see “Meta-Analysis” above).

Across habitats the extent to which whole-organism versus biochemical
processes acclimated varied ("Model 3"; [@fig-btrait]A-C; and see
Supplement, @fig-S2) for more detailed trait analysis). Biochemical
processes were more effective at acclimating compared to whole-organism
traits for ectotherms occupying marine habitats ([@fig-btrait]B),
whereas both whole-organism and biochemical traits acclimated in
freshwater ectotherms ([@fig-btrait]C). Unsurprisingly, neither broad
trait category showed evidence of acclimation in terrestrial ectotherms
([@fig-btrait]A).

Acclimation capacity did not vary consistently by life-history stage
with no differences between $lnRR_{Q_{10_{acclimation}}}$ and
$lnRR_{Q_{10_{acute}}}$ between adult and juveniles, except for marine
habitats where adults appeared to be able to acclimate to a greater
extent, although the effect size was small and in the same direction and
magnitude of juveniles ("Model 4"; [@fig-cvrlh]A-C). When comparing
between acute and acclimation responses between adults and juveniles
there were also no differences across habitats (Adult-Juvenile (Acute):
`r mean(acute_all_lh)`, 95% CI: `r quantile(acute_all_lh, 0.025)[1]` to
`r quantile(acute_all_lh, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(acute_all_lh)`; Adult-Juvenile (Acclimation):
`r mean(acclim_all_lh)`, 95% CI: `r quantile(acclim_all_lh, 0.025)[1]`
to `r quantile(acclim_all_lh, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(acclim_all_lh)`).

```{r, lncvrq10}
#| label: lncvrq10
#| echo: false
#| warning: false
#| error: false
#| include: false
#| cache: false

#############
  ## lnCVRQ10
#############
##################################################
  # Organise the data and prune tree
#################################################
  lnCVRQ10_data <- data_long2 %>% filter(grepl("lnCVR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period)) %>% 
                              mutate(spp = species_full, c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE),
                                    max_temp_c = scale(temp_2, scale = FALSE))

# Check and prune tree
  tree_lncvr <- tree_checks(data = lnCVRQ10_data, tree = tree_p0.7, dataCol = "species_full", type = "prune")
  tree_checks(data = lnCVRQ10_data, tree = tree_lncvr, dataCol = "species_full", type = "check") # Good

# Create the phylo correlation matrix
  phylo_cor_CVR <- vcv(tree_lncvr, cor = TRUE)
          A_CVR <- vcv(tree_lncvr, cor = FALSE) # For bayes, need cov matrix

###################################################
  # Model 2.1: Intercept only model.
  # Used for heterogeneity analysis
###################################################
rerun = FALSE
if(rerun){
  model2.1 <- metafor::rma.mv(effect_size ~ 1, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_CVR), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data)

  model2.1_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  1 + (~1 | record_num) + (1 | gr(species_full, cov = A_CVR)) + (1 | spp) + (1|trait), data2 = list(A_CVR=A_CVR), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.99, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnCVRQ10_data)

  # Save model1.1
  saveRDS(model2.1, file = "./output/models/lnCVRQ10_2.1")
  saveRDS(model2.1_bayes, file = "./output/models/lnCVRQ10_2.1_bayes")
} else{ 
        model2.1 <- readRDS("./output/models/lnCVRQ10_2.1")
  model2.1_bayes <- readRDS("./output/models/lnCVRQ10_2.1_bayes")
}

# Heterogeneity analysis
   w_cvr <- with(lnCVRQ10_data, 1/V) # weights
  sv_cvr <- sum(w_cvr*(dim(lnCVRQ10_data)[1]-1)) / (sum(w_cvr)^2 + sum(w_cvr^2))

# extract posterior sds, but square them to turn to variance
  posteriors_2.1 <- posterior_samples(model2.1_bayes, pars = c("^sd", "sigma"))^2
  colnames(posteriors_2.1) <- gsub("__Intercept", "", colnames(posteriors_2.1))
  colnames(posteriors_2.1) <- gsub("sd_", "", colnames(posteriors_2.1))

    res_cvr <- posteriors_2.1[,"sigma"]
   stdy_cvr <- posteriors_2.1[,"record_num"]
  phylo_cvr <- posteriors_2.1[,"species_full"]
    spp_cvr <- posteriors_2.1[,"spp"]
  trait_cvr <- posteriors_2.1[,"trait"]
  total_cvr <- phylo_cvr + spp_cvr + res_cvr + sv_cvr + stdy_cvr + trait_cvr

  # Calculate I2
     tot_i2_cvr <- (sv_cvr / total_cvr)*100    ; mean(tot_i2_cvr)  ; quantile(tot_i2_cvr, c(0.025, 0.975))
   phylo_i2_cvr <- (phylo_cvr / total_cvr)*100 ; mean(phylo_i2_cvr); quantile(phylo_i2_cvr, c(0.025, 0.975))
     spp_i2_cvr <- (spp_cvr / total_cvr)*100   ; mean(spp_i2_cvr)  ; quantile(spp_i2_cvr, c(0.025, 0.975))
    stdy_i2_cvr <- (stdy_cvr / total_cvr)*100  ; mean(stdy_i2_cvr) ; quantile(stdy_i2_cvr, c(0.025, 0.975))
   trait_i2_cvr <- (trait_cvr / total_cvr)*100 ; mean(trait_i2_cvr); quantile(trait_i2_cvr, c(0.025, 0.975))
     res_i2_cvr <- (res_cvr / total_cvr)*100   ; mean(res_i2_cvr)  ; quantile(res_i2_cvr, c(0.025, 0.975))

  # Table
     i2_lncvr <- data.frame(name = c("Phylogeny", "Species", "Study", "Trait"),
                           est = c(mean(phylo_i2_cvr),mean(spp_i2_cvr), mean(stdy_i2_cvr), mean(trait_i2_cvr)),
                           lci = c(quantile(phylo_i2_cvr, c(0.025, 0.975))[1], quantile(spp_i2_cvr, c(0.025, 0.975))[1],
                                   quantile(stdy_i2_cvr, c(0.025, 0.975))[1], quantile(trait_i2_cvr, c(0.025, 0.975))[1]),
                           uci = c(quantile(phylo_i2_cvr, c(0.025, 0.975))[2], quantile(spp_i2_cvr, c(0.025, 0.975))[2],
                                   quantile(stdy_i2_cvr, c(0.025, 0.975))[2], quantile(trait_i2_cvr, c(0.025, 0.975))[2]))

###################################################
  # Model 2.2: Model controlling for acclimation time, 
  # type and habitat assuming type can vary by habitat
###################################################

if(rerun){
  # More realistic model 
  model2.2 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + habitat + type:habitat, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_CVR), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data)

   model2.2_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + type + habitat + type:habitat + (~1 | record_num) + (1 | gr(species_full, cov = A_CVR)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_CVR=A_CVR), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnCVRQ10_data)

  # Save model1.2
  saveRDS(model2.2, file = "./output/models/lnCVRQ10_2.2")
  saveRDS(model2.2_bayes, file = "./output/models/lnCVRQ10_2.2_bayes")

  } else{
        model2.2 <- readRDS("./output/models/lnCVRQ10_2.2")
  model2.2_bayes <- readRDS("./output/models/lnCVRQ10_2.2_bayes")
}

######################################################
  # Bayesian Contrasts
######################################################

  posteriors_2.2 <- posterior_samples(model2.2_bayes, pars = "^b")
  f_acute_cvr  <- posteriors_2.2[,"b_Intercept"]; pmcmc(f_acute_cvr)
  f_acclim_cvr <- posteriors_2.2[,"b_Intercept"] + posteriors_2.2[,"b_typeacclim"]
  m_acute_cvr  <- posteriors_2.2[,"b_Intercept"] + posteriors_2.2[,"b_habitatm"]
  m_acclim_cvr <- posteriors_2.2[,"b_Intercept"] + posteriors_2.2[,"b_habitatm"] + posteriors_2.2[,"b_typeacclim"] + posteriors_2.2[,"b_typeacclim:habitatm"]
  t_acute_cvr  <- posteriors_2.2[,"b_Intercept"] + posteriors_2.2[,"b_habitatt"]
  t_acclim_cvr <- posteriors_2.2[,"b_Intercept"] + posteriors_2.2[,"b_habitatt"] + posteriors_2.2[,"b_typeacclim"] + posteriors_2.2[,"b_typeacclim:habitatt"]

# We'll take the marginalised mean for each type as there is no difference between acclimation and acute really except for in freshwater, but this effect size is so tiny. It's a 10% reduction at high temps and acclim is only a 2% reduction in variance at high temps.
  # First get sample sizes
    N <- lnCVRQ10_data %>% group_by(habitat, type) %>% summarise(n=n())

  # Calculated weighted mean by sample size. 
    f_cvr_mean <- (f_acute_cvr * N[1,"n"][[1]] + f_acclim_cvr * N[2,"n"][[1]]) / sum(N[1:2, "n"]); pmcmc(f_cvr_mean)
    m_cvr_mean <- (m_acute_cvr * N[3,"n"][[1]] + m_acclim_cvr * N[3,"n"][[1]]) / sum(N[3:4, "n"]); pmcmc(m_cvr_mean)
    t_cvr_mean <- (t_acute_cvr * N[5,"n"][[1]] + t_acclim_cvr * N[6,"n"][[1]]) / sum(N[5:6, "n"]); pmcmc(t_cvr_mean)

  # Lets now convert these to a %
    f_cvr_mean_ef  <- abs((exp(f_cvr_mean) - 1)*100) # ~8.4% reduction
    m_cvr_mean_ef  <- abs((exp(m_cvr_mean) - 1)*100) # ~15% reduction
    t_cvr_mean_ef  <- abs((exp(t_cvr_mean) - 1)*100) # ~30% reduction

###################################################
  # Model 2.3: Model controlling for acclimation time, 
  # type and trait category assuming type can vary by habitat
###################################################
  # Add two general trait categories to test for trait differences (whole-organism vs biochemical)
    lnCVRQ10_data <- lnCVRQ10_data %>% 
                    mutate(trait_broad = ifelse(trait_category %in% c("cardiac", "rest_MR", "sprint", "muscle", "endurance", "max_MR"), 
                                                "tissue/organism",
                                         ifelse(trait_category %in% c("ATPase", "metabolic_enzyme", "antiox", "mito_oxidation", "mito_leak"), 
                                                "biochemical", trait_category)))
 # We can also run subsets for each habitat
   tree_m_cvr <- tree_checks(lnCVRQ10_data %>% filter(habitat == "m" & trait_broad != "other"), tree = tree_lncvr, dataCol = "species_full", type = "prune")
   tree_f_cvr <- tree_checks(lnCVRQ10_data %>% filter(habitat == "f" & trait_broad != "other"), tree = tree_lncvr, dataCol = "species_full", type = "prune")
   tree_t_cvr <- tree_checks(lnCVRQ10_data %>% filter(habitat == "t" & trait_broad != "other"), tree = tree_lncvr, dataCol = "species_full", type = "prune")

 # cor matrices
   m_phylo_cvr <- vcv(tree_m_cvr, corr = TRUE)
   f_phylo_cvr <- vcv(tree_f_cvr, corr = TRUE)
   t_phylo_cvr <- vcv(tree_t_cvr, corr = TRUE)

if(rerun){

  # Full Bayesian model   
   model2.3_btrait_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type*trait_broad*habitat + 
                                           (1 | record_num) + 
                                           (1 | gr(species_full, cov = A)) + 
                                           (1 + max_temp_c | spp) +(1|trait) + 
                                           (1 + trait_broad |gr(obs, cor = FALSE)),  
                                         data2 = list(A=A), family = gaussian(), 
                                         prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                                         control = list(adapt_delta = 0.98, max_treedepth = 12), 
                                         chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = filter(lnCVRQ10_data, trait_broad != "other"))
    saveRDS(model2.3_btrait_bayes, "./output/models/model2.3_btrait_bayes")

  # Metafor models for creating plots. Subset by habitat to make this easier
      # Marine
         model2.3_m_broad <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_broad + type:trait_broad, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_broad| obs),
                              R = list(species_full = m_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN","HCS"), phi = 0,
                              data = filter(lnCVRQ10_data, habitat == "m" & trait_broad != "other"))
         saveRDS(model2.3_m_broad, "output/models/model2.3_m_broad")

      # Freshwater
         model2.3_f_broad <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_broad + type:trait_broad, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_broad| obs),
                              R = list(species_full = f_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN","HCS"),  phi = 0,
                              data = filter(lnCVRQ10_data, habitat == "f" & trait_broad != "other"))
          saveRDS(model2.3_f_broad, "output/models/model2.3_f_broad")

      # Terrestrial
         model2.3_t_broad <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_broad + type:trait_broad, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_broad| obs),
                              R = list(species_full = t_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN","HCS"), phi = 0,
                              data = filter(lnCVRQ10_data, habitat == "t" & trait_broad != "other"))
         saveRDS(model2.3_t_broad, "output/models/model2.3_t_broad")

  } else{
    model2.3_btrait_bayes <- readRDS("./output/models/model2.3_btrait_bayes")
    model2.3_m_broad <- readRDS("output/models/model2.3_m_broad")
    model2.3_f_broad <- readRDS("output/models/model2.3_f_broad")
    model2.3_t_broad <- readRDS("output/models/model2.3_t_broad")
}

###########################################################
  # Processing Bayesian Posteriors
###########################################################
      # Create new data for predictions
       trait_b_data <- data.frame(  type = c("acclim", "acute"),
                              max_temp_c = 0,
                             c_accl_time = 0,
                                       V = 0,
                                 habitat = rep(c("t", "m", "f"), each = 4),
                             trait_broad = rep(c("tissue/organism", "biochemical"), each = 2))

       # Predict from model. Note this is the summary.
         pred_lncvr_traitb <- fitted(model2.3_btrait_bayes, 
                                     newdata = trait_b_data, re_formula = NA,
                                     summary = TRUE, 
                                     robust = FALSE)
         pred_lncvr_traitb <- cbind(trait_b_data, pred_lncvr_traitb)

       # Predict to get the posterior. Note that we have the full posterior distribution here for each of the 12 categories in the new data. Here, we want to look at the difference between acute and acclimation within each of the categories. 
            pred_lncvr_traitb_post <- fitted(model2.3_btrait_bayes, 
                                     newdata = trait_b_data, re_formula = NA,
                                     summary = FALSE, 
                                     robust = FALSE)
        contrast_lncvr_trait_post <-  pred_lncvr_traitb_post[,c(1,3,5,7,9,11)] - pred_lncvr_traitb_post[,c(2,4,6,8,10,12)] # Acclimation - acute.

        # Contrast statistics
            contrast_lncvr_trait_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("tissue/organism", "biochemical"), each = 1),
                                                    diff = sapply(data.frame(contrast_lncvr_trait_post), mean),
                                                     l95 = sapply(data.frame(contrast_lncvr_trait_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lncvr_trait_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lncvr_trait_post), function(x) p_value(pmcmc(x))))

   # Calculate the design matrix using this new data. Need to be careful here to check that the levels are set up the same as the posterior.

###################################################
  # Model 2.4: Model controlling for acclimation time, 
  # type and life history stage assuming type can vary by habitat
###################################################

if(rerun){
  # More realistic model 
  model2.4 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type*habitat*life.hist, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_CVR), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data)
  # Save model
  saveRDS(model2.4, file = "./output/models/lnCVRQ10_2.4")

   model2.4_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type*habitat*life.hist + 
                                 (~1 | record_num) + 
                                 (1 | gr(species_full, cov = A_CVR)) + 
                                 (1 + max_temp_c| spp) + 
                                 (1|trait) +
                                 (1|obs), 
                               data2 = list(A_CVR=A_CVR), family = gaussian(), 
                               prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                               control = list(adapt_delta = 0.98, max_treedepth = 12), 
                               chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnCVRQ10_data)

  # Save model
  saveRDS(model2.4_bayes, file = "./output/models/lnCVRQ10_2.4_bayes")

  } else{
        model2.4 <- readRDS("./output/models/lnCVRQ10_2.4")
  model2.4_bayes <- readRDS("./output/models/lnCVRQ10_2.4_bayes")
}

###########################################################
  # Processing Bayesian Posteriors
###########################################################
      # Create new data for predictions
       lh_b_data <- data.frame( c_accl_time = 0, 
                          max_temp_c = 0, 
                                type = relevel(factor(c("acclim", "acute")), ref = "acute"), 
                             habitat = rep(c("m", "f", "t"), each = 4), 
                           life.hist = rep(c("j", "a"), each = 2))

            # I have no idea why the fitted function is not working on this model. We're doing this the old fashioned way
            X <- model.matrix(~max_temp_c + c_accl_time + type*habitat*life.hist, data = lh_b_data)

            # Extract posteriors
              post_lh <- posterior_samples(model2.4_bayes, pars = "^b")

        # Now we can do all the calculations in one shot.# Predict to get the posterior. Note that we have the full posterior distribution here for each of the 12 categories in the new data. Here, we want to look at the difference between acute and acclimation within each of the categories. 
            pred_lh <- apply(t(post_lh),2 , function(x) as.matrix(X)%*%x)

        contrast_lncvr_lh_post <-  pred_lh[c(1,3,5,7,9,11),] - pred_lh[c(2,4,6,8,10,12),] # Acclimation - acute.

        # Contrast statistics
            contrast_lncvr_lh_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     stage = rep(c("j", "a"), each = 1),
                                                    diff = apply(data.frame(contrast_lncvr_lh_post), 1, mean),
                                                     l95 = apply(data.frame(contrast_lncvr_lh_post), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lncvr_lh_post), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lncvr_lh_post), 1, function(x) p_value(pmcmc(x))))
```

```{r, fig-2}
#| label: fig-2
#| fig-cap: "A Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. B Mean acute and acclimation $lnCVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A and B. In panels A and B, thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

##################
  # Figure 3a plot
##################

  col <- c("#0871B9", "#F3B40D", "#1BB908")
  k_lnrr <- manual_extract(model1.2, mod = "habitat", group = "record_num", exclude = NULL)

  p3a <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 20, legend.pos = "top.left", trunk.size = 1.8, branch.size = 2.8) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +

  annotate("text", x = 1.5, y = -0.30, label =TeX(paste0("$\\beta = $", round(mean(f_acclim - f_acute), 2), "[", round(quantile(f_acclim - f_acute, 0.025), 2)," to ", round(quantile(f_acclim - f_acute, 0.975), 2), "] ", " $\\,p_{MCMC}$ < 0.001")), hjust = 0, size = 8) + 
  annotate("text", x = 1.20, y = -0.5, label =paste0(round(mean(change_f), 3), "%"), hjust = 0, size = 8) + labs(shape = "Type") +

  annotate("text", x = 2.5, y = -0.30, label =TeX(paste0("$\\beta = $", round(mean(m_acclim - m_acute), 2), "[", round(quantile(m_acclim - m_acute, 0.025), 2)," to ", round(quantile(m_acclim - m_acute, 0.975),2), "] ", " $\\,p_{MCMC}$ = 0.01")), hjust = 0, size = 8) + 
  annotate("text", x = 2.20, y = -0.5, label =paste0(round(mean(change_m), 3), "%"), hjust = 0, size = 8) +

  annotate("text", x = 3.5, y = -0.30, label =TeX(paste0("$\\beta = $", round(mean(t_acclim - t_acute), 2), "[", round(quantile(t_acclim - t_acute, 0.025), 2)," to ", round(quantile(t_acclim - t_acute, 0.975),2), "] ", "$\\,p_{MCMC}$ = 0.36")), hjust = 0, size = 8) + 
  annotate("text", x = 3.20, y = -0.5, label =paste0(round(mean(change_t), 3), "%"), hjust = 0, size = 8) +

  scale_fill_manual(values = col) + scale_color_manual(values = rep("black",3)) + ylim(-0.5, 1.5)

  p3a <- p3a + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top")

##################
  # Figure 3b plot
################## 
  # Clear from the model that there really is no difference between the change in variance for acute or acclimation CVR. But, there is a clear reduction in between individual variance for terrestrial species. As temperature get hotter then the variance in physiological rates decreases. This is not true of aquatic systems.

  k_cvr <- manual_extract(model2.2, mod = "habitat", group = "record_num", exclude = NULL)

  x1 = 1.4
  x2 = 2.4
  x3 = 3.4

  p3b <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), by = "type", at = list(type = c("acute", "acclim")), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 1.8, branch.size = 2.8,  k.pos = "none")  +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +

  scale_fill_manual(values = col) + scale_color_manual(values = rep("black",3)) +

  annotate("text", x = x1, y = -1.5, label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 3), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 8) + 
         annotate("text", x = x1, y = 0.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = x1-0.20, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(f_cvr_mean), digit = 2))), hjust = 0, size = 8) + 
  annotate("text", x = x1-0.20, y = 1, label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 8) +

  annotate("text", x = x2, y = -1.5, label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 3), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 8) + 
     annotate("text", x = x2, y = 0.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = x2-0.20, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(m_cvr_mean), digit = 2))), hjust = 0, size = 8) +
  annotate("text", x = x2-0.20, y = 1, label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 8) +

  annotate("text", x = x3+0.10, y = -1.5, label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 3), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 8) + 
  annotate("text", x = x3+0.10, y = 0.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = x3-0.10, y = 1, label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 8) + 
  annotate("text", x = x3-0.10, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(t_cvr_mean), digit = 3))), hjust = 0, size = 8) +

    labs(shape = "Type") +
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.3, size = 5) + ylim (-1.5, 1.5)

  p3b <- p3b + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_blank(),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top") + guides(shape = "none")

  fig3 <- (p3a + p3b) + plot_annotation(tag_levels = "A")
  ggsave(fig3, filename = "./output/figures/fig3_r1.png", width = 21.88, height = 9)

  fig3 <- magick::image_read("./output/figures/fig3_r1.png")
  fig3
```

### *Relative variation in physiological rates is reduced in terrestrial and marine ectotherms*

Relative variance in physiological rates ($lnCVR_{Q_{10}}$) showed a
decrease with increasing temperature across all habitat types, which was
especially pronounced in terrestrial and marine ectotherms. The total
proportion of heterogeneity in $lnCVR_{Q_{10}}$ was lower compared to
$lnRR_{Q_{10}}$ ($I^2_{sv}$ = `r mean(tot_i2_cvr)`, 95% CI:
`r quantile(tot_i2_cvr, 0.025)` to `r quantile(tot_i2_cvr, 0.975)`),
with most variation being driven by between-study and trait differences
("Model 1"; see Supplement @fig-s3). There were no differences between
$lnCVR_{Q_{10_{acute}}}$ and $lnCVR_{Q_{10_{acclimation}}}$. Reductions
in variance were driven by slower increases in the variance in
physiological rates as mean physiological rates increased with scaling
relationships between log standard deviation and log mean all being
smaller than 1 and varying across habitats (see *Supplementary
Materials*; @tbl-slopes & @fig-mv). There was a `r mean(t_cvr_mean_ef)`%
(95% CI: `r quantile(t_cvr_mean_ef, 0.025)[1]` to
`r quantile(t_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(t_cvr_mean)`) reduction in relative physiological rate variance
for terrestrial ectotherms and a `r mean(m_cvr_mean_ef)`% (95% CI:
`r quantile(m_cvr_mean_ef, 0.025)[1]` to
`r quantile(m_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(m_cvr_mean)`) reduction in relative variation for marine
ectotherms when temperatures increased by 10$^\circ$C. In contrast,
freshwater ectotherms exhibited a smaller reduction in relative
physiological rate variance at high temperatures
(`r mean(f_cvr_mean_ef)`%, 95% CI: `r quantile(f_cvr_mean_ef, 0.025)[1]`
to `r quantile(f_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ =
`r pmcmc(f_cvr_mean)`) (results from "Model 2").

Across habitats whole-organism versus biochemical processes showed a
similar decrease in relative variance in physiological rates
([@fig-btrait]D-F; see Supplement @fig-S2) for more detailed trait
analysis), except for biochemical processes in ectotherms occupying
freshwater habitats, which showed small, albeit significant increase in
relative variance when acclimated ([@fig-btrait]F). Reduced relative
variance was particularly prominent for resting metabolic rates and
sprint speed although traits differed in whether they exhibited a
reduction in variation in physiological rates at high temperatures (see
*Supplemental Material*, @fig-S2) ("Model 3"). Each life-history stage
exhibited the same pattern of relative variance change in each of the
habitats ("Model 4"; @fig-cvrlh).

```{r, fig-btrait}
#| label: fig-btrait
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnCVR_{Q_{10}}$ (D-F) effect sizes for tissue/whole-orgamism traits and biochemical traits across marine (A & D), freshwater (B & E) and terrestrial (C & F) habitats. Across all plots, thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. k = total number of effect size estimates while the numbers in brackets indicate the number of species. For ease of visualisation, raw data for both trait categories are presented but points are not distinguished by different symbols. "
#| echo: false
#| warning: false

add_theme  <- function(){theme(axis.title  = element_text(size = 30),
                               axis.text.y = element_text(size = 30),
                               axis.text.x = element_text(size = 30),
                              legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 36))}

branch.size = 2.8
trunk.size = 1.8

n_lnrr_btrait  <- lnRRQ10_data  %>% filter(trait_broad != "other") %>% 
              group_by(habitat, trait_broad, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(trait_broad, type, habitat)
n_lncvr_btrait <- lnCVRQ10_data %>% filter(trait_broad != "other") %>% 
              group_by(habitat, trait_broad, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(trait_broad, type, habitat)

       col <- c("#0871B9", "#F3B40D", "#1BB908")
     x.pos <- c(0.7, 1.30, 1.70, 2.30)
size_annot <- 5.5

############
# lnRR
############
m_r <- n_lnrr_btrait %>% filter(habitat == "m")
f_r <- n_lnrr_btrait %>% filter(habitat == "f")
t_r <- n_lnrr_btrait %>% filter(habitat == "t")

t_trait_b <- contrast_lnrr_trait_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)
m_trait_b <- contrast_lnrr_trait_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)
f_trait_b <- contrast_lnrr_trait_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)

yloc = c(1.5, 2.5)

p1 <- orchard_plot(model1.3_m_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + labs(title = "Marine") + add_theme() + ylim(-0.8, 1.8) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", m_r$k, " (", m_r$spp, ")")), size = size_annot) + scale_fill_manual(values = rep(col[2],4)) + scale_color_manual(values = rep("black",4)) + annotate("text", x = yloc, y = -0.8, label = TeX(paste0("$\\beta$ = ", m_trait_b$diff, " [", m_trait_b$l95, " to ", m_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b$p)), hjust = 0, size = 8)

p2 <- orchard_plot(model1.3_f_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Freshwater")  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-0.5, 2) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", f_r$k, " (", f_r$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1)) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + scale_fill_manual(values = rep(col[1],4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -0.5, label = TeX(paste0("$\\beta$ = ", f_trait_b$diff, " [", f_trait_b$l95, " to ", f_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b$p)), hjust = 0, size = 8)

p3 <- orchard_plot(model1.3_t_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "t", legend.pos = "none"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Terrestrial") + add_theme()  +  ylim(-0.3, 1.8) + theme(legend.position = "none") + annotate("text", x = x.pos, y = 1.5, label = TeX(paste0("\\it{k} = ", t_r$k, " (", t_r$spp, ")")), size = size_annot) + scale_fill_manual(values = rep(col[3], 4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -0.25, label = TeX(paste0("$\\beta$ = ", t_trait_b$diff, " [", t_trait_b$l95, " to ", t_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_b$p)), hjust = 0, size = 8)

############
# CVR
############
m_cv <- n_lncvr_btrait %>% filter(habitat == "m")
f_cv <- n_lncvr_btrait %>% filter(habitat == "f")
t_cv <- n_lncvr_btrait %>% filter(habitat == "t")

t_trait_b_cvr <- contrast_lncvr_trait_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)
m_trait_b_cvr <- contrast_lncvr_trait_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)
f_trait_b_cvr <- contrast_lncvr_trait_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(trait)

p1_cvr <- orchard_plot(model2.3_m_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + ylim(c(-2, 2.2)) + add_theme() + ylim(-1.7, 1.6)  + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", m_cv$k, " (", m_cv$spp, ")")), size = size_annot) + theme(legend.position = "none", axis.text.y = element_blank()) + scale_fill_manual(values = rep(col[2],4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -1.7, label = TeX(paste0("$\\beta$ = ", m_trait_b_cvr$diff, " [", m_trait_b_cvr$l95, " to ", m_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b_cvr$p)), hjust = 0, size = 8)

p2_cvr <- orchard_plot(model2.3_f_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type", angle = 45, g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)  + add_theme() + ylim(-1.6, 1.3) + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", f_cv$k, " (", f_cv$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1), axis.text.y = element_blank()) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + scale_fill_manual(values = rep(col[1], 4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -1.6, label = TeX(paste0("$\\beta$ = ", f_trait_b_cvr$diff, " [", f_trait_b_cvr$l95, " to ", f_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b_cvr$p)), hjust = 0, size = 8)

p3_cvr <- orchard_plot(model2.3_t_broad, mod = "trait_broad", at = list(c_accl_time = 0, habitat = "t"), angle = 45, group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)+ add_theme()  + ylim(-2, 2) + annotate("text", x = x.pos, y = 1.5, label = TeX(paste0("\\it{k} = ", t_cv$k, " (", t_cv$spp, ")")), size = size_annot) +  theme(legend.position = "none") + scale_fill_manual(values = rep(col[3], 4)) + scale_color_manual(values = rep("black", 4)) + annotate("text", x = yloc, y = -2, label = TeX(paste0("$\\beta$ = ", t_trait_b_cvr$diff, " [", t_trait_b_cvr$l95, " to ", t_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_b_cvr$p)), hjust = 0, size = 8)

############
# Full fig
############
figbtrait <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(figbtrait, filename = "./output/figures/fig4_btrait_r1.png", width = 28, height = 14, device = "png")

figbtrait <- magick::image_read("./output/figures/fig4_btrait_r1.png")
figbtrait
```

```{r, fig-cvrlh}
#| label: fig-cvrlh
#| fig-cap: "Estimated mean acclimation and acute $lnRR_{Q_{10}}$ (A-C) and $lnCVR_{Q_{10}}$ (D-F) for adult (a) and juvenile (j) life-history stages for marine (A & D), freshwater (B & E) and terrestrial (C & F) ectotherms. Across all plots, thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. k = total number of effect size estimates while the numbers in brackets indicate the number of species. For ease of visualisation, raw data for both adult and juvenile life-history stages are presented but points are not distinguished by different symbols. "
#| echo: false
#| warning: false

add_theme  <- function(){theme(axis.title  = element_text(size = 30),
                               axis.text.y = element_text(size = 30),
                               axis.text.x = element_text(size = 30),
                              legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 36))}

branch.size = 2.8
trunk.size = 1.8

n_lnrr_lh  <- lnRRQ10_data  %>% group_by(habitat, life.hist, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(type, life.hist, habitat)
n_lncvr_lh <- lnCVRQ10_data %>% group_by(habitat, life.hist, type) %>% summarise(k = n(), spp = length(unique(spp))) %>% arrange(type, life.hist, habitat)

x.pos = c(0.7, 1.30, 1.70, 2.30)
size_annot <- 5.5

###############
## lnRR Q10
###############

t_trait_b <- contrast_stage_lnrr_lh_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
m_trait_b <- contrast_stage_lnrr_lh_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
f_trait_b <- contrast_stage_lnrr_lh_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)

yloc = c(1.5, 2.5)

m_r <- n_lnrr_lh %>% filter(habitat == "m")
f_r <- n_lnrr_lh %>% filter(habitat == "f")
t_r <- n_lnrr_lh %>% filter(habitat == "t")

p1 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "m"), group = "spp", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + labs(title = "Marine") + add_theme() + ylim(-0.4, 1.5) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", m_r$k, " (", m_r$spp, ")")), size = size_annot) + annotate("text", x = yloc, y = -0.4, label = TeX(paste0("$\\beta$ = ", m_trait_b$diff, " [", m_trait_b$l95, " to ", m_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b$p)), hjust = 0, size = 8)

p2 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Freshwater")  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-0.4, 1.5) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", f_r$k, " (", f_r$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1)) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + annotate("text", x = yloc, y = -0.4, label = TeX(paste0("$\\beta$ = ", f_trait_b$diff, " [", f_trait_b$l95, " to ", f_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b$p)), hjust = 0, size = 8)

p3 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "t", legend.pos = "none"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Terrestrial") + add_theme()  +  ylim(-0.3, 1.8) + theme(legend.position = "none") + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", t_r$k, " (", t_r$spp, ")")), size = size_annot) + annotate("text", x = yloc, y = -0.3, label = TeX(paste0("$\\beta$ = ", t_trait_b$diff, " [", t_trait_b$l95, " to ", t_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_b$p)), hjust = 0, size = 8) +scale_x_discrete(labels = c("A" = "Adult","J" = "Juvenile"))

###############
# ln CVR
###############
m_cv <- n_lncvr_lh %>% filter(habitat == "m")
f_cv <- n_lncvr_lh %>% filter(habitat == "f")
t_cv <- n_lncvr_lh %>% filter(habitat == "t")

  t_trait_cvr <- contrast_lncvr_lh_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
m_trait_b_cvr <- contrast_lncvr_lh_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
f_trait_b_cvr <- contrast_lncvr_lh_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)

p1_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + ylim(c(-2, 2.2)) + add_theme() + ylim(-1.5, 1.1)  + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", m_cv$k, " (", m_cv$spp, ")")), size = size_annot) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = yloc, y = -1.3, label = TeX(paste0("$\\beta$ = ", m_trait_b_cvr$diff, " [", m_trait_b_cvr$l95, " to ", m_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b_cvr$p)), hjust = 0, size = 8)

p2_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type", angle = 45, g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)  + add_theme() + ylim(-1.3, 1.1) + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", f_cv$k, " (", f_cv$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1), axis.text.y = element_blank()) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + annotate("text", x = yloc, y = -1.3, label = TeX(paste0("$\\beta$ = ", f_trait_b_cvr$diff, " [", f_trait_b_cvr$l95, " to ", f_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b_cvr$p)), hjust = 0, size = 8)

p3_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "t"), angle = 45, group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)+ add_theme()  + ylim(-1.55, 1.1) + annotate("text", x = x.pos, y = 0.6, label = TeX(paste0("\\it{k} = ", t_cv$k, " (", t_cv$spp, ")")), size = size_annot) +  theme(legend.position = "none") + annotate("text", x = yloc, y = -1.55, label = TeX(paste0("$\\beta$ = ", t_trait_cvr$diff, " [", t_trait_cvr$l95, " to ", t_trait_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_cvr$p)), hjust = 0, size = 8) + scale_x_discrete(labels = c("A" = "Adult","J" = "Juvenile"))

fig4 <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig4_lh_r1.pdf", width = 28, height = 14, device = "pdf")

fig4 <- magick::image_read("./output/figures/fig4_lh_r1.pdf")
fig4
```

```{r, wildlnRR10}
#| label: wildlnRR10
#| cache: true
#| include: false
#| echo: false
#| warning: false
#| 
#######################################
  ## lnRRQ10 from Wild Populations ##
#######################################
##################################################
  # Organise the data and prune tree
##################################################

# Bring in Climate data
climate <- read.csv("./output/climate_data/ERA5_air_and_SST_timeseries.csv")

# Create a match column.
climate2 <- climate %>% mutate(match = interaction(Species, Latitude, Longitude)) %>% dplyr::select(Species, Latitude, Longitude, match, cv, sd, mean, acf_lag1, acf_all)

# Subset the data
lnRRQ10_data_wild <- data_long2 %>% filter(grepl("lnRR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period) & source == "w") %>% 
                              mutate(spp = species_full, c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE), match = interaction(spp, lat, long), 
                              max_temp_c = scale(temp_2, scale = FALSE))

# Now, join climate data to full data. This will give us all the climate variability data for each site across time
lnRRQ10_data_wild <- left_join(lnRRQ10_data_wild, climate2, by = "match")

# Check and prune tree
tree_lnrr_wild <- tree_checks(data = lnRRQ10_data_wild, tree = tree_p0.7, dataCol = "species_full", type = "prune")
tree_checks(data = lnRRQ10_data_wild, tree = tree_lnrr, dataCol = "species_full", type = "check") # Good

# Create the phylo correlation matrix
phylo_cor_wild <- vcv(tree_lnrr_wild, cor = TRUE)
        A_wild <- vcv(tree_lnrr_wild, cor = FALSE) # For bayes, need cov matrix

# Scale data
        lnRRQ10_data_wild <- lnRRQ10_data_wild %>% mutate(cv_c = scale(cv),
                                                          acf_all_c = scale(acf_all))

# Check the environmental variability across habitats. What is the average SD in temps

envi_var <- lnRRQ10_data_wild %>% group_by(habitat) %>% summarise(mean_SD = mean(sd), 
                                                                  mean_pred = mean(acf_all),
                                                                  stdy = length(unique(record_num)), 
                                                                        n = n())

###################################################
  # model_envi: Check how tempeartures vary by habitat
  # Expect terrestrial and freshwater systems tp be more variable 
  # Than marine. Predcition for autocorrelation is less clear
###################################################
# Lets have a look at modelling whether the environmental variability and predcitability vary by habitats
if(rerun){
     sd_mod <- bf(log(sd) ~ habitat  + (1|record_num) + (1|spp))

  model_envi <- brms::brm(sd_mod, family = gaussian(),  control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 1, iter = 5000, warmup = 1000, cores = 4, data = lnRRQ10_data_wild)
  saveRDS(model_envi, file = "./output/models/model_envi")
} else {
  model_envi <- readRDS("./output/models/model_envi")
}

####################################################
  # Model 3.3: Model environmental variability by acute 
  #  acclim 
####################################################
  rerun = FALSE
if(rerun){
  # More realistic model with 
  model3.3          <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c + type + type:cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild)

   model3.3_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~ max_temp_c + c_accl_time + cv_c + type + type:cv_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_wild=A_wild), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_wild)

  # Save model3.3
  saveRDS(model3.3, file = "./output/models/lnRRQ10_3.3")
  saveRDS(model3.3_bayes, file = "./output/models/lnRRQ10_3.3_bayes")

  } else{
        model3.3 <- readRDS("./output/models/lnRRQ10_3.3")
  model3.3_bayes <- readRDS("./output/models/lnRRQ10_3.3_bayes")
}

############################################################################################
# Model 3.9: Acclimatization lnRRQ10 only. Test if environmental variance explains 
# differently in each habitat
############################################################################################          
lnRRQ10_data_wild_acclim <- lnRRQ10_data_wild %>%  filter(type == "acclim")

if(rerun){
  # More realistic model with 
  model3.9  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim)

   model3.9_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_acclim)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_wild_acclim=A_wild_acclim), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnRRQ10_data_wild_acclim)

  # Save model3.8
  saveRDS(model3.9, file = "./output/models/lnRRQ10_3.9")
  saveRDS(model3.9_bayes, file = "./output/models/lnRRQ10_3.9_bayes")

  } else{
        model3.9 <- readRDS("./output/models/lnRRQ10_3.9")
  model3.9_bayes <- readRDS("./output/models/lnRRQ10_3.9_bayes")
}

#################################################
### Bayesian Slopes from Model 3.9
#################################################
 posteriors <- posterior_samples(model3.9_bayes, "^b")

    slope_f <-     posteriors[,"b_cv_c"]; pmcmc(slope_f); mean(slope_f); quantile(slope_f, c(0.025, 0.975))
    slope_m <-     posteriors[,"b_cv_c"] + posteriors[,"b_cv_c:habitatm"]; pmcmc(slope_m); quantile(slope_m, c(0.025, 0.975))
    slope_t <-     posteriors[,"b_cv_c"] + posteriors[,"b_cv_c:habitatt"]; pmcmc(slope_t); quantile(slope_t, c(0.025, 0.975))

##################################################
    # Model 3.8, 3.9 and 3.10 - Habitat subset
##################################################
    # Subset data to habitat
    lnRRQ10_data_wild_acclim_t <- lnRRQ10_data_wild_acclim %>% filter(habitat == "t")
    lnRRQ10_data_wild_acclim_f <- lnRRQ10_data_wild_acclim %>% filter(habitat == "f")
    lnRRQ10_data_wild_acclim_m <- lnRRQ10_data_wild_acclim %>% filter(habitat == "m")

if(rerun){
    # Terrestrial Models
    model3.9_t  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim_t)
    saveRDS(model3.9_t, "./output/models/model3.9_t")

    # Freshwater Models

    model3.9_f  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim_f)

    saveRDS(model3.9_f, "./output/models/model3.9_f")

    # Marine Models

    model3.9_m  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim_m)
    saveRDS(model3.9_m, "./output/models/model3.9_m")

  } else {
     model3.9_t <- readRDS("./output/models/model3.9_t")
     model3.9_f <- readRDS("./output/models/model3.9_f")
     model3.9_m <- readRDS("./output/models/model3.9_m")
}

```

```{r wildlncvr}
#| label: wildlncvr
#| cache: true
#| include: false
#| echo: false
#| warning: false
#| 
#######################################
  ## lnCVRQ10 from Wild Populations ##
#######################################
##################################################
  # Organise the data and prune tree
##################################################

# Bring in  historial climate data
  climate <- read.csv("./output/climate_data/ERA5_air_and_SST_timeseries.csv")

# Create a match column.
  climate2 <- climate %>% mutate(match = interaction(Species, Latitude, Longitude)) %>% 
                          dplyr::select(Species, Latitude, Longitude, match, cv, sd, mean, acf_lag1, acf_all)

# Populations and species?
  pop_spp_w <- climate2[complete.cases(climate2),]

# Subset the data
  lnCVRQ10_data_wild <- data_long2 %>% 
    filter(grepl("lnCVR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period) & source == "w") %>% 
                        mutate(spp = species_full, 
                                     c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE), 
                                     match = interaction(spp, lat, long),
                                     max_temp_c = scale(temp_2, scale = FALSE))

# Now, join climate data to full data. This will give us all the climate variability data for each site across time
  lnCVRQ10_data_wild <- left_join(lnCVRQ10_data_wild, climate2, by = "match")

# Look at the environment details across different habitats. 
  lnCVRQ10_data_wild  %>% group_by(habitat) %>% summarise(mean(sd))

# Check and prune tree
  tree_lncvr_wild <- tree_checks(data = lnCVRQ10_data_wild, tree = tree_p0.7, dataCol = "species_full", type = "prune")
  tree_checks(data = lnCVRQ10_data_wild, tree = tree_lncvr_wild, dataCol = "species_full", type = "check") # Good

# Create the phylo correlation matrix
  phylo_cor_wild_cvr <- vcv(tree_lncvr_wild, cor = TRUE)
          A_wild_cvr <- vcv(tree_lncvr_wild, cor = FALSE) # For bayes, need cov matrix

# Scale environment variables
    lnCVRQ10_data_wild <- lnCVRQ10_data_wild %>% mutate(     cv_c = scale(cv),
                                                        acf_all_c = scale(acf_all),
                                                        acf_lag1_c = scale(acf_lag1))

####################################################
  # Model 4.3: Model environmental variability by acute/acclim
####################################################
rerun = FALSE
if(rerun){
  # More realistic model with 
  model4.3          <- metafor::rma.mv(effect_size ~ c_accl_time + cv_c + type + type:cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 | spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_cvr), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild)

   model4.3_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  c_accl_time + cv_c + type + type:cv_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_cvr)) + (1 | spp) + (1|trait), data2 = list(A_wild_cvr=A_wild_cvr), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnCVRQ10_data_wild)

  # Save model4.3
  saveRDS(model4.3, file = "./output/models/lnCVRQ10_4.3")
  saveRDS(model4.3_bayes, file = "./output/models/lnCVRQ10_4.3_bayes")

  } else{
        model4.3 <- readRDS("./output/models/lnCVRQ10_4.3")
  model4.3_bayes <- readRDS("./output/models/lnCVRQ10_4.3_bayes")
}

############################################################################################
# Model 4.9: Acclimatization lnCVRQ10 only. Test if environmental variance explains 
# differently in each habitat
############################################################################################          
  lnCVRQ10_data_wild_acclim <- lnCVRQ10_data_wild %>%  filter(type == "acclim")

  # Create the phylo correlation matrix

  # Check and prune tree
tree_lncvr_wild_acclim <- tree_checks(data = lnCVRQ10_data_wild_acclim, tree = tree_p0.7, dataCol = "species_full", type = "prune")
tree_checks(data = lnCVRQ10_data_wild_acclim, tree = tree_lncvr_wild_acclim, dataCol = "species_full", type = "check")

phylo_cor_wild_acclim <- vcv(tree_lncvr_wild_acclim, cor = TRUE)
        A_wild_acclim <- vcv(tree_lncvr_wild_acclim, cor = FALSE) # For bayes, need cov matrix

if(rerun){
  # More realistic model with 
  model4.9  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 | spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim)

   model4.9_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + cv_c + habitat + habitat:cv_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_acclim)) + (1 | spp) + (1|trait), data2 = list(A_wild_acclim=A_wild_acclim), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.99, max_treedepth = 12), chains = 4, thin = 1, iter = 5000, warmup = 1000, cores = 8, data = lnCVRQ10_data_wild_acclim)

  # Save model3.8
  saveRDS(model4.9, file = "./output/models/lnCVRQ10_4.9")
  saveRDS(model4.9_bayes, file = "./output/models/lnCVRQ10_4.9_bayes")

  } else{
        model4.9 <- readRDS("./output/models/lnCVRQ10_4.9")
  model4.9_bayes <- readRDS("./output/models/lnCVRQ10_4.9_bayes")
}

#################################################
### Bayesian Slopes from Model 4.9
#################################################
 posteriors4.9 <- posterior_samples(model4.9_bayes, "^b")

    slope_f_4.9 <- posteriors4.9[,"b_cv_c"]; mean(slope_f_4.9); pmcmc(slope_f_4.9); quantile(slope_f_4.9, c(0.025, 0.975))
    slope_m_4.9 <- posteriors4.9[,"b_cv_c"] + posteriors4.9[,"b_cv_c:habitatm"]; mean(slope_m_4.9); pmcmc(slope_m_4.9); quantile(slope_m_4.9, c(0.025, 0.975))
    slope_t_4.9 <- posteriors4.9[,"b_cv_c"] + posteriors4.9[,"b_cv_c:habitatt"]; mean(slope_t_4.9); pmcmc(slope_t_4.9); quantile(slope_t_4.9, c(0.025, 0.975))

##################################################
    # Model 4.8, 4.9 and 4.10 - Habitat subset
    # Used for making plots
##################################################
    # Subset data to habitat
    lnCVRQ10_data_wild_acclim_t <- lnCVRQ10_data_wild_acclim %>% filter(habitat == "t")
    lnCVRQ10_data_wild_acclim_f <- lnCVRQ10_data_wild_acclim %>% filter(habitat == "f")
    lnCVRQ10_data_wild_acclim_m <- lnCVRQ10_data_wild_acclim %>% filter(habitat == "m")

if(rerun){
    # Terrestrial Models
    model4.9_t  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim_t)
    saveRDS(model4.9_t, "./output/models/model4.9_t")

    # Freshwater Models

    model4.9_f  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim_f)

    saveRDS(model4.9_f, "./output/models/model4.9_f")

    # Marine Models

    model4.9_m  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + cv_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim_m)
    saveRDS(model4.9_m, "./output/models/model4.9_m")

  } else {
     model4.9_t <- readRDS("./output/models/model4.9_t")
     model4.9_f <- readRDS("./output/models/model4.9_f")
     model4.9_m <- readRDS("./output/models/model4.9_m")
}

```

### *Past climate does not influence acclimation capacity or expected change in variance*

Thermal variability (i.e., $CV$) experienced by a population in the past
did not explain acclimation capacity ([@fig-5]A) or changes in
physiological rate variance ([@fig-5]B) among terrestrial, marine or
freshwater populations ("Model 5").

```{r, fig-5}
#| label: fig-5
#| fig-align: center
#| fig-cap: "Predicted mean acclimation (thick black line) $lnRR_{Q_{10_{acclim}}}$ (A) and $lnCVR_{Q_{10_{acclim}}}$ (B) as a function of the Thermal Coefficient of Variation (CV) for wild populations across marine, freshwater and terrestrial habitats. Dashed lines indicate 95% confidence intervals and dotted lines indicate 95% prediction intervals. Model slope ($\\beta$) along with the 95% CI and $p_{MCMC}$ values for the slopes are shown for each habitat."

size_axis = 20
col <- c("#0871B9", "#F3B40D", "#1BB908")

####################
# lnRR CV
####################
 dat_text <- data.frame(label = c(text(slope_f), text(slope_m), text(slope_t)),
                      label2 = c(text(slope_f, b = FALSE), text(slope_m,b = FALSE), text(slope_t,b = FALSE)),
                      p = c(p_text(slope_f), p_text(slope_m), p_text(slope_t)),
                      title = c("Freshwater", "Marine", "Terrestrial"), 
                      condition = c("f", "m", "t"),
                      x = rep(0, 3),
                      y = rep(-1.1, 3))
y_adj = -3
# Terrestrial
  p_t_rr <-  bubble_plot(model3.9_t, mod = "cv_c", group = "spp", ylab = TeX("$lnRR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g = TRUE) + 
    geom_point(data = lnRRQ10_data_wild_acclim_t, aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), fill = col[3], colour = "black", stroke = 0.2, shape = 21) + labs(title = dat_text$title[3]) +
  annotate("text", x = dat_text$x[3], y = y_adj, label = TeX(dat_text$p[3]), size = 8) +
  annotate("text", x = dat_text$x[3], y = y_adj + 0.5, label = TeX(dat_text$label2[3]), size = 8) +
  annotate("text", x = dat_text$x[3], y = y_adj + 1, label = TeX(dat_text$label[3]), size = 8) +
theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis),
      axis.title.x = element_blank(),
      plot.title = element_text( size = 36, hjust = 0)) + ylim(-3,3)

# Marine
p_m_rr <-  bubble_plot(model3.9_m, mod = "cv_c", group = "spp", ylab = TeX("$lnRR_{Q_{10_{acclim}}}"), xlab = "Thermal Coefficient of Variation (CV)", g = TRUE) + 
    geom_point(data = lnRRQ10_data_wild_acclim_m, aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), fill = col[2], colour = "black", stroke = 0.2, shape = 21) +  labs(title = dat_text$title[2]) +
  annotate("text", x = dat_text$x[2]-0.5, y = y_adj, label = TeX(dat_text$p[2]), size = 8) +
  annotate("text", x = dat_text$x[2]-0.5, y = y_adj+0.5, label = TeX(dat_text$label2[2]), size = 8) +
  annotate("text", x = dat_text$x[2]-0.5, y = y_adj+1, label = TeX(dat_text$label[2]), size = 8) +
theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis), 
      axis.title.x = element_blank(),
      plot.title = element_text( size = 36, hjust = 0)) + ylim(-3, 3)

# Freshwater
p_f_rr <-  bubble_plot(model3.9_f, mod = "cv_c", group = "spp", ylab = TeX("$lnRR_{Q_{10_{acclim}}}"), xlab = "Thermal Coefficient of Variation (CV)", g = TRUE) + 
    geom_point(data = lnRRQ10_data_wild_acclim_f, aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), fill = col[1], colour = "black", stroke = 0.2, shape = 21) +  labs(title = dat_text$title[1]) +
  annotate("text", x = dat_text$x[1]+0.6, y = y_adj, label = TeX(dat_text$p[1]), size = 8) +
  annotate("text", x = dat_text$x[1]+0.6, y = y_adj+0.5, label = TeX(dat_text$label2[1]), size = 8) +
  annotate("text", x = dat_text$x[1]+0.6, y = y_adj+ 1, label = TeX(dat_text$label[1]), size = 8) +
theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis), 
      axis.title.x = element_blank(),
      plot.title = element_text( size = 36, hjust = 0)) + ylim(-3, 3)

####################
# lnCVR CV
####################
dat_textcvr <- data.frame(label = c(text(slope_f_4.9), 
                                 text(slope_m_4.9), 
                                 text(slope_t_4.9)),
                      label2 = c(text(slope_f_4.9, b = FALSE), 
                                 text(slope_m_4.9,b = FALSE), 
                                 text(slope_t_4.9,b = FALSE)),
                           p = c(p_text(slope_f_4.9), 
                                 p_text(slope_m_4.9), 
                                 p_text(slope_t_4.9)),
                       title = c("Freshwater", "Marine", "Terrestrial"), 
                      condition = c("f", "m", "t"),
                      x = rep(0, 3),
                      y = rep(-4, 3))

# Freshwater
  yadj2 = -6
  x_f = 0.6
  p_f_cv <-  bubble_plot(model4.9_f, mod = "cv_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g =FALSE, k = FALSE) + 
      geom_point(data = lnCVRQ10_data_wild_acclim_f, aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), fill = col[1], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textcvr$x[1]+x_f, y = yadj2, label = TeX(dat_textcvr$p[1]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_f, y = yadj2+0.8, label = TeX(dat_textcvr$label2[1]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_f, y = yadj2+1.6, label = TeX(dat_textcvr$label[1]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)

# Marine
  x_m = -0.4
  p_m_cv <-  bubble_plot(model4.9_m, mod = "cv_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g =FALSE, k = FALSE) + 
    geom_point(data = lnCVRQ10_data_wild_acclim_m, 
                 aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), 
                 fill = col[2], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textcvr$x[1]+x_m, y = yadj2, label = TeX(dat_textcvr$p[2]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_m, y = yadj2+0.8, label = TeX(dat_textcvr$label2[2]), size = 8) +
    annotate("text", x = dat_textcvr$x[1]+x_m, y = yadj2+1.6, label = TeX(dat_textcvr$label[2]), size = 8) +
    theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)

# Terrestrial

x_t = 0
p_t_cv <-  bubble_plot(model4.9_t, mod = "cv_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), xlab = "Thermal Coefficient of Variation (CV)", g = FALSE, k = FALSE) + 
  geom_point(data = lnCVRQ10_data_wild_acclim_t, 
               aes(x = cv_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[3], colour = "black", stroke = 0.2, shape = 21) +
  annotate("text", x = dat_textcvr$x[3]+x_t, y = yadj2, label = TeX(dat_textcvr$p[3]), size = 8) +
  annotate("text", x = dat_textcvr$x[3]+x_t, y = yadj2+0.8, label = TeX(dat_textcvr$label2[3]), size = 8) +
  annotate("text", x = dat_textcvr$x[3]+x_t, y = yadj2+1.6, label = TeX(dat_textcvr$label[3]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis), 
      plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
  ylim(-6, 3.5)

####################
# Full plot
####################

  p_cv <- (p_t_rr + p_m_rr + p_f_rr) / (p_t_cv + p_m_cv + p_f_cv) + 
            plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 36))

  ggsave(p_cv, width = 18, height = 10, filename = "./output/figures/fig_p_cv_r1.png")

  p_cv <- image_read("./output/figures/fig_p_cv_r1.png")
  p_cv 
```

### *Changes in physiological rate variance under climate change*

```{r, global}
#| label: global
#| echo: false
#| include: false

# Here, we want to make global predictions, at the lat/long in the data for how the opportunity of selection is expected to change with climate change. How can we do this? Q10 effects already have temperature controls within the effect size. So, what we want is to be able to understand how Q10 changes with latitude and longitude. Once we know that, we an extract how temperature is expected to change from our current levels to future levels. This will be in degrees C. Then, assuming no plasticity and evolution (big assumption) we can get a baseline change in variance for each site. We first need to adjust the predicted Q10 value so that it's at 1C temperature, not 10, we can do this through this formula Qt = (Q10 / 10)*T, where T is the temperature change in C. This will adjust the predicted Q10 value to the change in 1 degree C. We can then multiply this by the temperature change expected at that location to get the % change in variance or acure/acclim response. We can then plot these globally.

#############################
rerun = FALSE
if(rerun){

  # Model lat and long. Need to exclude missing acclimdation data. That will make it easier to create predcitions and explore the model fit. Include max temp to correct predictions for possible decrease at higher temps.

   model5_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type*habitat + t2(Longitude, Latitude) +
                               (1 | record_num) + 
                               (1 | gr(species_full, cov = A_wild_cvr)) + 
                               (1 + max_temp_c | spp) + 
                               (1 | trait) + 
                               (1 | obs), 
                      data2 = list(A_wild_cvr=A_wild_cvr), family = gaussian(), 
                      prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                      control = list(adapt_delta = 0.99, max_treedepth = 12), 
                      chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, 
                      data = lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),])

  saveRDS(model5_bayes, file = "./output/models/lnCVRQ10_5_bayes")

  } else{
  model5_bayes <- readRDS("./output/models/lnCVRQ10_5_bayes")
}

# Lets explore the model and how it's predicting the real data
  data_w <- lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),]

# combine with predictions
  data_w_pred <- cbind(data_w, predict(model5_bayes))

# Fitted vs actual data
  ggplot(data = data_w_pred, aes(x = Estimate, y = effect_size))      + 
        geom_point() + geom_smooth(method = "lm")                     + 
        labs(x = "Predicted Effect Size", y = "Observed Effect Size") + 
        theme_classic()                                               + 
        geom_abline(intercept = 0, slope = 1)

# Seems to be OK, but over and under-predicting at times. Lets explore the data for odd ball predictions to see if there are any obvious reasons why. There are very few for the most part but lets see why sign changes occur which are most striking differences. Sign changes indicate the observed data is completely opposite in sign to the predicted data

#  Spot check of a bunch of studies doesn't show any obvious issues with the data. Note that data is repeated because acute and acclimated rates are calculated. We expect the rows of raw data for acute and acclim to be repeated even though these data are NOT all used to make the calculation.
  sign_change_dat <- data_w_pred  %>% filter(sign(Estimate) != sign(effect_size))
  #View(sign_change_dat)

# Fitted by Lat/long Longitude, Latitude
  ggplot(data = data_w_pred, aes(y = Estimate, x = Latitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Predicted Effect Size") + theme_classic()

  ggplot(data = data_w_pred, aes(y = Estimate, x = Longitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Predicted Effect Size") + theme_classic()

# Observed by Lat/long
  ggplot(data = data_w_pred, aes(y = effect_size, x = Latitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Latitude", y = "Observed Effect Size") + theme_classic()

  ggplot(data = data_w_pred, aes(y = effect_size, x = Longitude)) + geom_point() + geom_smooth(method = "loess") + labs(x = "Longitude", y = "Observed Effect Size") + theme_classic()

# How much variation does this model explain. Need to do this manually because function doesn't seme to work on the model object. We need to remember that predcit will include random effect variances so we may need to exclude this
  re_post <- posterior_samples(model5_bayes, pars = "^sd_")
  v_fix <- apply(predict(model5_bayes, summary = FALSE), 1, function(x) var(x)) # 0.71 (but remember includes random effect variances)

# sigma2_v = typical sampling error variance
sub_data <- lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),]
  sigma2_v <- sum(1 / sub_data$V) * (dim(sub_data)[1] - 1) /
              (sum(1 / sub_data$V)^2 - sum((1 / sub_data$V)^2))

# Now check that var fix from raw data and remove sampling variance. 
  v_fix2 <- var(sub_data$effect_size) - sigma2_v  #0.58

# Extract the random effect variances
  v_trait <- re_post$sd_trait__Intercept^2
    v_spp <- re_post$sd_spp__Intercept^2
    v_phy <- re_post$sd_species_full__Intercept^2
    v_std <- re_post$sd_record_num__Intercept^2
      v_r <- re_post$sd_obs__Intercept^2

# Lets get the total variance for the random effects only
  tot_re <- v_trait + v_spp + v_phy + v_std + v_r

# Now, calculate r2 marginal, by subtracting the random effect variance from the fixed effect variance (which includes random effects). Need to check this as it may not be this intuitive
    r2 <- (v_fix - tot_re - sigma2_v) / (v_fix - sigma2_v) # Need to remove sampling variance too

    mean(r2) # Seems a bit high 0.53
  quantile(r2, probs = c(0.025, 0.975)); #0.30-0.72

# Now lets check with our different method. This should be close to the above. Which it is, so I think we're ok in our intuition.
  r2_2 <- (v_fix2 - tot_re) / (v_fix2) # Already removed sampling variance from v_fix2

  mean(r2_2) #  0.50
  quantile(r2_2, probs = c(0.025, 0.975)); #0.33-0.62

######################################################################
# Model predictions for terrestrial, marine and freshwater ectotherms
######################################################################
# Now we need to make predictions for new data. We first need to create new data. We need to do this for each habitat type because they are all a little different. Lets do terrestrial first.

#############################
## Terrestrial ectotherms
#############################

  terrest_data <- sub_data  %>% filter(habitat == "t")  %>% distinct(spp, species_full, habitat, type, Latitude, Longitude)
  new_data_t <- data.frame(
      max_temp_c = 0, 
      c_accl_time = 0,
      V = 0,
      trait = 0,
      obs = 0, 
      record_num = 0
    )

  # Create new data frame with only unique values for prediction.
  new_data_t2  <- cbind(terrest_data, new_data_t)

  # We'll now make the predictions from the model object using the new data frame
  terre_pred <- predict(model5_bayes, newdata = new_data_t2, summary = TRUE, allow_new_levels = TRUE, sample_new_levels = "uncertainty", robust = FALSE, intervals = "confidence")

  # We'll also keep the whole posterior which we'll use later on. We need to transpose so that the data match. Each numbered columns is one sample of the posterior.
  terre_pred_no_sum <- cbind(new_data_t2, t(predict(model5_bayes, newdata = new_data_t2, summary = FALSE, allow_new_levels = TRUE, sample_new_levels = "uncertainty")))

  # Lets merge predictions with the original data
  terrest_data <- cbind(terrest_data, terre_pred)

  # Test manual calculations. We can also set summary = FALSE and then do the calculations ourselves. Note we want robust = FALSE to make the comparisons more similar as otherwise the calculation are different when using summary TRUE/FALSE . These are extremely close, and I suspect are only different because sample_new_levels is sampling differently each time predict is run, so will be slightly different from run-to-run
  terre_pred2 <- predict(model5_bayes, newdata = new_data_t2, summary = FALSE, allow_new_levels = TRUE, sample_new_levels = "uncertainty", robust = FALSE)
  plyr::ldply(apply(terre_pred2, 2, function(x) post_summary(x)))

#############################
## Marine ectotherms  
#############################
  marine_data <- sub_data  %>% filter(habitat == "m") %>% distinct(spp, species_full, habitat, type, Latitude, Longitude)

  new_data_m <- data.frame(
       max_temp_c = 0, 
      c_accl_time = 0,
                V = 0,
            trait = 0,
              obs = 0, 
       record_num = 0
    )

  # Create new data frame with only unique values for prediction.
  new_data_m  <- cbind(marine_data, new_data_m)

  # We'll now make the predictions from the model object using the new data frame
  marine_pred <- predict(model5_bayes, newdata = new_data_m, summary = TRUE, allow_new_levels = TRUE, sample_new_levels = "uncertainty", robust = FALSE, intervals = "confidence")

  # We'll also keep the whole posterior which we'll use later on. We need to transpose so that the data match. Each numbered columns is one sample of the posterior.
  marine_pred_no_sum <- cbind(new_data_m, t(predict(model5_bayes, newdata = new_data_m, summary = FALSE, allow_new_levels = TRUE, sample_new_levels = "uncertainty")))

  # Lets merge predictions with the original data
  marine_data <- cbind(marine_data, marine_pred)

#############################
## Freshwater ectotherms
#############################
  fresh_data <- sub_data  %>% filter(habitat == "f") %>% distinct(spp, species_full, habitat, type, Latitude, Longitude)

  new_data_f <- data.frame(
       max_temp_c = 0, 
      c_accl_time = 0,
                V = 0,
            trait = 0,
              obs = 0, 
       record_num = 0
    )

  # Create new data frame with only unique values for prediction.
  new_data_f  <- cbind(fresh_data, new_data_f)

  # We'll now make the predictions from the model object using the new data frame
  fresh_pred <- predict(model5_bayes, newdata = new_data_f, summary = TRUE, allow_new_levels = TRUE, sample_new_levels = "uncertainty", robust = FALSE, intervals = "confidence")

  # We'll also keep the whole posterior which we'll use later on. We need to transpose so that the data match. Each numbered columns is one sample of the posterior.
  fresh_pred_non_sum <- cbind(new_data_f, t(predict(model5_bayes, newdata = new_data_f, summary = FALSE, allow_new_levels = TRUE, sample_new_levels = "uncertainty")))

  # Lets merge predictions with the original data
  fresh_data <- cbind(fresh_data, fresh_pred)

# Now that we have all the predictions we can merge the data together. Note that species itself is not really that useful because each prediction at each latitude and longitude should incorporate the species effect. So we'll remove it. While I retain spp, it's still not totally specific to the species, but left it there so it's clear how these data match with the wild data. Note there are only 134 species because this is subset of the wild data.
         all_pred <- rbind(terrest_data, marine_data, fresh_data) %>% 
                     dplyr::select(-species_full)  %>% 
                     mutate(uniques = interaction(spp, Latitude, Longitude))
  all_pred_no_sum <- rbind(terre_pred_no_sum, marine_pred_no_sum, fresh_pred_non_sum) %>%      dplyr::select(-species_full)
        spp_wild  <- length(unique(sub_data$spp))
```

```{r, future_clim}
#| label: future_clim
#| echo: false
#| warning: false
#| include: false

# Now we want to take the predictions from `global` and use this for future climate sceneios. We have the future temperatures expected for each location. We can then use the predicted Q10 values to calculate the expected change in variance. We can then plot these globally.

# First, lets get the current and future temperature data for 2081-2085. We'll use the mean. 
  future_temp <- read.csv("./output/climate_data/CanESM2_air_and_SST_timeseries.csv")

# Now, lets summarise the mean for the current temps (i.e., 2018-2022)
current_temps <-   future_temp  %>% 
                   dplyr::select(Species, Latitude, Longitude, grep("X2018|X2019|X2020|X2021|X2022", colnames(future_temp)))  %>% 
                                mutate(mean = apply(.[, grep("X2018|X2019|X2020|X2021|X2022", colnames(.))], 1, function(x) mean(x, na.rm = TRUE)), Celc_curr = mean - 273.15, match = interaction(Species, Latitude, Longitude))

# Lets clean this data up. We only need the mean temperature here, but this needs to be converted to Celcius as right now it's in Kelvins. We'll now take the last 5 years of the future temperature 2096-2100
  future_temps  <- future_temp  %>% 
                  dplyr::select(Species, Latitude, Longitude,  grep("X2096|X2097|X2098|X2099|X2100", colnames(future_temp)))  %>% 
                  mutate(mean = apply(.[, grep("X2096|X2097|X2098|X2099|X2100", colnames(.))], 1, function(x) mean(x, na.rm = TRUE)), Celc_future = mean - 273.15,  match = interaction(Species, Latitude, Longitude))

# Now we can merge the climate datasets together so that we have the species, lat and long, current and future climate. Then we can calculate the expected 'increase' in T into the future

  climate_data <- left_join(current_temps, future_temps, by = "match")  %>% dplyr::select(match, Latitude.x, Longitude.x, Species.x, Celc_curr, Celc_future)  %>% mutate(T_change = Celc_future - Celc_curr)  %>% rename(Species = Species.x, Latitude = Latitude.x, Longitude = Longitude.x)

# Lets look at future and current. Most, I would expect, should be above the line indicating higher future temperatures.
  ggplot(data = climate_data, aes(x = Celc_curr, y = Celc_future)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Current Temperature (C)", y = "Future Temperature (C)") + theme_classic() + geom_abline(intercept = 0, slope = 1)

# Lets have a global plot of temperature changes to see it sptially

world_map = map_data("world") %>% 
  filter(! long > 180)

  map2_temp_proj <- ggplot() +
    geom_map(
      data = world_map, map = world_map,
      aes(long, lat, map_id = region),
      color = "black", fill = "lightgray") +
    expand_limits(x = world_map$long, y = world_map$lat) +
    coord_map("moll") +
    geom_point(
      data = climate_data[complete.cases(climate_data$T_change),],
      aes(Longitude, Latitude, color = T_change),
      alpha = 0.7, size = 5
    ) + labs(x = "Longitude", y = "Latitude", color = TeX("$\\Delta$ Temp. ($^\\circ$C)"), title = "Current (2018-2022) vs Future (2096-2100)") + scale_colour_gradient2(trans = "reverse") + theme_map(base_size = 22) + theme(plot.title = element_text(hjust = 0.5))

# Now lets merge all_pred with the climate projections
  # Simplify columns 
  climate_data_new <- climate_data  %>% dplyr::select(-Latitude, -Longitude)

  # Merge with model predictions. This is the predicted lnCVRQ10 for 10C change AT the specific latitude and longitude. We must assume that there is no plasticity or adaptation at each site to make projections.
  all_pred2 <- all_pred %>% mutate(match = interaction(spp, Latitude, Longitude)) %>% 
              left_join(., climate_data_new, by = "match")  %>% 
              mutate(lnCVR_future = (Estimate / 10) * T_change,
                      lnCVR_F_l95 = (Q2.5 / 10) * T_change,
                      lnCVR_F_u95 = (Q97.5 / 10) * T_change, 
                      per_change = (exp(lnCVR_future) - 1) * 100,
                      per_change_l95 = (exp(lnCVR_F_l95) - 1) * 100,
                      per_change_u95 = (exp(lnCVR_F_u95) - 1) * 100)

  # First, lets merge the climate data with the predictions, but here predictions are not summarised. We have the entire posteior

    all_pred2_no_sum <-  all_pred_no_sum %>% 
                        mutate(match = interaction(spp, Latitude, Longitude))  %>% 
                        left_join(., climate_data_new, by = "match")

  # Second, lets convert each estimate to lncvr to 1 degree C instead of 10...
   all_pred2_no_sum2 <- all_pred2_no_sum %>%  
                        mutate_if(colnames(.) %in% c(1:3200), lncvr_1c)

  # Third, lets multiply each estimate by the temperature change at that location. I've checked 4 different cells and they are calculated correctly.
    all_pred2_no_sum2[,colnames(all_pred2_no_sum2) %in% c(1:3200)]  <- all_pred2_no_sum2[,colnames(all_pred2_no_sum2) %in% c(1:3200)] * all_pred2_no_sum2$T_change

  # Fourth, we can calculate the percentage change on the posterior
    all_pred2_no_sum2[,paste0("perc", c(1:3200))]  <- (exp(all_pred2_no_sum2[,colnames(all_pred2_no_sum2) %in% c(1:3200)]) - 1)*100

  # Now we can summarise each. Here, what we are doing is taking the mean for each column. Effectively pooling across all the iterations for each dataset. This gives us an estimation of the mean change in variance across all locations/species, but it gives us 3200 point estimates of this mean. We can then use this to calculate the 95% CI's.
    mean_perc <-  apply(all_pred2_no_sum2[,paste0("perc", c(1:3200))], 1, mean, na.rm = TRUE) 
    mean(mean_perc)
    quantile(mean_perc, 0.025)
    quantile(mean_perc, 0.975)
    hist(mean_perc)

  # Now, lets do this but for each habitat, the good old fashioned way. I think that these are so variable because the expected change in the future is quite dramatic, so the variability is high.
  mean_perc_habitat <-  lapply(split(all_pred2_no_sum2,  all_pred2_no_sum2$habitat), function(x) apply(x[,paste0("perc", c(1:3200))], 1, mean, na.rm = TRUE))
  lapply(mean_perc_habitat, mean)
  lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))
  #lapply(mean_perc_habitat, hist)

# Create a table to summarise everything
tables_future <- all_pred2 %>% group_by(habitat) %>% summarise(mean = mean(per_change, na.rm = TRUE), 
                                                                l95 = mean(per_change_l95, na.rm = TRUE), 
                                                                u95 = mean(per_change_u95, na.rm = TRUE))

# Check on how many to decrease
  decreases <- all_pred2 %>% group_by(match)  %>% summarise(n = if_else(any(per_change < 0), 1, 0))

```

Measurements of acute and acclimation responses from wild ectotherms
were much less common than from captive populations ($N_{species}$ =
`r length(unique(all_pred$spp))`, from
`r length(unique(all_pred$uniques))` wild populations). Globally, there
was a clear bias towards species in the Northern Hemisphere
([@fig-fig6]A-C). Projected changes in physiological rate variance were
highly variable across the globe, with some regions showing a decrease
in physiological rate variance, while others showing an increase
([@fig-fig6]D). However, out of the `r length(unique(all_pred2$match))`
populations, relative variance was predicted to decrease in
`r (sum(decreases$n, na.rm = TRUE) / length(decreases$n[complete.cases(decreases$n)]))*100`%
of the locations.

Using the ERA5 climate model, predictions of current global changes in
physiological rate variance were generally conservative with our model
explaining \~ 50% of the variation in the observed data ($R^2$ =
`r round(mean(r2_2), 2)`, 95% CI: `r round(quantile(r2_2, 0.025)[1], 2)`
to `r round(quantile(r2_2, 0.975)[1], 2)`). Climate change is predicted
to result in a `r abs(lapply(mean_perc_habitat, mean)$f)`% reduction in
relative variance for freshwater systems (95% CI:
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[1]`
to
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[2]`%,
$p_{MCMC}$ = `r pmcmc(mean_perc_habitat$f)`), a
`r abs(lapply(mean_perc_habitat, mean)$m)`% reduction in relative
variance for marine systems (95% CI:
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[1]`
to
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[2]`%,
$p_{MCMC}$ = `r pmcmc(mean_perc_habitat$m)`), and a
`r abs(lapply(mean_perc_habitat, mean)$t)`% reduction in relative
variance for terrestrial systems (95% CI:
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[1]`
to
`r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[2]`%,
$p_{MCMC}$ = `r p_value(pmcmc(mean_perc_habitat$t))`) under a RCP8.5
climate scenario. All results are taken from "Model 6".

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "Model predictions for the expected change in acclimation $lnCVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).  Predicted change in relative physiological rate variance for each population based on current temperatures (average from 2018-2022; A-C) as well as the expected change from current temperatures based on future temperature predictions (average from 2096-2100, D). Future climate predictions are the reduction in relative variance expected under a RCP8.5 climate scenario relative to current climate conditions (% change)."

#devtools::install_github("hughjonesd/ggmagnify") #https://github.com/hughjonesd/ggmagnify

world_map = map_data("world") %>% 
  filter(! long > 180)

map1_current <- ggplot() +
  geom_map(
    data = world_map, map = world_map,
    aes(long, lat, map_id = region),
    color = "black", fill = "lightgray") +
  expand_limits(x = world_map$long, y = world_map$lat) +
  coord_map("moll") +
  geom_point(
    data = all_pred  %>% filter(type == "acclim"),
    aes(Longitude, Latitude, color = Estimate),
    alpha = 0.7, size = 5
  ) + labs(x = "Longitude", y = "Latitude", color = TeX("$lnCVR_{Q_{10}}$"), title = "Current (2018-2022)") + scale_color_viridis() + theme_map(base_size = 22) +  theme(plot.title = element_text(hjust = 0.5))

zoom_areas <- data.frame(
     x = c(-130, -15, -60, 145),
  xend = c( -50, 48, -35, 155),
     y = c(  25, 34, 0, -45),
  yend = c(  55,  69.8, -60, -10),
  label = c("North America", "Europe", "South America", "Australia")
)

zoom_plots <- lapply(1:nrow(zoom_areas), function(i) {
  map1_current +
    geom_polygon(fill = "gray80", color = "black") +
    coord_cartesian(xlim = c(zoom_areas[i, "x"], zoom_areas[i, "xend"]), 
                    ylim = c(zoom_areas[i, "y"], zoom_areas[i, "yend"])) +
                    theme(legend.position = "none",
                        plot.title = element_blank())
})

map2_future <- ggplot() +
    geom_map(
      data = world_map, map = world_map,
      aes(long, lat, map_id = region),
      color = "black", fill = "lightgray") +
    expand_limits(x = world_map$long, y = world_map$lat) +
    coord_map("moll") +
    geom_point(
      data = all_pred2[complete.cases(all_pred2$per_change),],
      aes(Longitude, Latitude, color = per_change),
      alpha = 0.7, size = 5
    ) + labs(x = "Longitude", y = "Latitude", color = TeX("% $\\Delta$"), title = "Future (2096-2100)") + scale_colour_viridis() + theme_map(base_size = 22) + theme(plot.title = element_text(hjust = 0.5))

# Make a multipanel plot. Make the world_map plot the biggest, and then add the zoomed in plots arond it using pactchwork

p <- (((zoom_plots[[1]] | zoom_plots[[2]]) / map1_current) + map2_future) + plot_layout(heights  = c(1, 2, 2)) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20, face = "bold"), plot.margin = unit(c(0, 0, 0, 0), "cm"))

ggsave(p, filename = "./output/figures/fig-preds.png", width = 7.8, height = 10.1)

image_read("./output/figures/fig-preds.png")
```

## **Discussion**

Understanding acclimation capacity and how variation in physiological
rates change across populations and species is important for predicting
the ecological and evolutionary consequences of climate change
[@Chevin2017; @Chevin2010; @Bolnick2011; @bush2016incorporating;
@seebacher2023can; @Sanderson2023; @urban2023and]. Here, acclimation
responses ($lnRR_{Q_{10}}$) varied across habitats while relative
variance ($lnCVR_{Q_{10}}$) in physiological rates of ectotherms
decreased across terrestrial, marine and freshwater habitats as
temperatures increase. Decreases in relative variance were particularly
pronounced in terrestrial ectotherms (\~28%). These effects are expected
to result in a decrease in the variability in physiological rates for
marine and terrestrial ectotherms between \~4-13% under future climate
change projections. Our results uncover an hitherto unrecognised dynamic
where the benefits of acclimation may be counteracted by shifts in trait
variance impacting how ectotherm populations across different habitats
will be able to cope with increased temperatures.

### ***Acclimation capacities and changes in relative variability in physiological rates vary among habitats***

We show that acclimation of physiological rates and changes in relative
variance differ across habitats. Our findings confirm previous results
that quantify the different capacity of terrestrial, marine and
freshwater ectotherms to acclimate [@Seebacher2015-wi;
@Gunderson2015-hz; @morley2019physiological]. Our analysis generally
supports that of @Seebacher2015-wi, @Gunderson2015-hz and
@morley2019physiological that all show a general inability of
terrestrial ectotherms to physiologically acclimate. The change in
acclimation $Q_{10}$ we found was similar to @Seebacher2015-wi for
freshwater organisms (\~17%), but higher in marine ectotherms (decrease
of 16% versus \~10% in @Seebacher2015-wi), and opposite in terrestrial
ectotherms (increase of \~6% compared to an \~8% decrease in
@Seebacher2015-wi), possibly due to the use of new $Q_{10}$ based effect
sizes that allow us to control for sampling variance. Greater capacity
for acclimation in aquatic organisms may be the result of fewer
opportunities for behavioural thermoregulation in aquatic environments
making physiological remodeling important for maintaining homeostasis
[@Gunderson2015-hz; @morley2019physiological]. While marine and
freshwater ectotherms were capable of partial acclimation it is still
important to recognise that, on average, the effect size was small
(amounting to $Q_{10}$ dropping from \~1.8 to 1.6).

Changes in relative variance in physiological rates across habitats was
nearly equivalent (marine) or larger (terrestrial) compared to their
capacity to acclimate. Only in freshwater ectotherms did acclimation in
physiological rates coincide with no change in relative variation.
Mechanistically, it is unclear what is contributing to the reduction in
relative variation in physiological rates at higher temperatures. One
possible explanation may be that biochemical reactions are operating
more efficiently or are relying on more conformationally stable enzymes
at higher temperatures impacting between individual variation
[@tattersall2012coping; @angilletta2009thermal; @somero1995proteins;
@fields2001protein]. Such a mechanism may also help explain the
differences in relative variance across habitat types. For example,
terrestrial ectotherms may rely on more specialised enzymatic variants
that operate over a more limited temperature range individually but
together allow for organisms to function over a wider thermal range
necessary to operate in more variable thermal environments
[@angilletta2009thermal].

Interestingly, we show that biochemical traits (i.e., enzymes) were more
likely to acclimate compared to whole-organisms traits as predicted, but
only for marine ectotherms. In contrast, both whole-organism and
biochemical traits were equally likely to acclimate in freshwater
ectotherms. Unsurprisingly, we observed no distinction between traits in
terrestrial ectotherms. We also expected relative variance to differ
between whole-organism and biochemical traits, but we did not find
strong support for this except for in freshwater ectotherms where
biochemical traits were more likely to maintain or increase relative
variance compared with whole-organisms traits. Acclimation responses in
marine and freshwater organisms may buffer / maintain variance in
physiological rates at higher temperatures due to the expression of
alternative enzyme forms or intracellular stabalising compounds/solutes
that impact enzyme function [@fields2001protein].

Our results further highlight the potential vulnerability of terrestrial
ectotherms to climate change, while calling into question if acclimation
in marine systems will be able to counteract the potentially negative
effects of reductions in relative variance. Assuming changes in relative
variation in physiological rates are underpinned by genetic variation,
and there is a genetic correlation with fitness, such reductions in
physiological variance could limit a populations ability to adapt to
future climates [@urban2023and]. Of course, considering how behavioural
modifications buffer organisms from high temperatures is also needed.
Terrestrial species likely have greater opportunity for behavioural
microhabitat selection in thermally heterogeneous environments which is
expected to decrease the impacts of climate change [@huey2012predicting;
@sunday2014thermal]. Importantly, while we seen a reduction in relative
variation, absolute variation in physiological rates did increase with
temperature (@fig-s1) -- albeit to a much lesser extent in terrestrial
and marine habitats compared to freshwater -- supporting the general
notion that capacity for adaptation may still be limited in terrestrial
and marine habitats. As of yet, it is unclear whether a focus on
relative or absolute variance should be prioritized. Future theoretical
models are needed to better understand the ecological and evolutionary
consequences of changes in both mean and variance.

### ***Plasticity and relative variance in physiological rates do not differ between life stages***

Acclimation capacities are expected to differ between life-stages
because of distinct of patterns of dispersal, habitat use and behaviour
that force earlier life stages to cope with more variable environmental
conditions which can also lead to developmental constraints on how
physiological systems respond later in life [@martin2015age;
@stearns1976life; @angilletta2009thermal; @noble2018developmental;
@o2019developmental; @Pottier2022]. In addition, plastic responses are
also expected to be costly [@angilletta2009thermal; @Dewitt1998], and
such costs can be magnified in late life reducing the capacity for
plasticity [e.g., @rossi2019ageing]. These processes can also result in
changes to intrapopulation variation in physiological rates at higher
temperatures but the direction of change between early and adult life
stages is likely to depend on the costs of adjusting physiological
processes, energy reserves at different life stages, and the extent to
which early life experiences constrain plasticity.

Despite these expectations, more broadly, we show little difference
between early and late life acclimation capacities and little change in
the relative variance in physiological rates across habitats. This may
not be too surprising given that such responses are likely context or
trait dependent [@carter2020life; @moghadam2019heat]. The lack of
differences we observed may also be due to the fact that both juvenile
and adult animals occupy similar thermal niches, disperse to a similar
extent and exhibit comparable thermoregulatory behaviors making
physiological responses to temperature similar. It is possible that
temperatures experienced earlier in development (e.g., egg or larval
stages) exhibit greater capacity for acclimation because of limited
dispersal capabilities and very different habitats. Some empirical
evidence supports this in insects. For example, @moghadam2019heat showed
that larval *Drosophila* were more plastic in their heat hardening
responses compared to adults. A focus on collecting more detailed
information on the differences in behaviour, dispersal and thermal
environments experienced across different life stages is likely to
provide a more complete picture on when plasticity differs across life
stages. To date, however, there are few studies that have focused on how
variability in responses change. We would also encourage more empirical
focus on this questions given the potential ecological and evolutionary
implications.

### ***No signature of past climate on capacity for physiological plasticity***

Theoretical models predict that plasticity should evolve in populations
experiencing greater environmental variability (spatial or temporal),
particularly when oscillations are predictable over time to make
environmental cues more reliable [@Chevin2017; @Lande2009; @Chevin2010;
@Nunney2016; @Murren2015; @hendry2016key; @Reed2010]. Higher spatial and
temporal heterogeneity in terrestrial habitats [@steele2019comparison]
might suggest that plasticity is more likely to evolve in terrestrial
environments. However, if thermal variability is too high and
unpredictable, the rates of acclimation slow, and there are costs
associated with re-modelling physiological processes [which seem likely,
@angilletta2009thermal] then we would instead expect canalised responses
[@angilletta2009thermal; @Seebacher2015-wi; @loughland2020differences;
@Leung2020; @rescan2022experimental; @leung2023phenotypic]. The lack of
acclimation in terrestrial ectotherms we observed is consistent with the
latter hypothesis, and is supported by other meta-analyses of heat
tolerance [@barley2021limited; @Gunderson2015-hz] suggesting that there
are costs to being plastic or that the environmental signals are
insufficient to trigger endocrine and epigenetic mechanisms that lead to
plasticity when environments are not predictable [@hendry2016key;
@Leung2020].

Whether a populations capacity for physiological plasticity and the
thermal variability (or predictability) it experiences are related has
received more mixed results. We show no relationship between acclimation
capacity and thermal variability in marine, freshwater and terrestrial
habitats. Our results are consistent with @Gunderson2015-hz who show no
relationship between plasticity in heat tolerance and latitude or
thermal seasonality. However, other analyses on heat tolerance limits
have found relationships between latitude (a proxy for seasonality)
[@morley2019physiological] or measures of thermal variability
[@verberk2024heat]. @Seebacher2015-wi also found evidence that
acclimation capacity is related to a populations thermal variability,
however, relationships depended on the habitat and traits in question.
Discrepancies across studies could be related to the taxa included in
analyses [e.g., @morley2019physiological], different traits or possibly
the fact that different climate projections/models are being used to
quantify thermal variability. Modelling realistic microenvironments
across such diverse taxa is also challenging because it is unclear what
the most appropriate spatial and temporal scale might be that is of
evolutionary relevance. As such, it is also likely that historical
temperature time series' are not representative of the selective
environment a population has experienced making relationships between
capacity for acclimation and temperature variability (or predictability)
difficult to pin down.

### ***Conclusions and future directions***

Enhanced knowledge of how variation in physiological rates vary across
populations and species, and the degree to which they can be adjusted in
response to the environment, may lead to more informed predictions about
the ecological and evolutionary dynamics of natural populations
[@forsman2015rethinking; @seebacher2023can; @cooke2021one;
@Sanderson2023]. While we show general patterns across taxa and habitats
it is important to recognise that this does not mean such patterns apply
to all populations. Substantial variation in acclimation responses and
changes in variance exist among populations and traits, as evidenced by
wide prediction intervals and substantial study- and trait-level
variance estimates. Nonetheless, quantitative measures on the changes in
the relative variance in physiological rates could be better
incorporated into physiological and ecological models to provide more
nuanced, and possibly more realistic, predictions about the impacts of
climate change on natural populations. While we do not yet understand
the relative contribution of environmental and genetic factors to
variance changes, models could better decouple how different levels of
heritability and total variance impact evolutionary and ecological
predictions. Our meta-analysis now provides the opportunity to
parameterise such models, and ensure they are better aligned with
empirical findings.

Nonetheless, it is clear that many fascinating questions remain
unanswered that will require greater focus on the consequences of
changes in variance (rather than just the mean). Particularly
interesting questions include: How do differences in physiological rate
variance change energy flow across trophic levels within communities?
What are the biochemical, cellular, and physiological mechanisms that
underlie differences in physiological rate variance across habitats? Are
changes in variance in one trait associated with changes in other
traits, or do some traits increase while others decrease? Are changes in
physiological rate variance correlated with changes in genetic
variation? Answers to these questions will require integrative
approaches that combine empirical and theoretical work across multiple
levels of biological organisation but will likely provide useful
advances in understanding the full consequences that climate change will
have on ectotherms across major ecosystems globally.

## Acknowledgements

We would like to thank Malgorzata Lagisz for helping sort out the
phylogeny used in our analysis. We are also grateful to three anonymous
reviewers who all provided highly constructive feedback that greatly
improved the manuscript.

## Funding

DWAN is supported by an Australian Research Council (ARC) Future
Fellowship (FT220100276) and ARC Discovery Project (DP210101152). SN is
supported by ARC Discovery Project Grants (DP210100812 and DP230101248).

## Data and code availability

All data and code used to reproduce analyses can be found on GitHub at:
https://github.com/daniel1noble/Q10_meta_analysis and is deposited in
Zenodo (https://doi.org/10.5281/zenodo.11123600).

## Author contributions

Conceptualization: DWAN, FK, FS, SN; Methodology: DWAN, AB, FK, FS, SN;
Investigation: DWAN, FK, FS, SN; Visualization: DWAN; Supervision: DWAN,
SN, FS; Writing—original draft: DWAN; Writing—review & editing: DWAN,
AB, FK, FS, SN.

## Conflict of interest

Authors declare that they have no competing interests.

## References

<br> 
<div id ="refs"></div>
<br>

## Supplemental Results and Figures

#### *Literature Search Protocol and PRISMA flow diagram*

We performed a literature search using the Web of Science database for
articles or proceedings papers published in English from 2013 to 2017
(the date after [@Seebacher2015-wi] searches were conducted) using the
following topic search string: *"(acclimat* AND (therm\* OR temp*) NOT
(plant* OR tree\* OR forest\* OR fung\* OR mammal\* OR marsup\* OR
bird\* OR human OR exercis\* OR train\* OR hypoxi*))"*. We further
limited to the following research areas: Anatomy Morphology;
Biodiversity Conservation; Biology; Ecology; Endocrinology Metabolism;
Entomology; Evolutionary Biology; Marine Freshwater Biology; Physiology;
Respiratory System, Reproductive Biology, Zoology.

Our search resulted in 1,321 papers for screening in Rayyan
[@Ouzzani2016-ge]. We also cross-checked papers we found in our searches
with a recent paper by @Havird2020-mj, which also updates the dataset of
@Seebacher2015-wi's. We included any papers that were missed between our
searches and those of @Havird2020-mj. @Havird2020-mj added 7 new studies
(mainly because they were focused on metabolic rates), and our searches
differed from theirs by only a single paper [i.e., @Bulgarella2015-yl].
Given the physiological traits we included were broader than
@Havird2020-mj, we had a substantial increase in additional papers that
we added to @Seebacher2015-wi's dataset. More specifically, in addition
to the `r length(unique(old_papers$record_num))` papers we included from
the @Seebacher2015-wi dataset, we extracted data from an extra
`r length(unique(new_papers$record_num))` papers (with a total of
`r dim(new_papers)[1]` effects; a
`r  (length(unique(new_papers$record_num)) / length(unique(old_papers$record_num)))*100`%
increase in the number of published articles). Note that
@Seebacher2015-wi included a total of 205 publications, however, not all
these contained the necessary statistics we needed to derive effect
sizes and associated sampling variances (see below). While we may have
missed papers, our goal was to obtain a large representative (and
unbiased) sample of acclimation research rather than a comprehensive
dataset. As such, our database represents the most up-to-date dataset
used since @Seebacher2015-wi to answer questions on physiological rates
across ectotherms.

We split the screening of titles and abstracts for the 1,321 papers
found in our search among DWAN, FK, FS, and SN evenly. To ensure
consistency among authors in title and abstract inclusion, relevant
authors went through a randomly selected set of papers together before
the formal screening to calibrate selection of papers based on our
inclusion criteria (see below). In cases of disagreement regarding
inclusion, we conservatively included the paper for full text screening
and discussed uncertain papers among authors to come to a decision.
After title and abstract screening, there was a total of 149 papers for
full text screening. Papers were included only if they: 1) measured a
physiological rate acutely at two temperatures on a sample of animals
chronically exposed to the same two temperatures for at least 1 week;
and 2) where physiological rates measured were burst and sustained
locomotion, metabolic rates (standard, resting, routine and maximal),
heart rates, and/or enzyme activities. Importantly, as in
@Seebacher2015-wi, we only included studies that manipulated
temperatures within normal thermal ranges for the species.

```{r, fig-prisma}
#| label: fig-prisma
#| fig-cap: "PRISMA flow diagram of the literature search and screening process."
#| echo: false

prisma <- magick::image_read("./output/figures/PRISMA.png")
prisma
```

#### *Further discussions on the assumptions of* $lnRR_{Q_{10}}$, $lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ estimates

$lnRR_{Q_{10}}$, $lnCVR_{Q_{10}}$ and $lnVR_{Q_{10}}$, as with $Q_{10}$
more generally, all assume that the effect of temperature on
physiological rates (or changes in relative variance) is log-linear.
While this is likely in our data given that we restricted our analysis
of $Q_{10}$ to standard operating temperatures for a given species, it
may not always be satisfied given the diversity of species in our
dataset. $Q_{10}$ [@van1884etudes] has been used exstensively in the
physiological literature to successfully address a multitude of
questions [e.g., @Havird2020-mj; @Seebacher2015-wi]. However, there is a
preference for using a Boltzmann – Arrhenius (BA) relationship (or its
more recent extension, the Sharpe-Schoolfield model [@molnar2017thermal;
@michaletz2024hotter]) to model thermal effects on physiological rates
[@gillooly2001effects; @michaletz2024hotter]. While debate still exists
over the utility of $Q_{10}$ when modelling temperature-dependence it is
important to recognise that both BA and $Q_{10}$ can exhibit
curvilinearity as temperatures increase (as discussed in
[@michaletz2024hotter]). @white2012information also showed that the BA
model may not always perform better. For example, in eukaryotes,
modelling thermal dependence using $Q_{10}$ provided a 5.8-fold better
fit to metabolic rate data than the BA relationship
[@white2012information]. Given that studies included in our analysis
never measured full performance curves at acute and acclimation
temperatures it was not possible for us to compare different models of
thermal dependence. Nonetheless, $Q_{10}$-based effect sizes remain the
most practical effect-size for comparing thermal dependence when using
existing empircial data, with the benefit that these effects having
convenient properties that make them suitable for meta-analysis.
Nonetheless, we control for possible violations of the log-linearity
assumption in our analyses.

#### *Exploring the impact of maximum treatment temperature on* $lnRR_{Q_{10}}$

```{r, fig-lnrrq10_temp}
#| label: fig-lnrrq10_temp
#| fig-cap: "Bubble plot of the relationship between $lnRR_{Q_{10}}$ and maximum temperature used in treatments within a study."

mod_table <- mod_results(model1.2, mod = "max_temp_c", group = "record_num", at = list(c_accl_time = 0))$mod_table  %>% mutate(moderator = moderator + mean(lnRRQ10_data$temp_2, na.rm = TRUE))

col <- c("#0871B9", "#F3B40D", "#1BB908")
lnrrq10_temp <- ggplot(data = lnRRQ10_data, aes(x = temp_2, y = effect_size)) + 
geom_point(aes(size = 1/sqrt(V), fill = habitat), alpha = 0.8, shape = 21, colour = "black") +
labs(x = TeX("Maximum Temperature ($^{\\circ}$C)"), y = TeX("$lnRR_{Q_{10}}"), size = "Precision (1/SE)", color = "Habitat") + 
ggplot2::geom_smooth(data = mod_table, 
ggplot2::aes(x = moderator , y = lowerPR), method = "loess", 
            formula = y ~ x, se = FALSE, lty = "dotted", colour = "black") + 
ggplot2::geom_smooth(data = mod_table, ggplot2::aes(x = moderator, y = upperPR), method = "loess", 
            formula = y ~ x, se = FALSE, lty = "dotted", colour = "black") + 
ggplot2::geom_smooth(data = mod_table, ggplot2::aes(x = moderator, y = lowerCL), method = "loess", 
            formula = y ~ x, se = FALSE, lty = "dashed", colour = "black") + 
ggplot2::geom_smooth(data = mod_table, ggplot2::aes(x = moderator, y = upperCL), method = "loess", 
            formula = y ~ x, se = FALSE, lty = "dashed", colour = "black") + 
ggplot2::geom_smooth(data = mod_table, ggplot2::aes(x = moderator, y = estimate), method = "loess", 
            formula = y ~ x, se = FALSE, colour = "black") + 
            scale_fill_manual(labels = c("f" = "Freshwater", "m" = "Marine", "t" = "Terrestrial"), values = col)+ labs(fill = "Habitat") +
theme_classic() + theme(axis.title  = element_text(size = 30), 
                        axis.text = element_text(size = 26), 
                        axis.text.y = element_text(size = 26),
                        legend.text = element_text(size = 24), 
                        legend.title = element_text(size = 26))

ggsave(lnrrq10_temp, file = "./output/figures/lnrrq10_temp.png", height = 6, width = 8)

lnrrq10_temp <- magick::image_read("./output/figures/lnrrq10_temp.png")
lnrrq10_temp

```

#### *How acclimation time is related to* $lnRR_{Q_{10}}$

```{r, fig-accl_time}
#| label: fig-accl_time
#| fig-cap: "Bubble plot of the relationship between $lnRR_{Q_{10}}$ and acclimation time for terrestrial (green), marine (orange) and freshwater (blue) habitats. Acclimation time is centered around the mean acclimation time (37.5 days) in the data. Not all studies reported acclimation time hence the total number of effects, k, was 1767."
  col <- c("#0871B9", "#F3B40D", "#1BB908")

  accl_time <- bubble_plot(model1.2, mod = "c_accl_time", group = "spp", ylab = TeX("$lnRR_{Q_{10}}$"), xlab = "Centered Acclimaton Time (days)") + geom_point(data = lnRRQ10_data, aes(x = c_accl_time, y = effect_size, fill = habitat), shape = 21, colour = "black") + scale_fill_manual(labels = c("f" = "Freshwater", "m" = "Marine", "t" = "Terrestrial"), values = col) + theme_classic() + theme(axis.title  = element_text(size = 30), 
                        axis.text = element_text(size = 26), 
                        axis.text.y = element_text(size = 26),
                        legend.text = element_text(size = 24), 
                        legend.title = element_text(size = 26))

  ggsave(accl_time, file = "./output/figures/accl_time.png", height = 6, width = 8.5)

accl_time <- magick::image_read("./output/figures/accl_time.png")
accl_time
```

#### *Comparing cool and warm acclimated acute responses*

```{r, @fig-Sacute}
#| label: fig-Sacute
#| fig.width: 12.7
#| fig.height: 8.7
#| fig-cap: "Mean acute $lnRR_{Q_{10}}$ for cool (blue) and warm (red) acclimated populations for terrestrial (diamonds), marine (square) and freshwater (circle) habitats. Note that points  in each category show the full distribution of data irrespective of habitat for simplicity. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. Note that means for all three habitats are displayed but there is weak evidence that the means differ between habitats given models with and without an interaction with habitat are equally supported. Note that x-axis is truncated for ease of visualisation. Sample sizes for each habitat for acute warm and cold are: marine [warm = (131, 38, 29), cold = (136, 41, 32)], freshwater [warm = (294, 76, 61), cold = (293, 77, 63)], terrestrial [warm = (83, 31, 35), cold = (84, 31, 35)]. Numbers within brackets are number of effects, number of studies and number of species."
#| echo: false

col_acute <- c("#3F5FAC", "#EF3D25")
orchard_plot(model1.s1_int, mod = "name", by = "habitat", xlab = TeX("$lnRR_{Q_{10}}$"), group = "spp", angle = 45, trunk.size = 1.8, branch.size = 2.8, alpha = 0.4, condition.lab = "Habitat") + scale_x_discrete(labels = c("LnRR_Q10_acute_t2" = "Warm - Acute", "LnRR_Q10_acute_t1" = "Cool - Acute")) + scale_fill_manual(values = col_acute) + scale_colour_manual(values = col_acute) + ylim (-0.5, 2.5) +  theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top") 
                               #annotate("text", label= paste0("k = ", k_lnrr_acute$k, " (", k_lnrr_acute$sdy, ")"), x = c(1.2, 2.2), y = 2)
```

#### *Acute and acclimation* $lnRR_{Q_{10}}$ for detailed trait categories across marine, freshwater and terrestrial taxa

In addition to the broader trait categories we fit models to understand
how acute and acclimation effect sizes varied across trait categories.
To achieve this, we categorized each effect size into one of 12 trait
categories. These categories included measures of whole organism
performance measures including cardiac (i.e., 'cardiac') and muscle
('muscle') function, sprint speed ('sprint') and endurance ('endurance')
and metabolic rates (i.e., maximal and resting metabolic rate; max MR',
'rest MR', respectively). Studies also quantified various enzymatic
reaction rates, including enzymes involved in general metabolic
responses (categorized as 'metabolic enzyme'), various parts of the
electron transport chain, including ATPase activity ('ATPase'),
mitochondrial leak ('Proton Leak') and oxidation ('OXPHOS', short for
Oxidative Phosphorylation), as well as antioxidant enzymes ('antiox').
All other traits not falling within these categories were placed into
'other'.

@fig-lnrrtrait; @fig-cvrtrait

```{r, detail_trait}
#| label: detail_trait
#| echo: false
#| include: false
#| warning: false

## Detailed trait category analyses for lnRR

  # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnRRQ10_data_m <- lnRRQ10_data %>% filter(habitat == "m" & !trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_f <- lnRRQ10_data %>% filter(habitat == "f" & !is.na(c_accl_time))
  lnRRQ10_data_t <- lnRRQ10_data %>% filter(habitat == "t" & !is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR", "other"))

  # Prune tree
   tree_m <- tree_checks(lnRRQ10_data_m, tree = tree_lnrr, dataCol = "species_full", type = "prune")
   tree_f <- tree_checks(lnRRQ10_data_f, tree = tree_lnrr, dataCol = "species_full", type = "prune")
   tree_t <- tree_checks(lnRRQ10_data_t, tree = tree_lnrr, dataCol = "species_full", type = "prune")

   # cor matrices
   m_phylo <- vcv(tree_m, corr = TRUE)
       A_m <- vcv(tree_m, corr = FALSE)
   f_phylo <- vcv(tree_f, corr = TRUE)
       A_f <- vcv(tree_f, corr = FALSE)
   t_phylo <- vcv(tree_t, corr = TRUE)
       A_t <- vcv(tree_t, corr = FALSE)

rerun = FALSE
if(rerun){
#################################
  # Detailed Trait Categories
  # For supplement
#################################
  # More realistic model with traits in marine species. Remove antioxidants because there are just too few effects. Note that, for orchardplot we can only fit one random slope, so we'll just use obs as results are unchanged and Bayesian models have it accounted for
   model1.3_m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = m_phylo), dfs = "contain", 
                              test = "t", 
                              data = lnRRQ10_data_m)

  # Robust variance estimator.
    model1.3_m_rve <- robust(model1.3_m, cluster = lnRRQ10_data_m$record_num) # Mostly Consistent but some changes to p for some traits.

  # Baeysian model
  model1.3_m_het <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = m_phylo), 
                              struct = c("GEN", "HCS"),  phi = 0, dfs = "contain", 
                              test = "t", data = lnRRQ10_data_m, control = list(iter.max=1000, rel.tol=1e-8))

  # Check k is the same for both models (yes) and check which model is best supported
    AICc1.3_m     <- AIC(model1.3_m) + AICc_correction(model1.3_m)
    AICc1.3_m_het <- AIC(model1.3_m_het) + AICc_correction(model1.3_m_het)
  # Clear support for this model. We'll go with this.
    AICc1.3_m - AICc1.3_m_het

  # Cannot model heterogeneous variance across traits if use sigma + TRUE, so just use obs.
   model1.3_m_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category + 
                                   (~1 | record_num) + 
                                   (1 | gr(species_full, cov = A_m)) + 
                                   (1 + max_temp_c | spp) +(1|trait) + 
                                   (1 + trait_category |gr(obs, cor = FALSE)),  
                                 data2 = list(A_m=A_m), family = gaussian(), 
                                 prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                                 control = list(adapt_delta = 0.98, max_treedepth = 12), 
                                 chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_m)

    # More realistic model with traits in freshwater species
  model1.3_f <- metafor::rma.mv(effect_size ~  max_temp_c+c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full =f_phylo), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_f)

  # Heteroscedastic variance model for freshwater
  model1.3_f_het <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full =f_phylo), struct = c("GEN", "HCS"), phi = 0, dfs = "contain", 
                              test = "t", data = lnRRQ10_data_f, method = "REML", control=list(iter.max=1000, rel.tol=1e-8))

  ## Check support for two models
  AICc_model1.3_f_het <- AIC(model1.3_f_het) + AICc_correction(model1.3_f_het) # Clearly better supported.
      AICc_model1.3_f <- AIC(model1.3_f) + AICc_correction(model1.3_f)
  delat_AICc_model1.3_f <- AICc_model1.3_f - AICc_model1.3_f_het

  # Run the Bayesian equivalent of the top supported model. 
   model1.3_f_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category + 
                                   (~1 | record_num) + 
                                   (1 | gr(species_full, cov = A_f)) + 
                                   (1 + max_temp_c| spp) + (1|trait) + 
                                   (1+trait_category|gr(obs, cor = FALSE)), 
                                 data2 = list(A_f=A_f), family = gaussian(), 
                                 prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                                 control = list(adapt_delta = 0.98, max_treedepth = 12), 
                                 chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_f)

    # More realistic model with traits in terrestrial species
  model1.3_t <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full =t_phylo), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_t)

  # Heteroscedastic variance model for terrestrial
  model1.3_t_het <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full =t_phylo), struct = c("GEN", "HCS"),  phi = 0, dfs = "contain", 
                              test = "t", data = lnRRQ10_data_t)

  # Check which model is better
    AICc_model1.3_t     <- AIC(model1.3_t) + AICc_correction(model1.3_t)
    AICc_model1.3_t_het <- AIC(model1.3_t_het) + AICc_correction(model1.3_t_het)

  #  Delta
    AICc_model1.3_t - AICc_model1.3_t_het

  # Bayesian model
   model1.3_t_bayes <- brms::brm(effect_size | se(sqrt(V)) ~ max_temp_c + c_accl_time + type + trait_category + type:trait_category + 
                                   (~1 | record_num) + 
                                   (1 | gr(species_full, cov = A_t)) + 
                                   (1 + max_temp_c| spp) + 
                                   (1 | trait) + 
                                   (1 + trait_category | gr(obs, cor = FALSE)), 
                                 data2 = list(A_t=A_t), family = gaussian(), 
                                 prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                                 control = list(adapt_delta = 0.98, max_treedepth = 12), 
                                 chains = 8, thin = 5, iter = 3000, warmup = 1000, cores = 8, data = lnRRQ10_data_t)

    # Save models
  saveRDS(model1.3_m, file = "./output/models/lnRRQ10_1.3m")
  saveRDS(model1.3_m_het, file = "./output/models/lnRRQ10_1.3m_het")
  saveRDS(model1.3_m_bayes, file = "./output/models/lnRRQ10_1.3_bayes_m")

  saveRDS(model1.3_f, file = "./output/models/lnRRQ10_1.3f")
  saveRDS(model1.3_f_het, file = "./output/models/lnRRQ10_1.3f_het")
  saveRDS(model1.3_f_bayes, file = "./output/models/lnRRQ10_1.3_bayes_f")

  saveRDS(model1.3_t, file = "./output/models/lnRRQ10_1.3t")
  saveRDS(model1.3_t_het, file = "./output/models/lnRRQ10_1.3t_het")
  saveRDS(model1.3_t_bayes, file = "./output/models/lnRRQ10_1.3_bayes_t")

  } else{
        model1.3_m <- readRDS("./output/models/lnRRQ10_1.3m")
    model1.3_m_het <- readRDS("./output/models/lnRRQ10_1.3m_het")
  model1.3_m_bayes <- readRDS("./output/models/lnRRQ10_1.3_bayes_m")
        model1.3_f <- readRDS("./output/models/lnRRQ10_1.3f")
    model1.3_f_het <- readRDS("./output/models/lnRRQ10_1.3f_het")
  model1.3_f_bayes <- readRDS("./output/models/lnRRQ10_1.3_bayes_f")
        model1.3_t <- readRDS("./output/models/lnRRQ10_1.3t")
    model1.3_t_het <- readRDS("./output/models/lnRRQ10_1.3t_het")
  model1.3_t_bayes <- readRDS("./output/models/lnRRQ10_1.3_bayes_t")
}

## Detailed trait category analyses for lnCVR

  # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnCVRQ10_data_m <- lnCVRQ10_data %>% filter(habitat == "m" & !trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnCVRQ10_data_f <- lnCVRQ10_data %>% filter(habitat == "f" & !is.na(c_accl_time))
  lnCVRQ10_data_t <- lnCVRQ10_data %>% filter(habitat == "t" & !is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR", "other"))

  # Prune tree
   tree_m_cvr <- tree_checks(lnCVRQ10_data_m, tree = tree_lncvr, dataCol = "species_full", type = "prune")
   tree_f_cvr <- tree_checks(lnCVRQ10_data_f, tree = tree_lncvr, dataCol = "species_full", type = "prune")
   tree_t_cvr <- tree_checks(lnCVRQ10_data_t, tree = tree_lncvr, dataCol = "species_full", type = "prune")

   # cor matrices
   m_phylo_cvr <- vcv(tree_m_cvr, corr = TRUE)
   f_phylo_cvr <- vcv(tree_f_cvr, corr = TRUE)
   t_phylo_cvr <- vcv(tree_t_cvr, corr = TRUE)

if(rerun){
#########################
  # Detailed Trait Categories
  # Models for the supplement
#########################
  # More realistic model 
  model2.3 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_CVR), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data)

   model2.3_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category + 
                                 (~1 | record_num) + 
                                 (1 | gr(species_full, cov = A_CVR)) + 
                                 (1 + max_temp_c| spp) + 
                                 (1|trait), 
                               data2 = list(A_CVR=A_CVR), family = gaussian(), 
                               prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), 
                               control = list(adapt_delta = 0.98, max_treedepth = 12), 
                               chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnCVRQ10_data)

  # Save model1.2
  saveRDS(model2.3, file = "./output/models/lnCVRQ10_2.3")
  saveRDS(model2.3_bayes, file = "./output/models/lnCVRQ10_2.3_bayes")

  model2.3f <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = f_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"), phi = 0, data = lnCVRQ10_data_f)

  saveRDS(model2.3f, file = "./output/models/lnCVRQ10_model2.3f")

  model2.3m <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category + type:trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 + trait_category| obs),
                              R = list(species_full = m_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnCVRQ10_data_m)
  saveRDS(model2.3m, file = "./output/models/lnCVRQ10_model2.3m")

  # Note that we need to drop interaction for terrestrial as it's not able to estimate. 
  model2.3t <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + trait_category, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c | spp,
                                            ~1 | trait,
                                            ~1 + trait_category | obs),
                              R = list(species_full = t_phylo_cvr), dfs = "contain", 
                              test = "t", struct = c("GEN", "HCS"),  phi = 0, data = lnCVRQ10_data_t)
  saveRDS(model2.3t, file = "./output/models/lnCVRQ10_model2.3t")

  } else{
#########################
  # Detailed Trait Categories
  # Models for the supplement
#########################
        model2.3 <- readRDS("./output/models/lnCVRQ10_2.3")
        model2.3f <- readRDS("./output/models/lnCVRQ10_model2.3f")
        model2.3m <- readRDS("./output/models/lnCVRQ10_model2.3m")
        model2.3t <- readRDS("./output/models/lnCVRQ10_model2.3t")
  model2.3_bayes <- readRDS("./output/models/lnCVRQ10_2.3_bayes")
}
```

```{r, fig-lnrrtrait}
#| label: fig-lnrrtrait
#| fig-cap: "Acute and Acclimation $lnRR_{Q_{10}}$ across detailed trait categories for A) marine, B) freshwater and C) terrestrial systems. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. The x-axis is truncated for ease of visualisation. See methods section “Moderator Variables” for a full description of the trait categories. "
#| echo: false

k_marine=manual_extract(model1.3_m_het, mod = "trait_category", group = "record_num", exclude = "Antiox")
k_fresh=manual_extract(model1.3_f_het, mod = "trait_category", group = "record_num")
k_terrs=manual_extract(model1.3_t_het, mod = "trait_category", group = "record_num")

ptm <- orchard_plot(model1.3_m_het, mod = "trait_category", at = list(type = c("acute", "acclim")), by = "type", group = "record_num", weights = "prop", k.pos = "none", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 0, condition.lab = "Type") +
  scale_x_discrete(labels = c("Mito_leak" = "Proton Leak",
                              "Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate",
                              "Mito_oxidation" = "OXPHOS")) + theme_classic() + theme(axis.text = element_text(size=12)) + ylim(-1.5, 2.5) +
                            annotate("text", label= paste0("k = ", k_marine$k, " (", k_marine$sdy, ")"), x = k_marine$moderator, y = 2.4) + add_theme()
#ptm$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

ptf <- orchard_plot(model1.3_f_het, mod = "trait_category", at = list(type = c("acute", "acclim")), by = "type", group = "record_num", weights = "prop", k.pos = "none", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 0, condition.lab = "Type") +
  scale_x_discrete(labels = c("Mito_leak" = "Proton Leak",
                              "Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate",
                              "Mito_oxidation" = "OXPHOS")) + theme_classic() + theme() + 
                              annotate("text", label= paste0("k = ", k_fresh$k, " (", k_fresh$sdy, ")"), x = k_fresh$moderator, y = 2.4) + add_theme()
#ptf$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

ptt <- orchard_plot(model1.3_t_het, mod = "trait_category", at = list(type = c("acute", "acclim")), by = "type", group = "record_num", weights = "prop", k.pos = "none",  xlab = TeX("$lnRR_{Q_{10}}$"), angle = 0, condition.lab = "Type") +
  scale_x_discrete(labels = c("Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate")) + theme_classic() + labs(legend = "Type")+ annotate("text", label= paste0("k = ", k_terrs$k, " (", k_terrs$sdy, ")"), x = k_terrs$moderator, y = 2.4) + add_theme() + ylim(-1,3)
#ptt$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

fig2 <- (ptm / ptf / ptt) + plot_annotation(tag_levels = "A")
ggsave(fig2, filename = "./output/figures/fig2_trait.png", height = 20, width = 15)

img_fig2 <- magick::image_read("./output/figures/fig2_trait.png")
img_fig2
```

```{r, fig-cvrtrait}
#| label: fig-cvrtrait
#| fig-cap: "Acute and acclimation $lnCVR_{Q_{10}}$ across traits for A) marine, B) freshwater and C) terrestrial systems. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. The x-axis is truncated for ease of visualisation. See methods section “Moderator Variables” for a full description of the trait categories. "
#| echo: false

k_marinecv=manual_extract(model2.3m, mod = "trait_category", group = "record_num", exclude = "Antiox")
k_freshcv=manual_extract(model2.3f, mod = "trait_category", group = "record_num")
k_terrscv=manual_extract(model2.3t, mod = "trait_category", group = "record_num")

ptm <- orchard_plot(model2.3m, mod = "trait_category", group = "record_num", weights = "prop",  xlab = TeX("$lnCVR_{Q_{10}}$"), angle = 0, k.pos = "none", condition.lab = "Type") +
  scale_x_discrete(labels = c("Mito_leak" = "Proton Leak",
                              "Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate",
                              "Mito_oxidation" = "OXPHOS")) + theme_classic() + add_theme() + annotate("text", label= paste0("k = ", k_marinecv$k, " (", k_marinecv$sdy, ")"), x = k_marinecv$moderator, y = 2.4) + ylim(-2, 2.6)
#ptm$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

ptf <- orchard_plot(model2.3f, mod = "trait_category",  group = "record_num", weights = "prop",  xlab = TeX("$lnCVR_{Q_{10}}$"), angle = 0, condition.lab = "Type",  k.pos = "none") +
  scale_x_discrete(labels = c("Mito_leak" = "Proton Leak",
                              "Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate",
                              "Mito_oxidation" = "OXPHOS")) + theme_classic()+ add_theme() + annotate("text", label= paste0("k = ", k_freshcv$k, " (", k_freshcv$sdy, ")"), x = k_freshcv$moderator, y = 2.4) + ylim(-2.5, 2.65)
#ptf$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

ptt <- orchard_plot(model2.3t, mod = "trait_category",  group = "record_num", weights = "prop",  xlab = TeX("$lnCVR_{Q_{10}}$"), angle = 0,  k.pos = "none", condition.lab = "Type") +
  scale_x_discrete(labels = c("Metabolic_enzyme" = "Metabolic Enzymes",
                              "Rest_MR" = "Resting Metabolic Rate",
                              "Max_MR" = "Maximum Metabolic Rate")) + theme_classic() +  annotate("text", label= paste0("k = ", k_terrscv$k, " (", k_terrscv$sdy, ")"), x = k_terrscv$moderator, y = 2.4) + labs(legend = "Type")+ add_theme() + ylim(-2, 2.65)
#ptt$layers[[1]] <- NULL # Lets just remove the data point geom layer as it's too busy

figcvrtrait <- (ptm / ptf / ptt) + plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(figcvrtrait, filename = "./output/figures/figcvrtrait.png", height = 20, width = 15)

img_figcvrtrait <- magick::image_read("./output/figures/figcvrtrait.png")
img_figcvrtrait
```

#### *Mean-variance relationships to understand patterns in* $lnCVR_{Q_{10}}$

```{r, mc_models}
#| label: mc_models
#| echo: false
#| include: false

 # Model mean variance relationships to estimate the relationship. Two additonal models to address reviewer comments about what exactly is changing in the effect size
rerun = FALSE
if(rerun) {

      r_11_m <- bf(log(r1.1_sd) ~ log(r1.1)*habitat + (1|record_num) + (1|spp) + (1|trait)) + gaussian()
      r_12_m <- bf(log(r1.2_sd) ~ log(r1.2)*habitat + (1|record_num) + (1|spp) + (1|trait)) + gaussian()
      r_21_m <- bf(log(r2.1_sd) ~ log(r2.1)*habitat + (1|record_num) + (1|spp) + (1|trait)) + gaussian()
      r_22_m <- bf(log(r2.2_sd) ~ log(r2.2)*habitat + (1|record_num) + (1|spp) + (1|trait)) + gaussian()

      mv_model_all <- brm(r_11_m + r_12_m + r_21_m + r_22_m + set_rescor(TRUE), 
                          chains = 4, cores = 4, iter = 4000, warmup = 1000, thin = 1, 
                          control = list(adapt_delta = 0.98), data = lnRRQ10_data)
       saveRDS(mv_model_all, file = "./output/models/mv_model_all")

} else {
        mv_model_all <- readRDS("./output/models/mv_model_all")
}

# Post processing
  model_mv1.1 <- posterior_samples(mv_model_all,  pars = "^b_logr11") # Acclimated at 1, measured at 1
  model_mv1.2 <- posterior_samples(mv_model_all,  pars = "^b_logr12") # Acclimated at 1, measured at 2; Acute
  model_mv2.2 <- posterior_samples(mv_model_all,  pars = "^b_logr22") # Acclimated at 2, measured at 2
  model_mv2.1 <- posterior_samples(mv_model_all,  pars = "^b_logr21") # Acclimated at 2, measured at 1; Acute

# Extract slopes
    slope_extract <- function(x, v){
        list = list()
         list[['fresh']] <-  x[,2]  # Slope for freshwater
        list[['marine']] <-  x[,2] + x[, grepl(":habitatm", colnames(x))] # Slope for marine
         list[['terrs']] <-  x[,2] + x[, grepl(":habitatt", colnames(x))] # Slope for terrestrial
         return(list)
    }

# From each model, lets get the slopes for each habitat
         r1.1 <-  slope_extract(model_mv1.1)
         r1.2 <-  slope_extract(model_mv1.2)
         r2.1 <-  slope_extract(model_mv2.1)
         r2.2 <-  slope_extract(model_mv2.2)

# Summary
         table <- data.frame(
                               mean = do.call("rbind", lapply(r1.1,  mean)),
                               L = do.call("rbind", lapply(r1.1,  function(x) quantile(x, 0.025))),
                               U = do.call("rbind", lapply(r1.1,  function(x) quantile(x, 0.975))))

         r_1.2 = data.frame( mean = do.call("rbind", lapply(r1.2,  mean)),
                               L = do.call("rbind", lapply(r1.2,  function(x) quantile(x, 0.025))),
                               U = do.call("rbind", lapply(r1.2,  function(x) quantile(x, 0.975))))
         r_2.1 = data.frame(mean = do.call("rbind", lapply(r2.1,  mean)),
                               L = do.call("rbind", lapply(r2.1,  function(x) quantile(x, 0.025))),
                               U = do.call("rbind", lapply(r2.1,  function(x) quantile(x, 0.975))))
          r_2.2 = data.frame(mean = do.call("rbind", lapply(r2.2,  mean)),
                               L = do.call("rbind", lapply(r2.2,  function(x) quantile(x, 0.025))),
                               U = do.call("rbind", lapply(r2.2,  function(x) quantile(x, 0.975))))

          table <- rbind(table, r_1.2,  r_2.1, r_2.2)
          table$habitat <- rep(c("Freshwater", "Marine", "Terrestrial"), 4)
          table$type <- c(rep("r1.1",3),
                          rep("r1.2", 3),
                          rep("r2.1",3),
                          rep("r2.2", 3))
          rownames(table) <- NULL

    table <- table %>% dplyr::select(type, habitat, mean, X2.5., X97.5.) %>% rename("2.5%" = X2.5.,
                                                                                    "97.5%" = X97.5.,
                                                                                    "Slope" = mean,
                                                                                    "Treatment" = type,
                                                                                    "Habitat" = habitat) %>% 
      mutate(Type = ifelse(Treatment %in% c("r1.1", "r2.2"), "Acclimation", ifelse(Treatment %in% c("r1.2", "r2.1"), "Acute", " "))) %>% relocate(Type, .before = Treatment) %>% dplyr::arrange(Habitat)
```

We explored mean-variance relationships for the acute and acclimation
responses across all traits and habitats. We estimated the scaling
relationship between log standard deviation in physiological rates
\[log(SD)\] and log mean physiological rates \[log(mean)\], accounting
for non-independence resulting from effects coming from the same
species, study and traits (i.e., random effects of species, study and
trait) as done in our main analyses. We also included an interaction
between log(mean) and habitat type to better understand how the scaling
relationship between log(SD) and log(mean) varies across habitats.

Overall, we found that the relationship between log(mean) and log(SD) of
the acute and acclimation responses was generally linear, suggesting
that the use of $lnCVR_{Q_{10}}$ was justified (@fig-mv). Overall, the
scaling relationship between log(SD) and log(mean) was sub-linear across
all habitats (@tbl-slopes), however, ectotherms from terrestrial
habitats had much shallower slopes than marine and freshwater
ectotherms, particularly at higher treatment temperatures, indicating
increased mean physiological rates generally do not result in higher
between individual variance in physiological rates (@tbl-slopes).
Interestingly, in marine ectotherms the slope was highest at cooler
temperatures, wheeras the slope was suppressed when acclimated and/or
measured at higher temperatures (i.e., r.1. compared to r1.2, r2.1,
r2.2) (@tbl-slopes). In freshwater ectotherms, there we some differences
in scaling relationships but they were all fairly comparable no matter
what acclimatin and test temperature (@tbl-slopes).

```{r, fig-mv}
  #| label: fig-mv
  #| fig-cap: "Mean-standard deviation relationships for the acute and acclimation responses across all habitats. Relationships are depicted for the low temperature treatment of the study (A) and high temperature treatment (B)."
    col <- c("#0871B9", "#F3B40D", "#1BB908")

  mv1 <- ggplot2::ggplot(lnRRQ10_data, aes(x = log(r1.1), y = log(r1.1_sd),color = habitat)) +
                  geom_point(aes(color = habitat), size = 2.2) + 
                  geom_smooth(method = "lm", se = FALSE) + 
                  labs(x = "log (Mean)", y = "log (SD)", color = "Habitat") + 
                  scale_colour_manual(values = col) + 
                  facet_wrap(~type) + theme_classic() + theme(axis.text = element_text(size=14), axis.title = element_text(size = 16), 
                  legend.text = element_text(size = 14), legend.title = element_text(size = 16), strip.text = element_text(size = 16))

  mv2 <- ggplot2::ggplot(lnRRQ10_data, aes(x = log(r1.2), y = log(r1.2_sd), color = habitat)) +
                  geom_point(aes(color = habitat), size = 2.2) + 
                  geom_smooth(method = "lm", se = FALSE) + 
                  labs(x = "log (Mean)", y = "log (SD)", color = "Habitat") + 
                  scale_colour_manual(values = col)  +
                  facet_wrap(~type) + theme_classic() + theme(axis.text = element_text(size=14), axis.title = element_text(size = 16), 
                  legend.text = element_text(size = 14), legend.title = element_text(size = 16), strip.text = element_text(size = 16))

 p_mv <-  (mv1 / mv2)  + plot_annotation(tag_levels = "A") + theme(plot.tag = element_text(size = 20))
  ggsave(p_mv, file = "./output/figures/plot_mv.png", height = 8, width = 8)

  p_mv <- magick::image_read("./output/figures/plot_mv.png")
  p_mv
```

```{r, tbl-slopes}
#| label: tbl-slopes
#| tbl-cap: "Slopes and 95% credible intervals (lower = 2.5% and upper = 97.5%) of log transformed standard deviation (log(SD)) and log transformed mean (log(mean)) for each of the four treatment types (r1.1, r1.2, r2.1, r2.2). Note that r1.1 and r2.2 represent measurements of physiological rates of acclimated animals and measured at their respective acclimation temperature. In contrast, r1.2 and r2.1 are acute measurements. See Figure 1 in main manuscript for full details on treatments."

flextable(table) %>% bold(part = "header") %>% autofit()
```

#### *Comparing raw variance changes using* $lnVR_{Q_{10}}$

```{r, lnvrq10}
#| label: lnvrq10
#| echo: false
#| warning: false
#| error: false
#| include: false
#| cache: true

#############
  ## lnVRQ10
#############
##################################################
  # Organise the data and prune tree
#################################################
  lnVRQ10_data <- data_long2 %>% filter(grepl("lnVR", name) & !is.na(effect_size) & !is.na(V) & !V == 0 & !is.na(acclim_period)) %>% 
                              mutate(spp = species_full, c_accl_time = acclim_period - mean(acclim_period, na.rm = TRUE),
                                    max_temp_c = scale(temp_2, scale = FALSE))

# Check and prune tree
  tree_lnVR <- tree_checks(data = lnVRQ10_data, tree = tree_p0.7, dataCol = "species_full", type = "prune")
  tree_checks(data = lnVRQ10_data, tree = tree_lnVR, dataCol = "species_full", type = "check") # Good

# Create the phylo correlation matrix
  phylo_cor_VR <- vcv(tree_lnVR, cor = TRUE)
          A_VR <- vcv(tree_lnVR, cor = FALSE) # For bayes, need cov matrix

###################################################
  # Model 3.1: Intercept only model.
  # Used for heterogeneity analysis
###################################################
rerun = FALSE
if(rerun){
  model3.1 <- metafor::rma.mv(effect_size ~ 1, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_VR), dfs = "contain", 
                              test = "t", data = lnVRQ10_data)

  model3.1_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  1 + (~1 | record_num) + (1 | gr(species_full, cov = A_VR)) + (1 | spp) + (1|trait), data2 = list(A_VR=A_VR), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnVRQ10_data)

  # Save model1.1
  saveRDS(model3.1, file = "./output/models/lnVRQ10_3.1")
  saveRDS(model3.1_bayes, file = "./output/models/lnVRQ10_3.1_bayes")
} else{ 
        model3.1 <- readRDS("./output/models/lnVRQ10_3.1")
  model3.1_bayes <- readRDS("./output/models/lnVRQ10_3.1_bayes")
}

# Heterogeneity analysis
   w_VR <- with(lnVRQ10_data, 1/V) # weights
  sv_VR <- sum(w_VR*(dim(lnVRQ10_data)[1]-1)) / (sum(w_VR)^2 + sum(w_VR^2))

# extract posterior sds, but square them to turn to variance
  posteriors_3.1 <- posterior_samples(model3.1_bayes, pars = c("^sd", "sigma"))^2
  colnames(posteriors_3.1) <- gsub("__Intercept", "", colnames(posteriors_3.1))
  colnames(posteriors_3.1) <- gsub("sd_", "", colnames(posteriors_3.1))

    res_VR <- posteriors_3.1[,"sigma"]
   stdy_VR <- posteriors_3.1[,"record_num"]
  phylo_VR <- posteriors_3.1[,"species_full"]
    spp_VR <- posteriors_3.1[,"spp"]
  trait_VR <- posteriors_3.1[,"trait"]
  total_VR <- phylo_VR + spp_VR + res_VR + sv_VR + stdy_VR + trait_VR

  # Calculate I2
     tot_i2_VR <- (sv_VR / total_VR)*100    ; mean(tot_i2_VR)  ; quantile(tot_i2_VR, c(0.025, 0.975))
   phylo_i2_VR <- (phylo_VR / total_VR)*100 ; mean(phylo_i2_VR); quantile(phylo_i2_VR, c(0.025, 0.975))
     spp_i2_VR <- (spp_VR / total_VR)*100   ; mean(spp_i2_VR)  ; quantile(spp_i2_VR, c(0.025, 0.975))
    stdy_i2_VR <- (stdy_VR / total_VR)*100  ; mean(stdy_i2_VR) ; quantile(stdy_i2_VR, c(0.025, 0.975))
   trait_i2_VR <- (trait_VR / total_VR)*100 ; mean(trait_i2_VR); quantile(trait_i2_VR, c(0.025, 0.975))
     res_i2_VR <- (res_VR / total_VR)*100   ; mean(res_i2_VR)  ; quantile(res_i2_VR, c(0.025, 0.975))

  # Table
     i2_lnVR <- data.frame(name = c("Phylogeny", "Species", "Study", "Trait"),
                           est = c(mean(phylo_i2_VR),mean(spp_i2_VR), mean(stdy_i2_VR), mean(trait_i2_VR)),
                           lci = c(quantile(phylo_i2_VR, c(0.025, 0.975))[1], quantile(spp_i2_VR, c(0.025, 0.975))[1],
                                   quantile(stdy_i2_VR, c(0.025, 0.975))[1], quantile(trait_i2_VR, c(0.025, 0.975))[1]),
                           uci = c(quantile(phylo_i2_VR, c(0.025, 0.975))[2], quantile(spp_i2_VR, c(0.025, 0.975))[2],
                                   quantile(stdy_i2_VR, c(0.025, 0.975))[2], quantile(trait_i2_VR, c(0.025, 0.975))[2]))

###################################################
  # Model 3.2: Model controlling for acclimation time, 
  # type and habitat assuming type can vary by habitat
###################################################

if(rerun){
  # More realistic model 
  model3.2 <- metafor::rma.mv(effect_size ~  max_temp_c + c_accl_time + type + habitat + type:habitat, V = V, 
                              random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 + max_temp_c| spp,
                                            ~1 | trait,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_VR), dfs = "contain", 
                              test = "t", data = lnVRQ10_data)

   model3.2_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + type + habitat + type:habitat + (~1 | record_num) + (1 | gr(species_full, cov = A_VR)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_VR=A_VR), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnVRQ10_data)

  # Save model1.2
  saveRDS(model3.2, file = "./output/models/lnVRQ10_3.2")
  saveRDS(model3.2_bayes, file = "./output/models/lnVRQ10_3.2_bayes")

  } else{
        model3.2 <- readRDS("./output/models/lnVRQ10_3.2")
  model3.2_bayes <- readRDS("./output/models/lnVRQ10_3.2_bayes")
}

######################################################
  # Bayesian Contrasts
######################################################

  posteriors_3.2 <- posterior_samples(model3.2_bayes, pars = "^b")
  f_acute_VR  <- posteriors_3.2[,"b_Intercept"]; pmcmc(f_acute_VR)
  f_acclim_VR <- posteriors_3.2[,"b_Intercept"] + posteriors_3.2[,"b_typeacclim"]
  m_acute_VR  <- posteriors_3.2[,"b_Intercept"] + posteriors_3.2[,"b_habitatm"]
  m_acclim_VR <- posteriors_3.2[,"b_Intercept"] + posteriors_3.2[,"b_habitatm"] + posteriors_3.2[,"b_typeacclim"] + posteriors_3.2[,"b_typeacclim:habitatm"]
  t_acute_VR  <- posteriors_3.2[,"b_Intercept"] + posteriors_3.2[,"b_habitatt"]
  t_acclim_VR <- posteriors_3.2[,"b_Intercept"] + posteriors_3.2[,"b_habitatt"] + posteriors_3.2[,"b_typeacclim"] + posteriors_3.2[,"b_typeacclim:habitatt"]

# We'll take the marginalised mean for each type as there is no difference between acclimation and acute really except for in freshwater, but this effect size is so tiny. It's a 10% reduction at high temps and acclim is only a 2% reduction in variance at high temps.
  # First get sample sizes
    N <- lnVRQ10_data %>% group_by(habitat, type) %>% summarise(n=n())

  # Calculated weighted mean by sample size. 
    f_VR_mean <- (f_acute_VR * N[1,"n"][[1]] + f_acclim_VR * N[2,"n"][[1]]) / sum(N[1:2, "n"]); pmcmc(f_VR_mean)
    m_VR_mean <- (m_acute_VR * N[3,"n"][[1]] + m_acclim_VR * N[3,"n"][[1]]) / sum(N[3:4, "n"]); pmcmc(m_VR_mean)
    t_VR_mean <- (t_acute_VR * N[5,"n"][[1]] + t_acclim_VR * N[6,"n"][[1]]) / sum(N[5:6, "n"]); pmcmc(t_VR_mean)

  # Lets now convert these to a %
    f_VR_mean_ef  <- abs((exp(f_VR_mean) - 1)*100) 
    m_VR_mean_ef  <- abs((exp(m_VR_mean) - 1)*100) 
    t_VR_mean_ef  <- abs((exp(t_VR_mean) - 1)*100)

```

Analysis of $lnVR_{Q_{10}}$ suggested that variance increases with
higher temperatures across all habitat types, with terrestrial
ectotherms having the smallest increase in variance (@fig-s1 & @tbl-s1).

```{r, fig-s1}
#| label: fig-s1
#| fig-align: center
#| fig-cap: "Estimated mean acute and acclimation $lnVR_{Q_{10}}$ for marine, freshwater and terrestrial habitats. The percentage change in variance is also back calculated. Note that these are raw variances and do not account for changes in mean physiological rates. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\Beta$ values are the contrasts between acute and acclimation means within each habitat with 'NS' signifiying . $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets. $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."

k_vr <- manual_extract(model3.2, mod = "habitat", group = "spp", exclude = NULL)

##################
  # Figure S1
################## 
  # Sensitivity analysis that looks at how variance changes but not controlling for the mean.
  col <- c("#0871B9", "#F3B40D", "#1BB908")
  p3 <- orchard_plot(model3.2, xlab = TeX("$lnVR_{Q_{10}}$"), mod = "habitat", by = "type", group  = "record_num", weights = "prop", angle = 45, trunk.size = 1.8, branch.size = 2.8) +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + ylim(-2,3) +

     annotate("text", x = 1.5, y = -2, label =TeX(paste0("$\\mu = $", round(mean(f_VR_mean), 2), " [", round(quantile(f_VR_mean, 0.025), 2)," to ", round(quantile(f_VR_mean, 0.975),2), "] ", " $\\,p_{MCMC}$", p_value(round(pmcmc(f_VR_mean), digit = 2)))), hjust = 0, size = 8) + 
    annotate("text", x = 1.2, y = 1.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = 1.20, y = -2, label =paste0(round(mean(f_VR_mean_ef), 1), "%"), hjust = 0, size = 8) + labs(shape = "Type") +

    annotate("text", x = 2.5, y = -2, label =TeX(paste0("$\\mu = $", round(mean(m_VR_mean), 2), " [", round(quantile(m_VR_mean, 0.025), 2)," to ", round(quantile(m_VR_mean, 0.975),2), "] ", " $\\,p_{MCMC}$", p_value(round(pmcmc(m_VR_mean), digit = 2)))), hjust = 0, size = 8) + 
    annotate("text", x = 2.2, y = 1.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = 2.20, y = -2, label =paste0(round(mean(m_VR_mean_ef), 1), "%"), hjust = 0, size = 8) + labs(shape = "Type") +

        annotate("text", x = 3.5, y = -2, label =TeX(paste0("$\\mu = $", round(mean(t_VR_mean), 2), " [", round(quantile(t_VR_mean, 0.025), 2)," to ", round(quantile(t_VR_mean, 0.975),2), "] ", " $\\,p_{MCMC}$ = ", p_value(round(pmcmc(t_VR_mean), digit = 2)))), hjust = 0, size = 8) + 
    annotate("text", x = 3.2, y = 1.5, label =TeX(paste0("$\\beta = $", "NS")), hjust = 0, size = 8) +
  annotate("text", x = 3.20, y = -2, label =paste0(round(mean(t_VR_mean_ef), 1), "%"), hjust = 0, size = 8) + labs(shape = "Type") + scale_fill_manual(values = c(col)) + annotate("text", y = 2.5, x = c(1.3, 2.3, 3.3)+0.1, label = TeX(paste0("\\it{k} = ", k_vr$k, " (", k_vr$sdy, ")")), size = 6)

  p3 <- p3 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 30),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24), legend.position = "top")

  ggsave(p3, filename = "./output/figures/figS1.pdf", width = 11.5, height = 8)

  figs1 <- image_read("./output/figures/figS1.pdf")
  figs1
```

```{r, tbl-s1}
#| label: tbl-s1
#| echo: false
#| tbl-cap: "Model estimates, standard error, and 95% credible intervals comparing changes in acute and acclimation $lnVR_{Q_{10}}$ across habitat types. Model estimates are based off 1,253 effect sizes from 139 studies."

 # Extract the fixed / population-level effect summaries of the posterior
    fixed <- do.call(rbind, apply(posterior_samples(model3.2_bayes, pars = "^b"), 2,  function(x) post_summary(x)))
    fixed <- cbind(Parameter = c("Intercept", "Acclimation Time (z scaled)", "Acclimation Effect", "Habitat (Marine)", "Habitat (Terrestrial)", "Acclimation*Marine", "Acclimation*Terrestrial"), fixed)
    rownames(fixed) <- NULL

 # Extract the random effect summaries of the posteriors (SDs)
    random <- do.call(rbind, apply(posterior_samples(model3.2_bayes, pars = "^sd"), 2,  function(x) post_summary(x)))
    random <- cbind(Parameter = c( "Study", "Phylogeny", "Species", "Trait"), random)
    rownames(random) <- NULL

# Create the table
    tabs1 <- rbind(fixed, random)

    flextable(tabs1) %>% bold(part = "header") %>% autofit() %>% add_body_row(values = as_paragraph(as_b("Fixed Effects")), top = TRUE, colwidths = 5) %>% compose(value = as_paragraph(as_b(c(("Random Effects"), rep("", 4)))), i = 8)
```

#### *Plots of* $I^2$ for multilevel models

```{r, fig-s3}
#| label: fig-s3
#| fig-cap: "$I^2$ estimates. A) $lnRR_{Q_{10}}$ B) $lnCVR_{Q_{10}}$ and C) $lnVR_{Q_{10}}$." 
#| fig-align: center

ps3_1 <- ggplot(i2_lnrr, aes(x = est, y = name)) + xlim(0, 50) + geom_point(size = 5) + geom_errorbar(aes(xmin = lci, xmax = uci), width = 0.15) + theme_bw() + labs(x = "Percentage of Variation (%)") + theme(axis.title.y = element_blank(), axis.text = element_text(size = 20), axis.title = element_text(size = 24)) + geom_text(aes(label = paste0(round(est, 2), "% ", "[", round(lci, 2), " - ", round(uci, 2), "%]")), position = position_nudge(x = 5, y = 0.2)) + annotate("text", x = 30, y = 1.5, label = TeX(paste0("$I^2_{sv}$ = ", round(mean(tot_i2), 2), "%")), size = 10)

ps3_2 <- ggplot(i2_lncvr, aes(x = est, y = name)) + xlim(0, 50) + geom_point(size = 5) + geom_errorbar(aes(xmin = lci, xmax = uci), width = 0.15) + theme_bw() + labs(x = "Percentage of Variation (%)") + theme(axis.title.y = element_blank(), axis.text = element_blank(), axis.title = element_text(size = 24)) + geom_text(aes(label = paste0(round(est, 2), "% ", "[", round(lci, 2), " - ", round(uci, 2), "%]")), position = position_nudge(x = 5, y = 0.2)) + annotate("text", x = 30, y = 1.5, label = TeX(paste0("$I^2_{sv}$ = ", round(mean(tot_i2_cvr), 2), "%")), size = 10)

ps3_3 <- ggplot(i2_lnVR, aes(x = est, y = name)) + xlim(0, 50) + geom_point(size = 5) + geom_errorbar(aes(xmin = lci, xmax = uci), width = 0.15) + theme_bw() + labs(x = "Percentage of Variation (%)") + theme(axis.title.y = element_blank(), axis.text = element_blank(), axis.title = element_text(size = 24)) + geom_text(aes(label = paste0(round(est, 2), "% ", "[", round(lci, 2), " - ", round(uci, 2), "%]")), position = position_nudge(x = 5, y = 0.2)) + annotate("text", x = 30, y = 1.5, label = TeX(paste0("$I^2_{sv}$ = ", round(mean(tot_i2_cvr), 2), "%")), size = 10)

figs3 <- (ps3_1 | ps3_2 | ps3_3) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20))
ggsave(figs3, width = 16, height = 5.7, filename = "./output/figures/figs3.png")

figs3 <- image_read("./output/figures/figs3.png")
figs3
```

#### *Environmental predictability*

Theoretical models highlight the importance of environmental
predictability in selecting for plastic responses. However, capturing
environmental predictability is challenging given that it is unclear
which timescale one should select. For example, is it more important to
look at correlation between temperatures monthly or seasonally. In
addition, such temporal resolution will likely depend on the species in
question given that for some species fine-grained thermal predictability
maybe more important compared to others.

With these limitations in mind, we used our temperature time series to
calculate auto regressive correlation in temperature across the entire
time series. We then modeled how this measure of thermal predictability
was related to plasticity.

```{r, acf}
#| label: acf
#| echo: false
#| warning: false
#| error: false
#| include: false
#| cache: true
############################################################################################
# Model 3.8: Acclimatization lnRRQ10 only. Test if environmental predictability explains 
# differently in each habitat
############################################################################################  
  lnRRQ10_data_wild_acclim <- lnRRQ10_data_wild %>%  filter(type == "acclim") %>% mutate(acf_lag1_c = scale(acf_lag1))

  # Create the phylo correlation matrix

  # Check and prune tree
tree_lnrr_wild_acclim <- tree_checks(data = lnRRQ10_data_wild_acclim, tree = tree_p0.7, dataCol = "species_full", type = "prune")
tree_checks(data = lnRRQ10_data_wild_acclim, tree = tree_lnrr_wild_acclim, dataCol = "species_full", type = "check")

phylo_cor_wild_acclim <- vcv(tree_lnrr_wild_acclim, cor = TRUE)
        A_wild_acclim <- vcv(tree_lnrr_wild_acclim, cor = FALSE) # For bayes, need cov matrix

if(rerun){
  # More realistic model with 
  model3.8  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c + habitat + habitat:acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim)

   model3.8_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + acf_lag1_c + habitat + habitat:acf_lag1_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_acclim)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_wild_acclim=A_wild_acclim), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data_wild_acclim)

  # Save model3.8
  saveRDS(model3.8, file = "./output/models/lnRRQ10_3.8")
  saveRDS(model3.8_bayes, file = "./output/models/lnRRQ10_3.8_bayes")

  } else{
        model3.8 <- readRDS("./output/models/lnRRQ10_3.8")
  model3.8_bayes <- readRDS("./output/models/lnRRQ10_3.8_bayes")
}

#################################################
### Bayesian Slopes from Model 3.8
#################################################
  posteriors3.8 <- posterior_samples(model3.8_bayes, "^b")

    slope_f_3.8 <- posteriors3.8[,"b_acf_lag1_c"] 
                    pmcmc(slope_f_3.8); quantile(slope_f_3.8, c(0.025, 0.975))
    slope_m_3.8 <- posteriors3.8[,"b_acf_lag1_c"] + posteriors3.8[,"b_acf_lag1_c:habitatm"] 
                    pmcmc(slope_m_3.8); quantile(slope_m_3.8, c(0.025, 0.975))
    slope_t_3.8 <- posteriors3.8[,"b_acf_lag1_c"] + posteriors3.8[,"b_acf_lag1_c:habitatt"] 
                    pmcmc(slope_t_3.8); quantile(slope_t_3.8, c(0.025, 0.975))

##################################################
    # Model 3.8, 3.9 and 3.10 - Habitat subset
##################################################
    # Subset data to habitat
    lnRRQ10_data_wild_acclim_t <- lnRRQ10_data_wild_acclim %>% filter(habitat == "t")
    lnRRQ10_data_wild_acclim_f <- lnRRQ10_data_wild_acclim %>% filter(habitat == "f")
    lnRRQ10_data_wild_acclim_m <- lnRRQ10_data_wild_acclim %>% filter(habitat == "m")

if(rerun){
    # Terrestrial Models
    model3.8_t  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim_t)
    saveRDS(model3.8_t, "./output/models/model3.8_t")

    # Freshwater Models
    model3.8_f  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim_f)
    saveRDS(model3.8_f, "./output/models/model3.8_f")

    # Marine Models
      model3.8_m  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnRRQ10_data_wild_acclim_m)
        saveRDS(model3.8_m, "./output/models/model3.8_m")

  } else {
     model3.8_t <- readRDS("./output/models/model3.8_t")
     model3.8_f <- readRDS("./output/models/model3.8_f")
     model3.8_m <- readRDS("./output/models/model3.8_m")
}

############################################################################################
  # Model 4.8: Acclimatization lnCVRQ10 only. Test if environmental predictability explains 
  # differently in each habitat
############################################################################################

if(rerun){
  # More realistic model with 
  model4.8  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c + habitat + habitat:acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim)

   model4.8_bayes <- brms::brm(effect_size | se(sqrt(V), sigma = TRUE) ~  max_temp_c + c_accl_time + acf_lag1_c + habitat + habitat:acf_lag1_c + (~1 | record_num) + (1 | gr(species_full, cov = A_wild_acclim)) + (1 + max_temp_c| spp) + (1|trait), data2 = list(A_wild_acclim=A_wild_acclim), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnCVRQ10_data_wild_acclim)

  # Save model4.8
  saveRDS(model4.8, file = "./output/models/lnRRQ10_4.8")
  saveRDS(model4.8_bayes, file = "./output/models/lnRRQ10_4.8_bayes")

  } else{
        model4.8 <- readRDS("./output/models/lnRRQ10_4.8")
  model4.8_bayes <- readRDS("./output/models/lnRRQ10_4.8_bayes")
}

#################################################
### Bayesian Slopes from Model 4.8
#################################################
 posteriors4.8 <- posterior_samples(model4.8_bayes, "^b")

    slope_f_4.8 <- posteriors4.8[,"b_acf_lag1_c"]; mean(slope_f_4.8); pmcmc(slope_f_4.8); quantile(slope_f_4.8, c(0.025, 0.975))
    slope_m_4.8 <- posteriors4.8[,"b_acf_lag1_c"] + posteriors4.8[,"b_acf_lag1_c:habitatm"]; mean(slope_m_4.8); pmcmc(slope_m_4.8); quantile(slope_m_4.8, c(0.025, 0.975))
    slope_t_4.8 <- posteriors4.8[,"b_acf_lag1_c"] + posteriors4.8[,"b_acf_lag1_c:habitatt"]; mean(slope_t_4.8); pmcmc(slope_t_4.8); quantile(slope_t_4.8, c(0.025, 0.975))

#################################################
### Same model as above, but seperate for habitat 
### For plotting, 4.8
#################################################
 lnCVRQ10_data_wild_acclim_t <- lnCVRQ10_data_wild_acclim %>% filter(habitat == "t")
 lnCVRQ10_data_wild_acclim_m <- lnCVRQ10_data_wild_acclim %>% filter(habitat == "m")
 lnCVRQ10_data_wild_acclim_f <- lnCVRQ10_data_wild_acclim %>% filter(habitat == "f")

if(rerun){
  # Terrestrial Model
    model4.8_t  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim_t)
    saveRDS(model4.8_t, "./output/models/model4.8_t")

  # Freshwater Model
    model4.8_f  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim_f)
    saveRDS(model4.8_f, "./output/models/model4.8_f")

  # Marine Model
    model4.8_m  <- metafor::rma.mv(effect_size ~ max_temp_c + c_accl_time + acf_lag1_c, V = V,
                                       random = list(~1 | record_num,
                                            ~1 | species_full,
                                            ~1 | trait,
                                            ~1 + max_temp_c| spp,
                                            ~1 | obs),
                              R = list(species_full = phylo_cor_wild_acclim), dfs = "contain", 
                              test = "t", data = lnCVRQ10_data_wild_acclim_m)
    saveRDS(model4.8_m, "./output/models/model4.8_m")
} else {
     model4.8_t <- readRDS("./output/models/model4.8_t")
     model4.8_m <- readRDS("./output/models/model4.8_m")
     model4.8_f <- readRDS("./output/models/model4.8_f")
}

```

```{r, fig-4}
#| label: fig-4
#| fig-align: center
#| fig-cap: "Predicted mean acclimation (thick black line) $lnRR_{Q_{10_{acclim}}}$ (A) and $lnCVR_{Q_{10_{acclim}}}$ (B) as a function of the Thermal Predictability for wild populations across marine, freshwater and terrestrial habitats. Dashed lines indicate 95% confidence intervals and dotted lines indicate 95% prediction intervals. Model slope ($\\beta$) along with the 95% CI and $p_{MCMC}$ values for the slopes are shown for each habitat."

size_axis = 20
col <- c("#0871B9", "#F3B40D", "#1BB908")

####################
# lnRR ACF
####################
 dat_textacf <- data.frame(label = c(text(slope_f_3.8), 
                                     text(slope_m_3.8), 
                                     text(slope_t_3.8)),
                          label2 = c(text(slope_f_3.8, b = FALSE), 
                                     text(slope_m_3.8, b = FALSE), 
                                     text(slope_t_3.8, b = FALSE)),
                               p = c(p_text(slope_f_3.8), 
                                     p_text(slope_m_3.8), 
                                     p_text(slope_t_3.8)),
                           title = c("Freshwater", "Marine", "Terrestrial"), 
                       condition = c("f", "m", "t"),
                               x = rep(0, 3))
  y_adj = -3
#############
# Terrestrial
#############
  p_t_rr_acf <-  bubble_plot(model3.8_t, mod = "acf_lag1_c", group = "spp", 
                             ylab = TeX("$lnRR_{Q_{10_{acclim}}}$"), 
                             xlab = "Thermal Predictability", g = TRUE) + 
    geom_point(data = lnRRQ10_data_wild_acclim_t, 
               aes(x = acf_lag1_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[3], colour = "black", stroke = 0.2, shape = 21) + 
    labs(title = dat_textacf$title[3]) +
    annotate("text", x = dat_textacf$x[3], y = y_adj, label = TeX(dat_textacf$p[3]), size = 8) +
    annotate("text", x = dat_textacf$x[3], y = y_adj + 0.5, label = TeX(dat_text$label2[3]), size = 8) +
    annotate("text", x = dat_textacf$x[3], y = y_adj + 1, label = TeX(dat_textacf$label[3]), size = 8) +
    theme( axis.text = element_text(size = size_axis), 
         axis.text.y = element_text(size = size_axis), 
          axis.title = element_text(size = size_axis),
        axis.title.x = element_blank(),
          plot.title = element_text( size = 36, hjust = 0)) + 
    ylim(-3,3)
#############
# Marine
#############
  x_m = -2.5
p_m_rr_acf <-  bubble_plot(model3.8_m, mod = "acf_lag1_c", group = "spp", 
                           ylab = TeX("$lnRR_{Q_{10_{acclim}}}"), 
                           xlab = "Thermal Predictability", g = TRUE) + 
    geom_point(data = lnRRQ10_data_wild_acclim_m, 
               aes(x = acf_lag1_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[2], colour = "black", stroke = 0.2, shape = 21) +  
    labs(title = dat_textacf$title[2]) +
    annotate("text", x = dat_textacf$x[2]+x_m, y = y_adj,     label = TeX(dat_textacf$p[2]),      size = 8) +
    annotate("text", x = dat_textacf$x[2]+x_m, y = y_adj+0.5, label = TeX(dat_textacf$label2[2]), size = 8) +
    annotate("text", x = dat_textacf$x[2]+x_m, y = y_adj+1,   label = TeX(dat_textacf$label[2]),  size = 8) +
    theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
         axis.title = element_text(size = size_axis), 
       axis.title.x = element_blank(),
         plot.title = element_text( size = 36, hjust = 0)) + 
    ylim(-3, 3)

#############
# Freshwater
#############
x_f = -0.2
p_f_rr_acf <-  bubble_plot(model3.8_f, mod = "acf_lag1_c", group = "spp", 
                           ylab = TeX("$lnRR_{Q_{10_{acclim}}}"), 
                           xlab = "Thermal Predictability", g = TRUE) + 
    geom_point(data = lnRRQ10_data_wild_acclim_f, 
               aes(x = acf_lag1_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[1], colour = "black", stroke = 0.2, shape = 21) +  
    labs(title = dat_textacf$title[1]) +
    annotate("text", x = dat_textacf$x[1]+x_f, y = y_adj,     label = TeX(dat_textacf$p[1]),      size = 8) +
    annotate("text", x = dat_textacf$x[1]+x_f, y = y_adj+0.5, label = TeX(dat_textacf$label2[1]), size = 8) +
    annotate("text", x = dat_textacf$x[1]+x_f, y = y_adj+ 1,  label = TeX(dat_textacf$label[1]),  size = 8) +
    theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
         axis.title = element_text(size = size_axis), 
       axis.title.x = element_blank(),
         plot.title = element_text( size = 36, hjust = 0)) + 
    ylim(-3, 3)

####################
# lnCVR CV
####################
dat_textcvr_acf <- data.frame(label = c(text(slope_f_4.8), 
                                 text(slope_m_4.8), 
                                 text(slope_t_4.8)),
                      label2 = c(text(slope_f_4.8, b = FALSE), 
                                 text(slope_m_4.8,b = FALSE), 
                                 text(slope_t_4.8,b = FALSE)),
                           p = c(p_text(slope_f_4.8), 
                                 p_text(slope_m_4.8), 
                                 p_text(slope_t_4.8)),
                       title = c("Freshwater", "Marine", "Terrestrial"), 
                      condition = c("f", "m", "t"),
                      x = rep(0, 3),
                      y = rep(-4, 3))
#############
# Freshwater
#############
  yadj2 = -6
  x_f = -0.2
  p_f_cv_acf <-  bubble_plot(model4.8_f, mod = "acf_lag1_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), 
                              xlab = "Thermal Predictability", g =FALSE, k = FALSE) + 
      geom_point(data = lnCVRQ10_data_wild_acclim_f, 
                 aes(x = acf_lag1_c, y = effect_size, size = 1/sqrt(V)), 
                 fill = col[1], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textcvr_acf$x[1]+x_f, y = yadj2, label = TeX(dat_textcvr_acf$p[1]), size = 8) +
    annotate("text", x = dat_textcvr_acf$x[1]+x_f, y = yadj2+0.8, label = TeX(dat_textcvr_acf$label2[1]), size = 8) +
    annotate("text", x = dat_textcvr_acf$x[1]+x_f, y = yadj2+1.6, label = TeX(dat_textcvr_acf$label[1]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)
#############
# Marine
#############
  x_m = -2.5
  p_m_cv_acf <-  bubble_plot(model4.8_m, mod = "acf_lag1_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), 
                              xlab = "Thermal Predictability", g =FALSE, k = FALSE) + 
    geom_point(data = lnCVRQ10_data_wild_acclim_m, 
                 aes(x = acf_lag1_c, y = effect_size, size = 1/sqrt(V)), 
                 fill = col[2], colour = "black", stroke = 0.2, shape = 21) + 
    annotate("text", x = dat_textcvr_acf$x[1]+x_m, y = yadj2, label = TeX(dat_textcvr_acf$p[2]), size = 8) +
    annotate("text", x = dat_textcvr_acf$x[1]+x_m, y = yadj2+0.8, label = TeX(dat_textcvr_acf$label2[2]), size = 8) +
    annotate("text", x = dat_textcvr_acf$x[1]+x_m, y = yadj2+1.6, label = TeX(dat_textcvr_acf$label[2]), size = 8) +
    theme(axis.text = element_text(size = size_axis), 
        axis.text.y = element_text(size = size_axis), 
        axis.title = element_text(size = size_axis), 
        plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
    ylim(-6, 3.5)
#############
# Terrestrial
#############
x_t = -0.2
p_t_cv_acf <-  bubble_plot(model4.8_t, mod = "acf_lag1_c", group = "spp", ylab = TeX("$lnCVR_{Q_{10_{acclim}}}$"), 
                            xlab = "Thermal Predictability", g = FALSE, k = FALSE) + 
  geom_point(data = lnCVRQ10_data_wild_acclim_t, 
               aes(x = acf_lag1_c, y = effect_size, size = 1/sqrt(V)), 
               fill = col[3], colour = "black", stroke = 0.2, shape = 21) +
  annotate("text", x = dat_textcvr_acf$x[3]+x_t, y = yadj2, label = TeX(dat_textcvr_acf$p[3]), size = 8) +
  annotate("text", x = dat_textcvr_acf$x[3]+x_t, y = yadj2+0.8, label = TeX(dat_textcvr_acf$label2[3]), size = 8) +
  annotate("text", x = dat_textcvr_acf$x[3]+x_t, y = yadj2+1.6, label = TeX(dat_textcvr_acf$label[3]), size = 8) +
  theme(axis.text = element_text(size = size_axis), 
      axis.text.y = element_text(size = size_axis), 
      axis.title = element_text(size = size_axis), 
      plot.title = element_text(face = "bold", size = 36, hjust = 0.5)) + 
  ylim(-6, 3.5)

####################
# Full plot
####################

p_acf <- (p_t_rr_acf + p_m_rr_acf + p_f_rr_acf) / (p_t_cv_acf + p_m_cv_acf + p_f_cv_acf) + plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 36))

ggsave(p_acf, width = 18, height = 10, filename = "./output/figures/fig_p_acf_r1.pdf")

  p_acf <- image_read("./output/figures/fig_p_acf_r1.pdf")
  p_acf 
```

#### *Publication bias analysis*

```{r, pubias}
#| label: pubias
#| echo: false
#| include: false
rerun = FALSE
if(rerun){
  # Use effective sample size instead because sampling variance estimates could induce a spurious correlation with our effect size
    lnRRQ10_data <- lnRRQ10_data %>%  
              mutate(n_ef = ifelse(name == "lnRR_Q10_acute_t1", (r1.1_N + r1.2_N) / (r1.1_N*r1.2_N), # acute 1
                            ifelse(name == "lnRR_Q10_acute_t2", (r2.1_N + r2.2_N) / (r2.1_N*r2.2_N), # acute 2
                                                                (r1.1_N + r2.2_N) / (r1.1_N*r2.2_N)))) # acclimation

    lnCVRQ10_data <- lnCVRQ10_data %>%  
               mutate(n_ef = ifelse(name == "lnRR_Q10_acute_t1", (r1.1_N + r1.2_N) / (r1.1_N*r1.2_N), # acute 1
                            ifelse(name == "lnRR_Q10_acute_t2", (r2.1_N + r2.2_N) / (r2.1_N*r2.2_N), # acute 2
                                                                (r1.1_N + r2.2_N) / (r1.1_N*r2.2_N)))) #acclimation

    lnVRQ10_data <- lnVRQ10_data %>%  
              mutate(n_ef = ifelse(name == "lnRR_Q10_acute_t1", (r1.1_N + r1.2_N) / (r1.1_N*r1.2_N), # acute 1
                            ifelse(name == "lnRR_Q10_acute_t2", (r2.1_N + r2.2_N) / (r2.1_N*r2.2_N), # acute 2
                                                                (r1.1_N + r2.2_N) / (r1.1_N*r2.2_N)))) #acclimation

    model_full_lnrr_pb <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + sqrt(n_ef) + habitat*life.hist*type + trait_category + habitat:type + habitat:life.hist + habitat:life.hist:type + (~1 | record_num) + (1 | gr(species_full, cov = A)) + (1 + max_temp_c | spp) + (1 | trait) + (1  | obs),  data2 = list(A=A), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 8, data = lnRRQ10_data)

    model_full_lncvr_pb <- brms::brm(effect_size | se(sqrt(V)) ~ max_temp_c +  c_accl_time + sqrt(n_ef) + habitat*life.hist*type + trait_category + (~1 | record_num) + (1 | gr(species_full, cov = A_CVR)) + (1 + max_temp_c| spp) + (1 | trait) + (1  | obs),  data2 = list(A_CVR=A_CVR), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 1, iter = 5000, warmup = 1000, cores = 8, data = lnCVRQ10_data)

    model_full_lnvr_pb <- brms::brm(effect_size | se(sqrt(V)) ~  max_temp_c + c_accl_time + sqrt(n_ef) + habitat*life.hist*type + trait_category + (~1 | record_num) + (1 | gr(species_full, cov = A_VR)) + (1 + max_temp_c | spp) + (1 | trait) + (1  | obs),  data2 = list(A_VR=A_VR), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 1, iter = 5000, warmup = 1000, cores = 8, data = lnVRQ10_data)

  saveRDS(model_full_lnrr_pb, file = "./output/models/lnRRQ10_pubbias")
  saveRDS(model_full_lncvr_pb, file = "./output/models/lnCVRQ10_pubbias")
  saveRDS(model_full_lnvr_pb, file = "./output/models/lnVRQ10_pubbias")
} else {
   model_full_lnrr_pb <- readRDS(file = "./output/models/lnRRQ10_pubbias")
  model_full_lncvr_pb <- readRDS(file = "./output/models/lnCVRQ10_pubbias")
  model_full_lnvr_pb <- readRDS(file = "./output/models/lnVRQ10_pubbias")
}

 #Pull the posterior for the slope with sqrt effective n
  post_pb_lnrr <- posterior_samples(model_full_lnrr_pb, "^b_sqrtn_ef")[,1]
 #Pull the posterior for the slope with sqrt effective n
  post_pb_lncvr <- posterior_samples(model_full_lncvr_pb, "^b_sqrtn_ef")[,1]
 #Pull the posterior for the slope with sqrt effective n
  post_full_lnvr_pb <- posterior_samples(model_full_lnvr_pb, "^b_sqrtn_ef")[,1]
```

```{r, fig-s2}
#| label: fig-s2
#| fig-cap: "Funnel plot of precision (1/sampling standard error) against effect size for A) log response ratio $Q_{10}$ ($lnRR_{Q_{10}}$), B) log coefficient of variance ratio $Q_{10}$ ($lnCVR_{Q_{10}}$) and C) log variance ratio $Q_{10}$ ($lnVR_{Q_{10}}$). Both acute ('black') and acclimation ('grey') effect sizes are plotted."
cols <- c("#303234", "#D3D3D3")
ps2_1 <- ggplot(lnRRQ10_data, aes(y = 1/sqrt(V), x = effect_size, color = type)) + geom_point() + theme_classic() + labs(y = TeX("Precision\\ $\\left( \\frac{1}{\\sqrt{V}} \\right)$"), x = TeX("$lnRR_{Q_{10}}$"))+geom_vline(xintercept = 0, linetype = 3) + theme(legend.title  = element_text(size = 15), legend.text = element_text(size = 15), axis.title = element_text(size = 15), axis.text = element_text(size = 15), legend.position = "none") + scale_color_manual(name = "Type", values = cols)

ps2_2 <- ggplot(lnCVRQ10_data, aes(y = 1/sqrt(V), x = effect_size, color = type)) + geom_point() + theme_classic() + labs(y = TeX("Precision\\ $\\left( \\frac{1}{\\sqrt{V}} \\right)$"), x = TeX("$lnCVR_{Q_{10}}$"))+geom_vline(xintercept = 0, linetype = 3) + theme(legend.title  = element_text(size = 15), legend.text = element_text(size = 15), axis.title = element_text(size = 15), axis.text = element_text(size = 15), legend.position = "none", axis.title.y = element_blank())+ scale_color_manual(name = "Type", values = cols)

ps2_3 <- ggplot(lnVRQ10_data, aes(y = 1/sqrt(V), x = effect_size, color = type)) + geom_point() + theme_classic() + labs(y = TeX("Precision\\ $\\left( \\frac{1}{\\sqrt{V}} \\right)$"), x = TeX("$lnVR_{Q_{10}}$"))+geom_vline(xintercept = 0, linetype = 3) + theme(legend.title  = element_text(size = 15), legend.text = element_text(size = 15), axis.title = element_text(size = 15), axis.text = element_text(size = 15),  axis.title.y = element_blank())+ scale_color_manual(name = "Type", values = cols)

figs2 <- (ps2_1 | ps2_2 | ps2_3) + plot_annotation(tag_levels = "A", tag_suffix = ")") & theme(plot.tag = element_text(size = 20))
ggsave(figs2, width = 11.5, height = 4.8, filename = "./output/figures/figS2.png")

figs2 <- image_read("./output/figures/figS2.png")
figs2
```

Funnel plots did not show any noticeable deviation from the typical
funnel shape for any of the effect size estimates (@fig-s2).
Meta-regression models including sampling standard error as a moderator
also suggested no relationship with effect size for $lnRR_{Q_{10}}$
($\beta$ = `r mean(post_pb_lnrr)`, 95% CI:
`r quantile(post_pb_lnrr, 0.025)` to `r quantile(post_pb_lnrr, 0.975)`,
$p_{MCMC}$ = `r p_value(pmcmc(post_pb_lnrr))`), $lnCVR_{Q_{10}}$
($\beta$ = `r mean(post_pb_lncvr)`, 95% CI:
`r quantile(post_pb_lncvr, 0.025)` to
`r quantile(post_pb_lncvr, 0.975)`, $p_{MCMC}$ =
`r p_value(pmcmc(post_pb_lncvr))`) or $lnVR_{Q_{10}}$ ($\beta$ =
`r mean(post_full_lnvr_pb)`, 95% CI:
`r quantile(post_full_lnvr_pb, 0.025)` to
`r quantile(post_full_lnvr_pb, 0.975)`, $p_{MCMC}$ =
`r p_value(pmcmc(post_full_lnvr_pb))`) was not significant indicating
little evidence for publication bias.